### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_1.Rnw
package skeleton for yuimadoc package, not yet working
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@11 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Introduction}

\def\yuima{{\bf yuima}}

The package $\yuima$ provides 
an object-oriented programming environment 
for simulation and statistical inference 
for stochastic processes by R. 
On this framework, 
$\yuima$ also supplies various functions 
to execute simulation and statistical analysis. 
Both categories of procedures may depend each other. 
Statistical inference often requires a simulation technique 
as a subroutine, and a certain simulation method 
needs to fix a tuning parameter by applying 
a statistical methodology. 
It is especially the case of stochastic processes 
because most of expected values involved 
do not admit an explicit expression. 
%It is therefore natural to 
$\yuima$ facilitates comprehensive, systematic approaches 
to the solution.

Stochastic differential equations are 
commonly used 
to model random evolution along continuous or 
practically continuous time, such as 
the random movements of a stock price. 
Theory of statistical inference for 
stochastic differential equations already has a fairly long history, 
more than three decades, but it is still developing quickly new 
methodologies and expanding the area. 
The formulas produced by the theory are usually very sophisticated, 
which makes it difficult for standard users not necessarily 
familiar with this field to enjoy utilities. 
For example, the asymptotic expansion method for computing 
option prices (i.e., expectation of an irregular functional of 
a stochastic process) provides precise approximation values 
instantaneously, taking advantage of the analytic approach, 
but the formula has a long expression like more than one page! 
$\yuima$ delivers up-to-date methods as a package 
onto the desk of the user working 
with simulation and/or statistics for stochastic differential equations.

Sampled data from a continuous-time process features 
the time stamps as well as the positions of the object. 
It is requiring a new theory of estimation. 
$\yuima$ framework can apply multi-dimensional time stamps 
of tick data and provides diverse functions handling such kind data 
to support statistical analysis.

### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_2.Rnw
package skeleton for yuimadoc package, not yet working
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@11 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Yuima object}
The main object in our framework is the \code{yuima} object.
The \code{yuima} object contains slots named \code{model}, \code{data},
\code{sampling}, and \code{characteristic}.

\begin{center}
\begin{tabular}{|c|}
\hline
\colorbox{black}{\textcolor{white}{  \code{yuima} \bf object}}\\
\hline
\hline
\code{model}\\
\hline
\code{data}\\
\hline
\code{sampling}\\
\hline
\code{characteristic}\\
\hline
\end{tabular}
\end{center}

The slot \code{data} will contain
a time series data and the slot \code{sampling} contains a description of
sampling methods and/or parameters of the data collection(e.g. poisson sampling with rate $p=0.1$.).

The \code{yuima.model} object is intended to be a description of model 
independent of the actual data. 
For example, in model selection or hypothesis testing for SDEs, 
it is possible to have different models to fit to the same set of data.

On the contrary, for the same model specified in \code{yuima.model}, 
different time series can be fitted. 
In this case the vector \code{model} has length 1, and 
the vector of \code{data} has length greater than 1.

The \code{yuima.characteristic} object is prepared for specifying the characteristics of the \code{yuima} object such as
the number of equations in the model and data, time and space unit scale for simulation or data.
For now, this slot is mainly for keeping consistency of the other three slots.
%For example, the number of pathes specified in the \code{yuima.model} object and the actual data in \code{yuima.data} is 
%not the same,

\code{setYuima} is a constructor for \code{yuima} class, 
which does not initialize but only creates four NULL slots:

\begin{itemize}
\item{data:}{ an object of class \code{yuima.data}}
\item{model:}{ an object of class \code{yuima.model}}
\item{sampling:}{ an object of class \code{yuima.sampling}}
\item{characteristic:}{ an object of class \code{yuima.characteristic}}
\end{itemize}
%We show how to construct \code{yuima} object in the next section.

%%Note that \code{yuima.sampling} is not implemented yet.

%-----------------------------------
 \begin{figure}[!htb]
  \centering
  \includegraphics[width=10cm,clip]{./EPS/functionDefinition.eps}
  \caption{function relation}
 \end{figure}
 %-----------------------------------
%-----------------------------------
 \begin{figure}[!htb]
  \centering
  \includegraphics[width=5cm,clip]{./EPS/classDefinition.eps}
  \caption{class definition}
 \end{figure}
 %-----------------------------------

\subsection{yuima.model}
\code{yuima.model} specifies the SDE of interest. The constructor \code{setModel} is
used to give a mathematical description of the stochastic differential equation. 
All functions in the package are assumed to get as much information
as possible from the model instead of replicating the same code everywhere.
If there are missing pieces of information, we may change or extend the description
of the model.
\\
\\
An object of \code{yuima.model} contains several slots listed below. 
To see inside its structure, we use the \R{} command \code{str}.
\begin{itemize}
\item \code{drift} is an \R{} expression which contains the drift specification.
\item \code{diffusion} is itself a list of 1 slot which describes the diffusion
  coefficient relative to first noise.

\item \code{parameter} which is a short name for ``parameters'' which 
  is a list of objects.

\item \code{all} contains the names of all the parameters found 
   in the diffusion and drift coefficient.
\item \code{common} contains the names of the parameters in common between the drift and diffusion coefficients.
\item \code{diffusion} contains the parameters belonging to the diffusion coefficient.
\item \code{drift} contains the parameters belonging to the drift coefficient.
\item \code{solve.variable} contains a vector of variable names, each element corresponds to the
   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order.

\item \code{state.variable} and \code{time.variable}, by default,
  are assumed to be $x$ and $t$ but the user can freely choose them.
  The \code{yuima.model} function assumes that the user either use default
  names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
  the rest of the symbols are considered parameters and distributed accordingly
  in the \code{parameter} slot.

\item \code{noise.number} indicates the number of sources of noise.

\item \code{equation.number} represents the number of equations, i.e. the number of one dimensional
  stochastic differential equations.

\item \code{dimension} reports the dimensions of the parameter space. It is a list of the
  same length of \code{parameter} with the same names.

\end{itemize}
We show an example to describe the following simple stochastic differential equation

\begin{equation}
\mbox{d}X_t=3X_t\mbox{d}t+\mbox{d}W_t.
\end{equation}

<<print=FALSE,echo=TRUE>>=
library(yuima)
@

<<echo=TRUE>>=
mod <- setModel(drift="3*x")
str(mod)
@

\subsection{yuima.data}
\code{yuima.data} is a class to represent the data to be used. The constructor \code{setData} is
used to give an observed data to the \code{yuima.data} object. \code{get.zoo.data} returns the \code{zoo.data} field of 
an \code{yuima.data} object.
As usual class methods, \code{yuima.data} provides \code{length}, \code{plot} and 
\code{dim} functions to get necessary information.

 The \code{zoo.data} slot may contain different kind of observed data. 
 In the \R{} system, there are different kinds of objects to describe time series data.
 Here is a very short list of possibilities
 \begin{center}
 \begin{tabular}{r l}
 \code{ts}, \code{mts}: & regularly spaced time series 1-dim or multidim\\
 \code{its}: & irregularly spaced time series\\
 \code{tseries}: & time index is real time of financial markets \\
 \code{zoo}: & more flexibility  in the management of the index of the time series.\\ 
 \end{tabular}
 \end{center}
 Because \R{} is used in many contexts, including finance, and users adopt one of 
 the above times series objects because they are used to it, 
 our framework accepts any of the above as observed data and treat the inference accordingly.
 \\
 \\
 An object of \code{yuima.data} contains several slots listed below. 
 To see inside its structure we use the \R{} command \code{str}.
 \begin{itemize}
 \item \code{original.data} is a copy of the original data. (it can be any type of data.)
 \item \code{zoo.data} is an bundled \code{zoo} version object of class \code{yuima.data} or \code{yuima}.
 \end{itemize}

 We show an example of \code{yuima.data} object.

<<print=FALSE,echo=TRUE>>=
library(zoo)
@

<<fig=TRUE,echo=TRUE,width=7,height=7,results=hide>>=
X <- ts(matrix(rnorm(200),100,2))
mydata <- setData(X)
str(get.zoo.data(mydata))
#plot(get.zoo.data(mydata))
dim(mydata)
length(mydata)
plot(mydata)
@

Then, we show how to construct \code{yuima} object with \code{yuima.data} and \code{yuima.model} object.

We first construct a two-dimensional homogeneous diffusion process:
\begin{eqnarray*}
\mbox{d} X_{1t} &=&
- 3 X_{1t} \mbox{d}t + \mbox{d} W_{1t}, \\
\mbox{d}X_{2t} &=&
-(X_{1t} + 2 X_{2t}) \mbox{d}t + 0.5 \mbox{d} W_{1t} + \mbox{d}W_{2t}.
\end{eqnarray*}
<<echo=TRUE>>=
# we set the drift coefficient
a <- c("-3*x1","-x1-2*x2")
# and also the diffusion coefficient
b <- matrix(c("1","0.5","0","1"),2,2)
# Then set
mod2 <- setModel(drift = a, diffusion = b, solve.variable = c("x1","x2"))
str(mod2)
@

Then, we can combine the \code{yuima.model} and \code{yuima.data}
objects to a \code{yuima} object.
<<echo=TRUE>>=
yuima.obj <- setYuima(data=mydata,model=mod2)
str(yuima.obj)
@

\subsection{Realworld data}
In this subsection, we show a simple example of handling the realworld data in
the yuima framework.
<<fig=TRUE,echo=TRUE,width=7,height=7>>=
## Use TTR package to get real market data from Yahoo! Finance
require(TTR)
sbux<-getYahooData("SBUX", start=20080401, end=20090331)
cbou<-getYahooData("CBOU", start=20080401, end=20090331)
sd <- setData(list(sbux,cbou))
real.yuima <- setYuima(data=sd)
summary(real.yuima)
head(get.zoo.data(real.yuima)[[1]])
head(get.zoo.data(real.yuima)[[2]])

plot(get.zoo.data(real.yuima)[[1]]$High,main="SBUX_High")
@

### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_3.Rnw
package skeleton for yuimadoc package, not yet working
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@11 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Simulation}
\subsection{Diffusion Process}
\code{simulate} is a function to solve SDE using the Euler-Maruyama method.
Here we illustrate how to generate a simulated path of the solution of
a given SDE. Let us consider a 3-dim SDE

 \begin{equation} \label{sabr}
 \begin{cases}
 & \mbox{d} X_t^1 = X_t^2 \left| X_t^1 \right|^{2/3} \mbox{d}W_t^1, \\
 & \mbox{d}X_t^2 = g(t)\mbox{d}X_t^3, \\
 & \mbox{d}X_t^3 = X_t^3( \mu  \mbox{d}t + \sigma (\rho \mbox{d}W_t^1 + \sqrt{1-\rho^2} 
   \mbox{d}W_t^2))
 \end{cases},
 \ \ \ \ (X_0^1, X_0^2, X_0^3) = (1.0,0.1,1.0)
 \end{equation}
 with $\mu = 0.1, \sigma = 0.2, \rho = -0.7 $ and $g(t) = 0.4 + (0.1 + 0.2t) e^{-2t}$,  for
 example, where $W = (W^1, W^2)$ is a 2-dim standard Brownian motion.
 The solution  $X = (X^1, X^2, X^3)$ takes values on $\mathbb{R}_+^3$.
 This is a stochastic integral equation defined as
 \begin{equation}
 X_t = X_0 +  \int_0^t b(s,X_s) \mbox{d}s + \int_0^t c(s,X_s) \mbox{d}W_s
 \end{equation}
 with 
 \[
  b(s,x) = \left( \begin{array}{@{\,}c@{\,}} 0 \\ g(s) \mu x^3 \\ \mu x^3 \end{array} \right),
 \ \ 
 c(s,x) = \left( \begin{array}{@{\,}cc@{\,}} x^2 | x^1|^{2/3} &  0 \\ 
  g(s)\sigma \rho x^3 &  g(s)\sigma \sqrt{1-\rho^2} x^3 \\ 
 \sigma \rho  x^3 & \sigma \sqrt{1-\rho^2} x^3 
 \end{array} \right)
 \]
 for $ x = (x^1, x^2, x^3)$. 
 The Euler-Maruyama method with discretization scheme
 \begin{equation}
 0 = \tau_0 < \tau_1 < \cdots < \tau_j < \tau_{j+1} < \cdots
 \end{equation}
 refers to an approximative solution given by
 \begin{equation}
 \tilde{X}_{\tau_{j+1}} = \tilde{X}_{\tau_j} + 
 b(\tau_j,\tilde{X}_{\tau_j})(\tau_{j+1} - \tau_j) + 
 c(\tau_j, \tilde{X}_{\tau_j} )(W_{\tau_{j+1}}- W_{\tau_j}), \ \ j \geq 0.
 \end{equation}
 If interval $\tau_{j+1} - \tau_j$ is independent to  $\{W_t\}_{t \geq \tau_j}$ for each $j$,
 $W_{\tau_{j+1}} - W_{\tau_j}$ has the same distribution as $\sqrt{\tau_{j+1} - \tau_j}N_j$,
 where $N_j$ is an iid sequence of standard normal variables.
 The usual discretization scheme is an equidistant one such as 
 \begin{equation}
 \tau_{j+1} - \tau_j = \text{Terminal}/\text{division}
 \end{equation}
 for all $j \geq 0$.

<<fig=TRUE,echo=TRUE,width=7,height=7>>=
 set.seed(123)

 mu <- 0.1
 sig <- 0.2 
 rho <- -0.7

 g <- function(t){
    return(0.4 + (0.1 + 0.2*t)* exp(-2*t))
 }

 f1 <- function(t, x1, x2, x3) {
    ret <- 0
    if(x1 > 0 && x2 > 0) ret <- x2*exp(log(x1)*2/3)
    return(ret)
 }

 f2 <- function(t, x1, x2, x3) {
     ret <- 0
     if(x3 > 0) ret <- rho*sig*x3
     return(ret)
 }

 f3 <- function(t, x1, x2, x3) {
     ret <- 0
     if(x3 > 0) ret <- sqrt(1-rho^2)*sig*x3
     return(ret)
 }

 diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3) * g(t)", 
     "f2(t,x1,x2,x3)", "0", "f3(t,x1,x2,x3)*g(t)", "f3(t,x1,x2,x3)"),  3, 2)

 sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, 
                       state.variable = c("x1", "x2", "x3"), solve.variable = c("x1", "x2", "x3"))

 # Change the parameters below if needed
 yuima.samp <- setSampling(Terminal=1,division=1000)

 # make yuima object
 yuima <- setYuima(model=sabr.mod, sampling=yuima.samp)

 # Solve SDEs using Euler-Maruyama method
 yuima <- simulate(yuima, xinit=c(1.0,0.1,1.0))
 plot(yuima)
@

 \subsection{Space-discretizing scheme}

 \code{simulate} provides also space-discretized Euler-Maruyama method to solve SDE.
 This is at the moment designed for 1 factor SDE, i.e., the case the driving
 Brownian motion $W$ is 1-dimensional. Here discretization scheme $\{\tau_j\}$ 
 is defined as

\begin{equation}
\tau_0 = 0, \ \ \tau_{j+1} = \inf\{ t > \tau_j; |W_t - W_{\tau_j}| = \epsilon \}
\end{equation}
for each $j \geq 0$. \code{simulate} takes
\begin{equation}
\epsilon^2 = \text{Terminal} / \text{division},
\end{equation}
which coinsides with the mean of the interval  $\tau_{j+1} - \tau_j$. 
This space-discretizing scheme is 3 times efficient than the usual time-discretizing
one in the sense of mean-squared error.

Let us consider the previous model (\ref{sabr}) with $\rho = -1$.

<<fig=TRUE,echo=TRUE,width=7,height=7>>=

set.seed(123)

rho <- -1

diff.coef.matrix <- c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3) * g(t)", 
    "f2(t,x1,x2,x3)")

sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, 
                      state.variable = c("x1", "x2", "x3"), solve.variable = c("x1", "x2", "x3"))

yuima <- setYuima(model=sabr.mod, sampling=yuima.samp)

#yuima <- simulate(yuima)
yuima.sd <- simulate(yuima, space.discretized = TRUE, xinit=c(1.0,0.1,1.0))

#plot(yuima)
plot(yuima.sd)
@

### 2009-11-07 pkg/yuimadocs/inst/doc/yuima_4.Rnw
package skeleton for yuimadoc package, not yet working
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@11 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{cce}
  cce is a function to estimate the covariance between two It\^o 
  processes when they are observed at discrete times nonsynchronously. 
  It can also apply to irregularly sampled one-dimensional data as a special case.

\subsection{Nonsynchronous covariance estimator}
Suppose that two It\^o processes are observed
only at discrete times in a nonsynchronous manner.
We are interested in estimating the covariance of the two processes accurately
in such a situation.
This type of problem arises typically in high-frequency financial time series.

Let $T \in (0,\infty)$ be a terminal time for possible observations. 
We consider a two dimensional It\^o process $(X^1,X^2)$  
satisfying the stochastic differential equations
\begin{eqnarray*}
\mbox{d}X_t^l &=&\mu_t^l \mbox{d}t + \sigma_t^l \mbox{d}W_t^l ,\quad
t\in[0,T]\\
X_0^l &=& x_0^l        \notag
\end{eqnarray*}
for $l=1,2$. 
Here $W^l$ denote standard Wiener processes with a 
progressively measurable correlation process 
$\mbox{d}\langle W_1 , W_2 \rangle_t = \rho_t \mbox{dt}$,  
$\mu^l_t$ and $\sigma_t^l$ are progressively measurable processes, 
and $x^l_0$ are initial random variables independent of $(W^1,W^2)$.  
Diffusion type processes are in the scope but this model 
can express more sophisticated stochastic structures.

The process $X^l$ is supposed to be observed at 
over the increasing sequence of times 
$T^{l,i}$ $(i\in\mathbb{Z}_{\geq0})$ starting at $0$, up to time T. 
Thus, the observables are $(T^{l,i},X^{l,i})$ with $T^{l,i}\leq T$. 
Each $T^{l,j}$ may be a stopping time, so possibly depends on 
the history of $(X^1,X^2)$ as well as the precedent stopping times.  
Two sequences of stopping times 
$T^{1,i}$ and $T^{2,j}$ are {\it nonsynchronous},  and 
irregularly spaced, in general. 
In particular, cce can apply to estimation of the quadratic variation of a single 
stochastic process sampled regularly/irregularly.

The parameter of interest is the quadratic covariation between $X^1$ and $X^2$:

\begin{equation}
\theta=
\langle X^1 , X^2 \rangle_T = \int_0^T \sigma_t^1 \sigma_t^2 \rho_t \mbox{d}t.
\end{equation}
The target variable $\theta$ is random in general.

It can be estimated with the nonsynchronous covariance estimator 
(Hayashi-Yoshida estimator) 
\begin{equation}
U_n= \sum_{i,j:T^{1,i}\leq T, T^{2,j}\leq T} (X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)
1_{\{ (T^{1,i-1},T^{1,i}] 
\bigcap (T^{2,j-1},T^{2,j}]  \neq \emptyset  \}}.
\end{equation}
That is, the product of any pair of increments $(X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)$ and
$(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)$ will make a contribution 
to the sum only when 
the respective observation intervals $(T^{1,i-1},T^{1,i}] $ and $ (T^{2,j-1},T^{2,j}] $ are overlapping 
with each other. 
%The estimator $U_n$ was proposed by Hayashi and Yoshida  and 
It is known that $U_n$ is consist and has asymptotically mixed normal 
distribution 
as $n\to\infty$ if the maximum length between two consecutive observing times 
tends to $0$. 
See \cite{hay-yos05,hay-yos04,hay-yos06,hay-yos08} for details.

\subsection{Example: data generation and estimation by yuima package}
We will demonstrate how to apply cce function to 
nonsynchronous high-frequency data by simulation. 
As an example, consider a two dimensional stochastic process 
$(X^1_t,X^2_t)$ satisfying the stochastic differential equation
\begin{equation}
\begin{split}
\mbox{d}X^1_t = \sigma_{1,t} \mbox{d}B^1_t, \\
\mbox{d}X^2_t = \sigma_{2,t} \mbox{d}B^2_t. 
\end{split}
\end{equation}
Here $B^1_t$ and $B^2_t$ denote two standard Wiener processes, 
however they are correlated as 
\begin{eqnarray}
B^1_t &=& W^1_t,\\
B^2_t &=&  \int_0^t \rho_s \mbox{d} W^1_s +
  \int_0^t \sqrt{ 1- \rho_s^2} \mbox{d} W^2_s,
\end{eqnarray}
where  $W^1_t$ and $W^2_t$ are independent Wiener processes,  and 
$\rho_t$ is the correlation function between $B^1_t$ and $B^2_t$. 
We consider $\sigma_{l,t},l=1,2$ and $\rho_t$ 
of the following form in this example:

\begin{eqnarray*}
 \sigma_{1,t} &=& \sqrt{1+t}, \\
 \sigma_{2,t} &=& \sqrt{1+t^2}, \\
 \rho_t &=& \frac{1}{\sqrt{2}}.
\end{eqnarray*}

To simulate the stochastic process $(X^1_t,X^2_t)$, we first build the model 
by setModel as follows. 
It should be noted that the method of generating 
nonsynchronous data can be replaced by a simpler one but 
we will take a general approach here to demonstrate 
a usage of 
the yuima comprehensive package for simulation and estimation of 
stochastic processes.

<<print=FALSE,echo=TRUE>>=
# diffusion coefficient for process 1
diff.coef.1 <- function(t,x1=0, x2=0) sqrt(1+t)
# diffusion coefficient for process 2
diff.coef.2 <- function(t,x1=0, x2=0) sqrt(1+t^2)
# correlation
cor.rho <- function(t,x1=0, x2=0) sqrt(1/2)
# coefficient matrix for diffusion term
diff.coef.matrix <- matrix( c( "diff.coef.1(t,x1,x2)", "diff.coef.2(t,x1,x2) * cor.rho(t,x1,x2)", "", "diff.coef.2(t,x1,x2) * sqrt(1-cor.rho(t,x1,x2)^2)"),2,2)
# Model sde using yuima.model
cor.mod <- setModel(drift = c("",""), diffusion = diff.coef.matrix, solve.variable=c("x1","x2"))
@

The parameter we want to estimate is the quadratic covariation between $X_1$ and $X_2$:

\begin{equation}
\theta = \langle X_1, X_2 \rangle_T =
 \int_0^T \sigma_{1,t} \sigma_{2,t} \rho_t \mbox{d} t.
\end{equation}
Later, we will compare estimated values with 
the true value of  $\theta$ given by 
<<echo=TRUE>>=
CC.theta <- function( T, sigma1, sigma2, rho)
{
 	tmp <- function(t) return( sigma1(t) * sigma2(t) * rho(t) )
 	integrate(tmp,0,T)
}
# Cumulative Covariance
theta <- CC.theta(T=1, sigma1=diff.coef.1, sigma2=diff.coef.2, rho=cor.rho)$value
cat(sprintf("theta=%5.3f\n",theta))
@ 
For the sampling scheme, we will consider the independent oisson sampling. 
That is, each configuration of the sampling times $T^{l,i}$ is realized 
as the Poisson random measure with intensity $np_l$, 
and the two random measures are independent each other as well as 
the stochastic processes. 
Then it is known that

\begin{equation}
n^{1/2} ( U_n -\theta) \rightarrow N(0,c),
\end{equation}
as $n\to\infty$, 
where 
\begin{equation}
 c = \left( \frac{2}{p_1} + \frac{2}{p_2} \right)
\int_0^T \left( \sigma_{1,t} \sigma_{2,t}  \right)^2 \mbox{d}t +
\left(
\frac{2}{p_1} + \frac{2}{p_2} - \frac{2}{p_1 + p_2}
\right)
\int_0^T
\left(
\sigma_{1,t} \sigma_{2,t} \rho_t
\right)^2 \mbox{d} t.
\end{equation}

We use a function \code{poisson.random.sampling} to get observations 
by the Poisson sampling.
This function returns a matrix combined times and observations of
Poisson random sampling.

<<echo=TRUE>>=
set.seed(111)

Terminal <- 1
p1 <- 0.2
p2 <- 0.3
rate<-c(p1, p2)
n <- 1000 
# Change to n<-1000
division <- 1200
# Change too division <- 1200

yuima.samp <- setSampling(Terminal=Terminal,division=division)
yuima <- setYuima(model=cor.mod, sampling=yuima.samp)

# solve SDEs using Euler-Maruyama method
Data <- simulate(yuima)

sampledData <- poisson.random.sampling(Data,rate=rate, n=n)
@ 
\code{cce} takes the sample and returns 
an estimate of the quadratic covariation.
<<cce, echo=TRUE>>=
cce(sampledData)[1,2] # CumulativeCovarianceEstimator
@ 
and in case of non missing data
<<cce2, echo=TRUE>>=
cce(Data)[1,2] # CumulativeCovarianceEstimator
@ 
The last value is just for reference; it may bear another kind error 
due to approximation error to the Poisson sampling.

In what follows, we will evaluate the difference between 
the estimated value and the true value, as well as the asymptotic variance, 
and compare the histogram of the estimation error after scaling 
with the limiting Gaussian distribution 
(Figure \ref{fig:histogram}). 
The function \texttt{var.c} calculates the theoretical 
asymptotic variance of the scaled estimation error.

<<cctheta2, echo=TRUE>>=
#error <- NULL   # difference between true parameter and estimator

# Cumulative Covariance
theta <- CC.theta(T=Terminal, diff.coef.1, diff.coef.2, cor.rho)$value
cat(sprintf("The true value of theta=%5.3f\n",theta))

# asymptotic variance
var.c <- function(T, p1,p2, sigma1, sigma2, rho)
{
  tmp_integrand1 <- function(t) (sigma1(t) * sigma2(t))^2
  i1 <- integrate(tmp_integrand1,0,T)
  tmp_integrand2 <- function(t) (sigma1(t) * sigma2(t) * rho(t))^2
  i2 <- integrate(tmp_integrand2,0,T)
  2*(1/p1 + 1/p2)* i1$value + 2*(1/p1+1/p2 - 1/(p1+p2)) * i2$value
}

vc <- var.c(T=Terminal, p1, p2, diff.coef.1, diff.coef.2, cor.rho)
sd <- sqrt(vc/n)
cat(sprintf("The standard deviation by normal approximation =%5.3f\n",sd))

# simulation
N <- 100
U <- NULL
for(i in 1:N)
{
  # solve SDEs using Euler-Maruyama method
  yuima.tmp <- simulate(yuima)
  yuimaData <- poisson.random.sampling(yuima.tmp,rate=rate,n=n)
  Un <- cce(yuimaData)[1,2] # CumulativeCovarianceEstimator
  U <- append(U, Un)
}

cat(sprintf("The sample mean =%5.3f\n", mean(U)))
cat(sprintf("The sample standard deviation =%5.3f\n", sqrt(var(U))))
@

\begin{figure}[h]
\centering
<<fig1, fig=TRUE,echo=TRUE,width=10,height=10,results=hide>>=
# plot sampled data histogram and Gaussian distribution (theoretical result).
library(KernSmooth)
h <- dpih(U)
bins <- seq(min(U)-5, max(U)+5+h,by=h)
xrange <- c(min(U)-5-h, max(U)+5+h)
yrange<-c(0,max(hist(U)$density))
hist(U,breaks=bins,xlab="",ylab="",xlim=xrange,ylim=yrange,probability=T)
curve(dnorm(x,mean=theta,sd=sd),ylab="",xlim=xrange,ylim=yrange,col="red", add=TRUE)
@  
\caption{A sampled data histogram and Gaussian distribution(theoretical result)}
\label{fig:histogram}
\end{figure}

\begin{thebibliography}{99}
\bibitem{hay-yos05}
Hayashi, T., Yoshida, N.: 
On Covariance Estimation of Non-synchronously Observed Diffusion Processes. 
Bernoulli, {\bf 11}, 2, 359-379 (2005)

\bibitem{hay-yos04}
Hayashi, T., Yoshida, N.: 
Asymptotic normality of a covariance estimator for nonsynchronously observed diffusion processes. 
Annals of the Institute of Statistical Mathematics, 
{\bf 60}, 2, %/ 2008”N6ŒŽ 
367-406 (2008)

\bibitem{hay-yos06}
Hayashi, T., Yoshida, N.: 
Nonsynchronous Covariance Estimator and Limit Theorem. 
Research Memorandum No.1020, Institute of Statistical Mathematics, 2006

\bibitem{hay-yos08}
Hayashi, T., Yoshida, N.: 
Nonsynchronous Covariance Estimator and Limit Theorem II. 
Research Memorandum No.1067, Institute of Statistical Mathematics, 2008

\end{thebibliography}

### 2009-11-10 pkg/yuimadocs/inst/doc/bib.Rnw
first version of yuima doc
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@14 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2009-11-07 pkg/yuimadocs/inst/doc/yuima_4.Rnw
@@
-\section{cce}
-  cce is a function to estimate the covariance between two It\^o 
-  processes when they are observed at discrete times nonsynchronously. 
-  It can also apply to irregularly sampled one-dimensional data as a special case.
-
-\subsection{Nonsynchronous covariance estimator}
-Suppose that two It\^o processes are observed
-only at discrete times in a nonsynchronous manner.
-We are interested in estimating the covariance of the two processes accurately
-in such a situation.
-This type of problem arises typically in high-frequency financial time series.
-
-Let $T \in (0,\infty)$ be a terminal time for possible observations. 
-We consider a two dimensional It\^o process $(X^1,X^2)$  
-satisfying the stochastic differential equations
-\begin{eqnarray*}
-\mbox{d}X_t^l &=&\mu_t^l \mbox{d}t + \sigma_t^l \mbox{d}W_t^l ,\quad
-t\in[0,T]\\
-X_0^l &=& x_0^l        \notag
-\end{eqnarray*}
-for $l=1,2$. 
-Here $W^l$ denote standard Wiener processes with a 
-progressively measurable correlation process 
-$\mbox{d}\langle W_1 , W_2 \rangle_t = \rho_t \mbox{dt}$,  
-$\mu^l_t$ and $\sigma_t^l$ are progressively measurable processes, 
-and $x^l_0$ are initial random variables independent of $(W^1,W^2)$.  
-Diffusion type processes are in the scope but this model 
-can express more sophisticated stochastic structures.
-
-The process $X^l$ is supposed to be observed at 
-over the increasing sequence of times 
-$T^{l,i}$ $(i\in\mathbb{Z}_{\geq0})$ starting at $0$, up to time T. 
-Thus, the observables are $(T^{l,i},X^{l,i})$ with $T^{l,i}\leq T$. 
-Each $T^{l,j}$ may be a stopping time, so possibly depends on 
-the history of $(X^1,X^2)$ as well as the precedent stopping times.  
-Two sequences of stopping times 
-$T^{1,i}$ and $T^{2,j}$ are {\it nonsynchronous},  and 
-irregularly spaced, in general. 
-In particular, cce can apply to estimation of the quadratic variation of a single 
-stochastic process sampled regularly/irregularly.
-
-The parameter of interest is the quadratic covariation between $X^1$ and $X^2$:
-
-\begin{equation}
-\theta=
-\langle X^1 , X^2 \rangle_T = \int_0^T \sigma_t^1 \sigma_t^2 \rho_t \mbox{d}t.
-\end{equation}
-The target variable $\theta$ is random in general.
-
-It can be estimated with the nonsynchronous covariance estimator 
-(Hayashi-Yoshida estimator) 
-\begin{equation}
-U_n= \sum_{i,j:T^{1,i}\leq T, T^{2,j}\leq T} (X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)
-1_{\{ (T^{1,i-1},T^{1,i}] 
-\bigcap (T^{2,j-1},T^{2,j}]  \neq \emptyset  \}}.
-\end{equation}
-That is, the product of any pair of increments $(X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)$ and
-$(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)$ will make a contribution 
-to the sum only when 
-the respective observation intervals $(T^{1,i-1},T^{1,i}] $ and $ (T^{2,j-1},T^{2,j}] $ are overlapping 
-with each other. 
-%The estimator $U_n$ was proposed by Hayashi and Yoshida  and 
-It is known that $U_n$ is consist and has asymptotically mixed normal 
-distribution 
-as $n\to\infty$ if the maximum length between two consecutive observing times 
-tends to $0$. 
-See \cite{hay-yos05,hay-yos04,hay-yos06,hay-yos08} for details.
-
-\subsection{Example: data generation and estimation by yuima package}
-We will demonstrate how to apply cce function to 
-nonsynchronous high-frequency data by simulation. 
-As an example, consider a two dimensional stochastic process 
-$(X^1_t,X^2_t)$ satisfying the stochastic differential equation
-\begin{equation}
-\begin{split}
-\mbox{d}X^1_t = \sigma_{1,t} \mbox{d}B^1_t, \\
-\mbox{d}X^2_t = \sigma_{2,t} \mbox{d}B^2_t. 
-\end{split}
-\end{equation}
-Here $B^1_t$ and $B^2_t$ denote two standard Wiener processes, 
-however they are correlated as 
-\begin{eqnarray}
-B^1_t &=& W^1_t,\\
-B^2_t &=&  \int_0^t \rho_s \mbox{d} W^1_s +
-  \int_0^t \sqrt{ 1- \rho_s^2} \mbox{d} W^2_s,
-\end{eqnarray}
-where  $W^1_t$ and $W^2_t$ are independent Wiener processes,  and 
-$\rho_t$ is the correlation function between $B^1_t$ and $B^2_t$. 
-We consider $\sigma_{l,t},l=1,2$ and $\rho_t$ 
-of the following form in this example:
-
-\begin{eqnarray*}
- \sigma_{1,t} &=& \sqrt{1+t}, \\
- \sigma_{2,t} &=& \sqrt{1+t^2}, \\
- \rho_t &=& \frac{1}{\sqrt{2}}.
-\end{eqnarray*}
-
-To simulate the stochastic process $(X^1_t,X^2_t)$, we first build the model 
-by setModel as follows. 
-It should be noted that the method of generating 
-nonsynchronous data can be replaced by a simpler one but 
-we will take a general approach here to demonstrate 
-a usage of 
-the yuima comprehensive package for simulation and estimation of 
-stochastic processes.
-
-<<print=FALSE,echo=TRUE>>=
-# diffusion coefficient for process 1
-diff.coef.1 <- function(t,x1=0, x2=0) sqrt(1+t)
-# diffusion coefficient for process 2
-diff.coef.2 <- function(t,x1=0, x2=0) sqrt(1+t^2)
-# correlation
-cor.rho <- function(t,x1=0, x2=0) sqrt(1/2)
-# coefficient matrix for diffusion term
-diff.coef.matrix <- matrix( c( "diff.coef.1(t,x1,x2)", "diff.coef.2(t,x1,x2) * cor.rho(t,x1,x2)", "", "diff.coef.2(t,x1,x2) * sqrt(1-cor.rho(t,x1,x2)^2)"),2,2)
-# Model sde using yuima.model
-cor.mod <- setModel(drift = c("",""), diffusion = diff.coef.matrix, solve.variable=c("x1","x2"))
-@
-
-The parameter we want to estimate is the quadratic covariation between $X_1$ and $X_2$:
-
-\begin{equation}
-\theta = \langle X_1, X_2 \rangle_T =
- \int_0^T \sigma_{1,t} \sigma_{2,t} \rho_t \mbox{d} t.
-\end{equation}
-Later, we will compare estimated values with 
-the true value of  $\theta$ given by 
-<<echo=TRUE>>=
-CC.theta <- function( T, sigma1, sigma2, rho)
-{
- 	tmp <- function(t) return( sigma1(t) * sigma2(t) * rho(t) )
- 	integrate(tmp,0,T)
-}
-# Cumulative Covariance
-theta <- CC.theta(T=1, sigma1=diff.coef.1, sigma2=diff.coef.2, rho=cor.rho)$value
-cat(sprintf("theta=%5.3f\n",theta))
-@ 
-For the sampling scheme, we will consider the independent oisson sampling. 
-That is, each configuration of the sampling times $T^{l,i}$ is realized 
-as the Poisson random measure with intensity $np_l$, 
-and the two random measures are independent each other as well as 
-the stochastic processes. 
-Then it is known that
-
-\begin{equation}
-n^{1/2} ( U_n -\theta) \rightarrow N(0,c),
-\end{equation}
-as $n\to\infty$, 
-where 
-\begin{equation}
- c = \left( \frac{2}{p_1} + \frac{2}{p_2} \right)
-\int_0^T \left( \sigma_{1,t} \sigma_{2,t}  \right)^2 \mbox{d}t +
-\left(
-\frac{2}{p_1} + \frac{2}{p_2} - \frac{2}{p_1 + p_2}
-\right)
-\int_0^T
-\left(
-\sigma_{1,t} \sigma_{2,t} \rho_t
-\right)^2 \mbox{d} t.
-\end{equation}
-
-We use a function \code{poisson.random.sampling} to get observations 
-by the Poisson sampling.
-This function returns a matrix combined times and observations of
-Poisson random sampling.
-
-<<echo=TRUE>>=
-set.seed(111)
-
-Terminal <- 1
-p1 <- 0.2
-p2 <- 0.3
-rate<-c(p1, p2)
-n <- 1000 
-# Change to n<-1000
-division <- 1200
-# Change too division <- 1200
-
-yuima.samp <- setSampling(Terminal=Terminal,division=division)
-yuima <- setYuima(model=cor.mod, sampling=yuima.samp)
-
-# solve SDEs using Euler-Maruyama method
-Data <- simulate(yuima)
-
-sampledData <- poisson.random.sampling(Data,rate=rate, n=n)
-@ 
-\code{cce} takes the sample and returns 
-an estimate of the quadratic covariation.
-<<cce, echo=TRUE>>=
-cce(sampledData)[1,2] # CumulativeCovarianceEstimator
-@ 
-and in case of non missing data
-<<cce2, echo=TRUE>>=
-cce(Data)[1,2] # CumulativeCovarianceEstimator
-@ 
-The last value is just for reference; it may bear another kind error 
-due to approximation error to the Poisson sampling.
-
-In what follows, we will evaluate the difference between 
-the estimated value and the true value, as well as the asymptotic variance, 
-and compare the histogram of the estimation error after scaling 
-with the limiting Gaussian distribution 
-(Figure \ref{fig:histogram}). 
-The function \texttt{var.c} calculates the theoretical 
-asymptotic variance of the scaled estimation error.
-
-<<cctheta2, echo=TRUE>>=
-#error <- NULL   # difference between true parameter and estimator
-
-# Cumulative Covariance
-theta <- CC.theta(T=Terminal, diff.coef.1, diff.coef.2, cor.rho)$value
-cat(sprintf("The true value of theta=%5.3f\n",theta))
-
-# asymptotic variance
-var.c <- function(T, p1,p2, sigma1, sigma2, rho)
-{
-  tmp_integrand1 <- function(t) (sigma1(t) * sigma2(t))^2
-  i1 <- integrate(tmp_integrand1,0,T)
-  tmp_integrand2 <- function(t) (sigma1(t) * sigma2(t) * rho(t))^2
-  i2 <- integrate(tmp_integrand2,0,T)
-  2*(1/p1 + 1/p2)* i1$value + 2*(1/p1+1/p2 - 1/(p1+p2)) * i2$value
-}
-
-vc <- var.c(T=Terminal, p1, p2, diff.coef.1, diff.coef.2, cor.rho)
-sd <- sqrt(vc/n)
-cat(sprintf("The standard deviation by normal approximation =%5.3f\n",sd))
-
-# simulation
-N <- 100
-U <- NULL
-for(i in 1:N)
-{
-  # solve SDEs using Euler-Maruyama method
-  yuima.tmp <- simulate(yuima)
-  yuimaData <- poisson.random.sampling(yuima.tmp,rate=rate,n=n)
-  Un <- cce(yuimaData)[1,2] # CumulativeCovarianceEstimator
-  U <- append(U, Un)
-}
-
-cat(sprintf("The sample mean =%5.3f\n", mean(U)))
-cat(sprintf("The sample standard deviation =%5.3f\n", sqrt(var(U))))
-@
-
-\begin{figure}[h]
-\centering
-<<fig1, fig=TRUE,echo=TRUE,width=10,height=10,results=hide>>=
-# plot sampled data histogram and Gaussian distribution (theoretical result).
-library(KernSmooth)
-h <- dpih(U)
-bins <- seq(min(U)-5, max(U)+5+h,by=h)
-xrange <- c(min(U)-5-h, max(U)+5+h)
-yrange<-c(0,max(hist(U)$density))
-hist(U,breaks=bins,xlab="",ylab="",xlim=xrange,ylim=yrange,probability=T)
-curve(dnorm(x,mean=theta,sd=sd),ylab="",xlim=xrange,ylim=yrange,col="red", add=TRUE)
-@  
-\caption{A sampled data histogram and Gaussian distribution(theoretical result)}
-\label{fig:histogram}
-\end{figure}
-
 \begin{thebibliography}{99}
@@
 Annals of the Institute of Statistical Mathematics, 
-{\bf 60}, 2, %/ 2008”N6ŒŽ 
+{\bf 60}, 2, %/ 2008Ç¯6·î 
 367-406 (2008)
@@
 
+\bibitem[Yoshida, 2005a]{Yos05a}
+Yoshida, N. (2005). 
+\newblock 
+General M-estimation for stochastic differential equation with 
+jumps by sampled data. 
+\newblock 
+in preparation
+
 \end{thebibliography}

### 2009-11-10 pkg/yuimadocs/R/zzz.R

@@ -1,0 +1,8 @@
+.noGenerics <- TRUE
+
+.onLoad <- function(libname, pkgname)
+{
+  cat("\nHow to use yuima? Type vignette(\"yuima\")\n")
+# library.dynam("cem", pkgname, libname) 
+}
+

### 2009-11-11 pkg/yuimadocs/inst/doc/yuima_5.Rnw
section 5
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@15 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{adaBayes}
  adaBayes is a function that computes Bayesian estimators for unknown parameters of 
  a stochastic differential equation based on discretely observed data.

\subsection{Bayesian type estimators for the stochastic differential equation}\label{190105-4}

Consider a diffusion process $X=(X_t)_{t\in\bbR_+}$ satisfying the 
stochastic differential equation 
\beas 
dX_t=a(X_t,\theta_2)dt+b(X_t,\theta_1)dw_t,
\sskip X_0=x_0,
\eeas
where $w_t$ is an{$r$}-dimensional standard Wiener process 
independent of the initial value $x_0$. 
We suppose that the parameters 
$\theta_1$ and $\theta_2$ are unknown 
but 
$\theta_i\in\Theta_i\subset\bbR^{m_i}$ 
for $i=1,2$. 
Also assume that 
$B(x,\theta_1)=bb'(x,\theta_1)$ is elliptic uniformly 
in $(x,\theta_1)$.

In order to estimate the unknown parameters 
with the discrete-time 
observations ${\bf x}_n=(X_{t_i})_{i=0}^n$, 
$t_i=ih$ with $h=h_n$ depending on $n\in\bbN$,  
we use the quasi-likelihood function 
\beas 
p_n({\bf x}_n,\theta)
&=&\prod_{i=1}^n\frac{1}{(2\pi h)^{d/2}|B(X_{t_{i-1}},\theta_1)|^{1/2}}
\\&&
\cdot 
\exp\left(-\frac{1}{2h}
B(X_{t_{i-1}},\theta_1)^{-1}\left[
(\Delta_iX-ha(X_{t_{i-1}},\theta_2) )^{\otimes2}\right]\right),
\eeas
where $\Delta_iX=X_{t_i}-X_{t_{i-1}}$. 
%

Two approaches are possible. 
One is the quasi-maximum likelihood approach  
and another is the adaptive Bayesian approach.

The quasi-maximum likelihood estimator 
that maximizes $p_n({\bf x}_n,\theta)$ 
in $\theta=(\theta_1,\theta_2)\in
=\overline{\Theta_1\times\Theta_2}$ 
is denoted by $\hat{\theta}_{n}=
(\hat{\theta}_{1,n},\hat{\theta}_{2,n})$.

The adaptive Bayes type estimator is defined as follows. 
First we choose arbitrary value $\theta_2^\star\in\Theta_2$ and 
pretend $\theta_1$ is the unknown parameter to 
make the Bayesian type estimator $\tilde{\theta}_1$ as 
\beas 
\tilde{\theta}_1
&=&
\Big[\int_{\Theta_1}p_n({\bf x}_n,(\theta_1,\theta_2^\star))
\pi_1(\theta_1)d\theta_1 \Big]^{-1}
\int_{\Theta_1} \theta_1 p_n({\bf x}_n,(\theta_1,\theta_2^\star))
\pi_1(\theta_1)d\theta_1, 
\eeas
where 
$\pi_1$ is a prior density on $\Theta_1$. 
According to the asymptotic theory, 
if $\pi_1$ is positive on $\Theta_1$, any function can be used. 
For estimation of $\theta_2$, we use $\tilde{\theta}_1$ 
to reform the quasi-likelihood function. That is,
the Bayes type estimator for $\theta_2$ is defined by 
\beas 
\tilde{\theta}_2
&=&
\Big[\int_{\Theta_2}p_n({\bf x}_n,(\tilde{\theta}_1,\theta_2))
\pi_2(\theta_2)d\theta_2 \Big]^{-1}
\int_{\Theta_2} \theta_2 p_n({\bf x}_n,(\tilde{\theta}_1,\theta_2))
\pi_2(\theta_2)d\theta_2, 
\eeas
where 
$\pi_2$ is a prior density on $\Theta_2$. 
In this way, we obtain the adaptive Bayes type estimator 
$\tilde{\theta}=(\tilde{\theta}_1,\tilde{\theta}_2)$ 
for $\theta=(\theta_1,\theta_2)$.

The asymptotic behavior of those estimators is known 
in various situations. The results depend on what kind 
asymptotics one considers.

See \cite{yos05}.

\subsection{Example}

### 2010-01-29 pkg/yuimadocs/inst/doc/yuima_4.Rnw
minor fix
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@62 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 # Change to n<-1000
-division <- 1200
-# Change too division <- 1200
+n <- 1200
+# Change too n <- 1200
 
-yuima.samp <- setSampling(Terminal=Terminal,division=division)
+yuima.samp <- setSampling(Terminal=Terminal,n=n)
 yuima <- setYuima(model=cor.mod, sampling=yuima.samp)
@@
 \end{figure}
-
-\begin{thebibliography}{99}
-\bibitem{hay-yos05}
-Hayashi, T., Yoshida, N.: 
-On Covariance Estimation of Non-synchronously Observed Diffusion Processes. 
-Bernoulli, {\bf 11}, 2, 359-379 (2005)
-
-\bibitem{hay-yos04}
-Hayashi, T., Yoshida, N.: 
-Asymptotic normality of a covariance estimator for nonsynchronously observed diffusion processes. 
-Annals of the Institute of Statistical Mathematics, 
-{\bf 60}, 2, %/ 2008”N6ŒŽ 
-367-406 (2008)
-
-\bibitem{hay-yos06}
-Hayashi, T., Yoshida, N.: 
-Nonsynchronous Covariance Estimator and Limit Theorem. 
-Research Memorandum No.1020, Institute of Statistical Mathematics, 2006
-
-\bibitem{hay-yos08}
-Hayashi, T., Yoshida, N.: 
-Nonsynchronous Covariance Estimator and Limit Theorem II. 
-Research Memorandum No.1067, Institute of Statistical Mathematics, 2008
-
-\end{thebibliography}

### 2010-03-07 pkg/yuimadocs/inst/doc/yuima_5.Rnw
minor addition on documentation
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@64 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\subsection{Technical Details}
+2010/03/07: Hideitsu Hino wrote\\
+[TBC: Description about ml.ql function. It's possible options. Optim
+function equipped with R which uses ``Nelder-Mead method'' for
+likelihood maximization. By specifying method=''Newton'', we can use
+Newton method for likelihood maximization.]
+
+[TBC: Description about adaBayes function. It's possible options. MCMC.]
+
+[TBC: Return value format of ml.ql and adaBayes. It is in the same form
+to the return value of ``mle'' function. That is, we can apply
+``confint'' to get their confidence intervals.]
+
 \subsection{Example}

### 2011-09-09 pkg/yuimadocs/inst/doc/JSS/article.Rnw
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@180 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Alexandre Brouste\\University of Le Mans \And 
        Stefano M. Iacus\\University of Milan \And 	
        Hiroki Masuda\\University of Kyushu \And
        Masaaki Fukasawa\\University of ???\AND
        Kengo Kamatani\\University of Tokyo \And
        Yusutaka Shimuzu\\University of Osaka \And
        Masayuki Uchida\\University of Osaka \And
        Nakahiro Yoshida\\University of Tokyo 
        }
\title{The Yuima Project: a Computational Framework for Simulation and Inference of Stochastic Differential Equations}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Alexandre Brouste, Stefano M. Iacus, Hiroki Masuda, Masayuki Uchida, Nakahiro Yoshida} %% comma-separated
\Plaintitle{A Capitalized Title: Something About a Package foo} %% without formatting
\Shorttitle{A Capitalized Title} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  The abstract of the article.
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{13}
%% \Issue{9}
%% \Month{September}
%% \Year{2004}
%% \Submitdate{2004-09-29}
%% \Acceptdate{2004-09-29}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Stefano M. Iacus\\
  Department of Economics, Management and Quantitative Methods\\
  University of Milan\\
  Via Conservatorio 7, 20122 Milan, Italy\\
  E-mail: \email{stefano.iacus@unimi.it}\\
  URL: \url{http://www.economia.unimi.it/~iacus/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/1/31336-5053
%% Fax: +43/1/31336-734

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\end{document}

### 2011-09-09 pkg/yuimadocs/inst/doc/JSS/article.Rnw
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@181 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\usepackage{Sweave}    
+%\usepackage{Rd}
+
+\usepackage{amsmath,amssymb}
 
@@
         }
-\title{The Yuima Project: a Computational Framework for Simulation and Inference of Stochastic Differential Equations}
+\title{The YUIMA Project: a Computational Framework for Simulation and Inference of Stochastic Differential Equations}
 
@@
 \Abstract{
-  The abstract of the article.
+The Yuima Project is an open source and collaborative effort  aimed at developing the \proglang{R} package named \pkg{yuima} for simulation and inference of stochastic differential equations.
+
+In the \pkg{yuima} package stochastic differential equations can be of very abstract type, multidimensional, driven by Wiener process of fractional Brownian motion with general Hurst parameter, with or without jumps specified as L\'evy noise.
+
+The \pkg{yuima} package is intended to offer the basic infrastructure on which complex models and inference procedures can be built on.
+This paper explains the design of the \pkg{yuima} package and provides some examples of applications.
 }
-\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
-\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
+\Keywords{inference for stochastic processes, simulation, stochastic differential equations}
+\Plainkeywords{inference for stochastic processes, simulation, stochastic differential equations} %% without formatting
 %% at least one keyword must be supplied
@@
   URL: \url{http://www.economia.unimi.it/~iacus/}
+\\
+\\
+  Nakahiro Yoshida\\
+  Graduate School of Mathematical Science\\
+  University of Tokyo\\
+  3-8-1 Komaba, Meguro-ku,  Tokyo 153-8914, Japan\\
+  E-mail: \email{nakahiro@ms.u-tokyo.ac.jp}\\
+  URL: \url{http://www.ms.u-tokyo.ac.jp/~nakahiro/hp-naka-e}
 }
@@
 
+\SweaveOpts{prefix.string=yuima} 
+\SweaveOpts{echo=TRUE}
+
+%% preamble for Rnw files
+<<print=FALSE, echo=FALSE>>=
+options(prompt="R> ")
+options(width=70)
+@
+
+\def\ve{{\varepsilon}}
+\def\de{{\rm d}}
+\def\dE{{\mathbb E}}
+\def\diag{\mathop{\rm diag}\nolimits}
+
 %% for those who use Sweave please include the following line (with % symbols):
 %% need no \usepackage{Sweave.sty}
-
 %% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section[About Java]{About \proglang{Java}}
+%\section[About Java]{About \proglang{Java}}
 %% Note: If there is markup in \(sub)section, then it has to be escape as above.
 
+\section{Introduction}
+The YUIMA Project\footnote{The Project has been  funded up to 2010 by the Japan Science Technology (JST) Basic Research Programs PRESTO, Grants-in-Aid for Scientific Research No.
+19340021.} is an open source\footnote{All code in the \pkg{yuima} package is subject to the GNU General Public License, Version 2, see \url{http://www.gnu.org/licenses/gpl-2.0.html}.}
+ academic project aimed at developing the \proglang{R} package named ``\pkg{yuima}'' for simulation and inference of stochastic differential equations. 
+The YUIMA Project is mainly developed by mathematicians and 
+statisticians who actively publish in the field of inference and simulation for stochastic
+differential equations.
+%The YUIMA Project Core Team, currently 
+%consists of the following people: A. Brouste, M. Fukasawa, H. Hino, S.M. Iacus, K. Kamatani, H.Masuda, Y. Shimizu, M. Uchida, N. Yoshida.
+
+ The \pkg{yuima} package  provides 
+an object-oriented programming environment 
+for simulation and statistical inference 
+for stochastic processes by \proglang{R}.
+The \pkg{yuima} package adopts the  $\tt S4$ system of classes and methods \citep{chambers98}.
+
+Under this framework, 
+the \pkg{yuima} package also supplies various functions 
+to execute simulation and statistical analysis. 
+Both categories of procedures may depend each other. 
+Statistical inference often requires a simulation technique 
+as a subroutine, and a certain simulation method 
+needs to fix a tuning parameter by applying 
+a statistical methodology. 
+It is especially the case of stochastic processes 
+because most of expected values involved 
+do not admit an explicit expression. 
+The \pkg{yuima} package facilitates comprehensive, systematic approaches 
+to the solution.
+
+Stochastic differential equations are 
+commonly used 
+to model random evolution along continuous or 
+practically continuous time, such as 
+the random movements of a stock price. 
+Theory of statistical inference for 
+stochastic differential equations already has a fairly long history, 
+more than three decades, but it is still developing quickly new 
+methodologies and expanding the area. 
+The formulas produced by the theory are usually very sophisticated, 
+which makes it difficult for practitioners not necessarily 
+familiar with this field to enjoy their utility. 
+For example, the asymptotic expansion method for computing 
+asian option prices (i.e., expectation of a functional of 
+a stochastic process) provides precise approximation values 
+instantaneously, taking advantage of the analytic approach, 
+but the formula, based on Malliavin calculus, has a long expression like more than one page!
+
+The \pkg{yuima} package delivers up-to-date methods as a package 
+onto the desk of the user working 
+with simulation and/or statistics for stochastic differential equations.
+
+Sampled data from a continuous-time process features 
+the time stamps as well as the positions of the object. 
+It is requiring a new theory of estimation. 
+The \pkg{yuima} framework can apply multi-dimensional time stamps 
+of tick data and provides diverse functions handling such kind data 
+to support statistical analysis.
+
+Although we assume that the reader of this paper has a basic knowledge of the \proglang{R} language, most of the examples are easy to be understood by anyone.
+
+\section{The \pkg{yuima} package}
+The package \pkg{yuima} depends on some other packages, like  \pkg{zoo}, which can be installed separately.
+The package \pkg{zoo} is used internally to store time series data. This dependence may change in the future adopting a more flexible class for internal storage of time series.
+
+\subsection{How to obtain the package}
+The \pkg{yuima} package is hosted on \textsf{R-Forge} and the web page
+of the Project is  \url{http://r-forge.r-project.org/projects/yuima}.
+The  \textsf{R-Forge} page contains the latest development version, and stable
+version of the package will also be available through \textsf{CRAN}.
+Development versions of the package are not supposed to be stable or functional, thus
+the occasional user should consider to install the stable version first.
+The package can be installed from \textsf{R-Forge} using
+\code{install.packages("yuima",repos="http://R-Forge.R-project.org")}.
+
+\subsection{The main object and classes}
+Before discussing the methods for simulation and inference for stochastic processes solutions to stochastic differential equations, here we discuss the main classes in the package.
+As mentioned there are different classes of objects defined in the \pkg{yuima} package and
+the main class is called the \code{yuima-class}. This class  is composed of several slots.
+Figure \ref{fig:classes} represents the different classes and their slots.
+\begin{figure}[th]
+\centering{\includegraphics[width=\textwidth]{yuima-class}}
+\caption{The main classes in the \pkg{yuima} package.}
+\label{fig:classes}
+\end{figure} 
+The different slots do not need to be all present at the same time.
+For example, in case one wants to simulate a stochastic process, only the slots \code{model} and \code{sampling} should be present, while the slot \code{data} will be filled by the simulator.
+We now discuss in details the different object separately.
+
+\subsection{The $\tt yuima.model$ class}
+At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified. All multidimensional and eventually as parametric models.
+\begin{itemize}
+\item diffusions $\displaystyle  \de X_t=a(t,X_t)dt + b(t,X_t)\de W_t $, where $W_t$ is a standard Brownian motion;
+\item fractional Gaussian noise, with $H$ the Hurst parameter
+$$ \de X_t=a(t,X_t)dt + b(t,X_t)\de W_t^{H}; $$ 
+\item diffusions with jumps and  L\'evy processes solution to
+$$
+\begin{aligned}
+\de X_t = & \,\,\, a(X_t)dt + b(X_t)\de W_t + \int\limits_{|z|>1}\!\!\! c(X_{t-},z)\mu(\de t,\de z) \\
+&{}+\!\! \int\limits_{0<|z|\le 1}\!\!\!  c(X_{t-},z)\{\mu(\de t,\de z) - \nu(\de z)\de t\}.
+\end{aligned}
+$$
+\end{itemize}
+
+The \code{yuima.model} class contains informations about the stochastic differential equation of interest. The constructor function \code{setModel} is
+used to give a mathematical description of the stochastic differential equation. 
+All functions in the package are assumed to get as much information
+as possible from the model instead of replicating the same code everywhere.
+If there are missing pieces of information, we may change or extend the description
+of the model.
+\\
+\\
+An object of class \code{yuima.model} contains several slots listed below. 
+To see inside its structure, one can use the \proglang{R} command \code{str} on a \pkg{yuima} object.
+\begin{itemize}
+\item \code{drift} is an \proglang{R} vector of expressions which contains the drift specification;
+\item \code{diffusion} is itself a list of 1 slot which describes the diffusion
+  coefficient matrix relative to first noise;
+\item \code{hurst} is the Hurst index of the fractional Brownian motion, by default \code{0.5}  meaning a standard Brownian motion;
+\item \code{parameter} which is a short name for ``parameters'' which 
+  is a list with the following entries:
+\begin{itemize}
+\item \code{all} contains the names of all the parameters found 
+   in the diffusion and drift coefficient;
+\item \code{common} contains the names of the parameters in common between the drift and diffusion coefficients;
+\item \code{diffusion} contains the parameters belonging to the diffusion coefficient;
+\item \code{drift} contains the parameters belonging to the drift coefficient;
+\item \code{jump} contains the parameters belonging to the  coefficient of the L\'evy noise;
+\item \code{measure} contains the parameters describing to the L\'evy measure;
+\end{itemize}
+\item \code{measure} specifies the measure of the L\'evy noise;
+\item \code{measure.type} a switch to specify how the L\'evy measure is described;
+\item \code{state.variable} and \code{time.variable}, by default,
+  are assumed to be \code{x} and \code{t} but the user can freely choose them.
+  The \code{yuima.model} class assumes that the user either use default
+  names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
+  the rest of the symbols are considered parameters and distributed accordingly
+  in the \code{parameter} slot.
+\item \code{jump.variable} the name of the variable used in the description of the L\'eve component;
+\item \code{solve.variable} contains a vector of variable names, each element corresponds to the
+   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order.
+\item \code{noise.number} indicates the number of sources of noise.
+\item \code{xinit} initial value of the stochastic differential equation;
+\item \code{equation.number} represents the number of equations, i.e. the number of one dimensional
+  stochastic differential equations.
+\item \code{dimension} reports the dimensions of the parameter space. It is a list of the
+  same length of \code{parameter} with the same names.
+\item \code{J.flag} for internal use only.
+\end{itemize}
+As seen in the above, the parameters space is accurately described internally in a \code{yuima} object because in inference for stochastic differential equations, estimators of different parameters have different properties.
+Usually, the rate of convergence for estimators in the diffusion coefficient are similar to the ones in the i.i.d.
+sampling while estimators of parameters in the drift coefficient are slower or, in some cases, not even consistent.
+The \pkg{yuima} always tries to implement the best statistical inference for the given model under the observed sampling scheme.
+
+\section{Model specification}
+In order to show how general is the approach in the \pkg{yuima} package we present some examples.
+
+\subsection{Diffusion processes}
+Assume that we want to describe the following stochastic differential equation
+$$\de X_t = -3 X_t \de t + \frac{1}{1+X_t^2}\de W_t$$
+This is done in \pkg{yuima} specifying the drift and diffusion coefficients as plain mathematical expressions
+<<print=FALSE,echo=FALSE,results=hide>>=
+library(yuima)
+@ 
+<<echo=TRUE, print=FALSE,results=hide>>=
+mod1 <- setModel(drift = "-3*x", diffusion = "1/(1+x^2)")
+@ 
+At this point, the package fills the proper slots of the \code{yuima} object
+{\scriptsize
+<<>>=
+str(mod1)
+@
+}
+
+For the above, it is possible to see that the jump coefficient is void and the Hurst parameter is set to 0.5, because this corresponds to the standard Brownian motion.
+Now, with \code{mod1} in hands, it is very easy to simulate a trajectory of the process as follows
+<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+set.seed(123)
+X <- simulate(mod1)
+plot(X)
+@
+
+\noindent
+The \code{simulate} function fills in addition the two slots \code{data} and \code{sampling} of the \code{yuima} object.
+{\scriptsize
+<<>>=
+str(X,vec.len=2)
+@
+}
+
+\subsection{Specification of parametric models}
+When a parametric model like
+$$\de X_t = -\theta X_t \de t + \frac{1}{1+X_t^\gamma}\de W_t$$
+is specified, \pkg{yuima} attempts to distinguish the parameters' names from the ones of the state and time variables
+<<echo=TRUE, print=FALSE,results=hide>>=
+mod2 <- setModel(drift = "-theta*x", diffusion = "1/(1+x^gamma)")
+@ 
+{\scriptsize
+<<>>=
+str(mod2)
+@
+}
+In order to simulate the parametric model it is necessary to specify the values of the parameters as the next code shows
+<<echo=TRUE, print=FALSE,fig=TRUE,height=4,results=hide>>=
+set.seed(123)
+X <- simulate(mod2,true.param=list(theta=1,gamma=3))
+plot(X)
+@
+
+\subsection{Multidimensional processes}
+Next is an example with two stochastic differential equations driven by three independent Brownian motions
+$$
+\begin{aligned}
+\de X^1_t &= -3 X^1_t \de t + \de W^1_t + X^2_t \de W^3_t\\
+\de X^2_t &= -(X^1_t + 2 X^2_t) \de t + X^1_t \de W^1_t + 3 \de W^2_t
+\end{aligned}
+$$
+but this has to be organized into matrix form
+$$
+\left(\begin{array}{c}\de X^1_t \\\de X^2_t\end{array}\right)=
+\left(\begin{array}{c} -3 X^1_t \\  -X^1_t - 2X^2_t\end{array}\right)\de t +
+\left[\begin{array}{ccc}1 & 0 & X^2_t \\X^1_t & 3 & 0\end{array}\right]
+\left(\begin{array}{c}\de W_t^1 \\ \de W_t^2 \\ \de W_t^3\end{array}\right)
+$$
+<<echo=TRUE, print=FALSE,results=hide>>=
+sol <- c("x1","x2") # variable for numerical solution
+a <- c("-3*x1","-x1-2*x2")   # drift vector 
+b <- matrix(c("1","x1","0","3","x2","0"),2,3)  #  diffusion matrix
+mod3 <- setModel(drift = a, diffusion = b, solve.variable = sol)
+@
+Again, this model can be easily simulated
+<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+set.seed(123)
+X <- simulate(mod3)
+plot(X, plot.type="single",lty=1:2)
+@
+
+\noindent
+But it is also possible to specify more complex models like the following
+$$
+ \begin{cases}
+ &\de X_t^1 = X_t^2 \left| X_t^1 \right|^{2/3} \de W_t^1, \\
+ &\de X_t^2 = g(t)\de X_t^3, \\
+ &\de X_t^3 = X_t^3( \mu  \de t + \sigma (\rho \de W_t^1 + \sqrt{1-\rho^2} \de W_t^2))
+ \end{cases},
+$$
+where  $g(t) = 0.4 + (0.1 + 0.2t) e^{-2t}$.
+
+\subsubsection{Fractional Gaussian noise}
+In order to specify a stochastic differential equation driven by fractional Gaussian noise it is necessary to specify the value of the Hurst parameter.
+For example, if we want to specify the following model
+$$\de Y_t =  3 Y_t \de t + \de W_t^H$$
+we proceed as follows
+<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+mod4 <- setModel(drift="3*y", diffusion=1, hurst=0.3, solve.var="y")
+set.seed(123)
+X <- simulate(mod4,sampling=setSampling(n=1000))
+plot(X)
+@
+
+\noindent
+In this case, the appropriate slot is now filled
+@ 
+{\scriptsize
+<<>>=
+str(mod4)
+@
+}
+
+\subsection{L\'evy processes}
+Jump processes can be specified in different ways in mathematics and hence in \pkg{yuima} package. 
+Let $Z_t$ be a  Compound Poisson Process (i.e. jumps size follow some distribution, like the Gaussian law and jumps occur at Poisson times).
+Then it is possible to consider the following SDE which involves jumps
+$$\de X_t =  a(X_t)dt + b(X_t)\de W_t + \de Z_t$$
+In the next example we consider a compound Poisson process with intensity $\lambda=10$ with Gaussian jumps.
+This model can be specified in \code{setModel} using the argument  \code{measure.type="CP"} 
+A simple Ornstein-Uhlembeck process with Gaussian jumps 
+$$\de X_t = -\theta X_t \de t + \sigma \de W_t + Z_t$$
+is specified as
+<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+mod5 <- setModel(drift=c("-theta*x"), diffusion="sigma",
+ jump.coeff="1", measure=list(intensity="10", df=list("dnorm(z, 0, 1)")),
+ measure.type="CP", solve.variable="x")
+set.seed(123)
+X <- simulate(mod5, true.p=list(theta=1,sigma=3),sampling=setSampling(n=1000))
+plot(X)
+@
+
+\noindent
+Another possibility is to specify the jump component via the L\'evy measure.
+Without going into too much details, here is an example of specification of a simple Ornstein-Uhlembeck  process with IG (Inverse Gaussian) L\'evy measure
+$$\de X_t = -x dt + dZ_t$$
+<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+mod6 <- setModel(drift="-x", xinit=1, jump.coeff="1", 
+  measure.type="code", measure=list(df="rIG(z, 1, 0.1)"))
+set.seed(123)
+X <- simulate(mod6, sampling=setSampling(Terminal=10, n=10000)) 
+plot(X)
+@
+
+\subsection{Specification of generic models}
+In general, the \pkg{yuima} package allows to specify a large family of models solutions to
+$$\de X_t =a(X_t)\de t +  b(X_t)\de W_t + c(X_t)\de Z_t$$
+using the following interface
+<<echo=TRUE, print=FALSE,eval=FALSE>>=
+setModel(drift, diffusion, hurst = 0.5, jump.coeff,
+ measure, measure.type, state.variable = "x",
+ jump.variable = "z", time.variable = "t", solve.variable, xinit)
+@
+The \pkg{yuima} package implements many multivariate Random Numbers Generators (RNG) which are needed to simulate L\'evy paths including
+\code{rIG} (Inverse Gaussian), \code{rNIG} (Normal Inverse Gaussian), \code{rbgamma} (Bilateral Gamma), \code{rngamma} (Gamma) and \texttt{rstable} (Stable Laws).
+Other user-defined RNG can be used freely.
+
+\subsection{Simulation of models and sampling}
+The \code{simulate} function simulates \code{yuima} models according to Euler-Maruyama scheme in the presence of non-fractional diffusion noise and L\'eve jumps and uses the Cholesky or the Wood and Chan methods for the fractional Gaussian noise.
+The \code{simulate} function accepts several arguments including the description of the sampling structure, which is an object of type \code{yuima.sampling}.
+The \code{setSampling} allow for the specification of different sampling parameters including random sampling.
+Further, the \code{subsampling} allow to subsample a trajectory of a simulated stochastic differential equation or a given time series in the \code{yuima.data} slot of a \code{yuima} object.
+Sampling and subsampling can be specified jointly as arguments to the \code{simulate} function.
+This is convenient if one wants to simulate data at very high frequency but then return only low frequency data for inference or other applications.
+
+\section{Miscellanea}
+The code \pkg{yuima} package offers several other utility and extensions from the main core classes. We will review some of these.
+
+\subsection{Asymptotic expansion}
+The \pkg{yuima} package can handle asymptotic expansion of functionals of  $d$-dimensional diffusion process 
+$$\de X_t^\ve = a(X_t^\ve,\ve)\de t + b(X_t^\ve,\ve)\de W_t, \qquad \ve \in(0,1]$$
+with $W_t$ and $r$-dimensional Wiener process, i.e. $W_t=(W_t^1, \ldots, W_t^r)$.
+The functional is expressed in the following abstract form
+$$
+F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve), \qquad W_t^0=t
+$$ 
+A typical example of application is the case of Asian option pricing. For example, in the Black \& Scholes model
+$$\de X_t^\ve = \mu X_t^\ve\de t + \ve X_t^\ve \de W_t$$ 
+the price of the option is of the form
+$$ \dE\left\{ \max\left( \frac{1}{T}\int_0^T X_t^\ve \de t - K,0\right)\right\}.$$
+Thus the functional of interest is
+$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$ 
+with
+$$
+f_0(x,\ve) = \frac{x}{T}, \quad f_1(x,\ve)=0, \quad  F(x, \ve) = 0
+$$
+in 
+$$
+F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve)
+$$ 
+So, the call option price requires the composition  of a smooth functional
+$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$
+with the irregular function
+$$
+\max(x-K,0)
+$$
+Monte Carlo methods require a huge number of simulations to get the desired accuracy of the
+calculation of the price, while asymptotic expansion of $F^\ve$ provides very accurate approximations.
+The \pkg{yuima} package provides functions to construct the functional $F^\ve$, and automatic asymptotic expansion based on Malliavin calculus starting from a \pkg{yuima} object.
+Next is an example
+<<echo=TRUE, print=FALSE, results=hide>>=
+diff.matrix <- matrix( c("x*e"), 1,1)
+model <- setModel(drift = c("x"), diffusion = diff.matrix)
+T <- 1
+xinit <- 1
+K <- 1
+f <- list( expression(x/T), expression(0))
+F <- 0
+e <- .3
+yuima <- setYuima(model = model, sampling = setSampling(Terminal=T, n=1000))
+yuima <- setFunctional( yuima, f=f,F=F, xinit=xinit,e=e)
+@
+this time the \code{setFunctional} command fills the appropriate slots
+<<echo=TRUE, print=FALSE>>=
+str(yuima@functional)
+@
+Then, it is as easy as
+<<echo=TRUE, print=FALSE>>=
+F0 <- F0(yuima)
+F0
+max(F0-K,0)  # asian call option price
+@
+to obtain the zero order approximation of the value of the functional.
+We can go up to the first order approximation adding one term to the expansion
+<<echo=TRUE, print=FALSE,results=hide>>=
+rho <- expression(0)
+get_ge <- function(x,epsilon,K,F0){
+  tmp <- (F0 - K) + (epsilon * x) 
+  tmp[(epsilon * x) < (K-F0)] <- 0
+  return( tmp )
+}
+epsilon <- e  # noise level
+g <- function(x) {
+  tmp <- (F0 - K) + (epsilon * x) 
+ tmp[(epsilon * x) < (K-F0)] <- 0
+ tmp
+}
+asymp <- asymptotic_term(yuima, block=10, rho, g)
+@
+and the final value is
+<<echo=TRUE, print=FALSE>>=
+asymp$d0 + e * asymp$d1  # asymp. exp. of asian call price
+@
+
+\subsection{Export of a \code{yuima} model}
+The \pkg{yuima} implements the function \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}. A simple writing like
+\code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper.
+
+\section{Inference for stochastic processes}
+The \pkg{yuima} implements several optimal techniques for parametric and nonparametric estimation of multidimensional stochastic differential equations.
+Although most of the examples in this section are given on simulated data, the main way to fill up the \code{data} slot of a \code{yuima} object is to use the function \code{setYuima}.
+The function \code{setYuima} sets various slots of the \code{yuima} object.
+In particular, to estimate a \code{yuima.model} called \code{mod} on the data \code{X} one can use a code like this \code{my.yuima <- setYuima(data=setData(X), model=mod)} and then pass \code{my.yuima} to the inference functions as described in what follows.
+
+\subsection{Quasi Maximum Likelihood estimation}
+Consider the multidimensional diffusion process
+$$
+\de X_t = b(\theta_2,X_t)\de t + \sigma(\theta_1, X_t) \de W_t
+$$
+where $W_t$ is an {$r$}-dimensional standard Wiener process 
+independent of the initial value $X_0=x_0$. 
+Quasi-MLE assumes the following approximation of the true log-likelihood for multidimensional diffusions
+{\small
+\begin{eqnarray}\label{qlik}
+\ell_n({\bf X}_n,\theta)
+=-\frac12\sum_{i=1}^n\left\{\log\text{det}(\Sigma_{i-1}(\theta_1))
++\frac{1}{\Delta_n}
+\Sigma_{i-1}^{-1}(\theta_1)[\Delta X_i-\Delta_n b_{i-1}(\theta_2)]^{\otimes 2}\right\}
+\end{eqnarray}}
+where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, $A^{\otimes 2}= A^T A$ and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
+Then, \cite{Yoshida92}, the QML estimator of $\theta$ is
+$$\tilde\theta_n=\arg\min_\theta \ell_n({\bf X}_n,\theta)$$
+As an example, we consider the simple model
+\begin{equation}
+\de X_t = -\theta_2  X_t  \de t + \theta_1  \de W_t
+\label{eq:model1}
+\end{equation}
+with $\theta_1=0.3$ and $\theta_2=0.1$
+<<echo=TRUE, print=FALSE,results=hide>>=
+ymodel <- setModel(drift = "-x*theta2", diffusion = "theta1",
+     time.variable = "t", state.variable = "x", solve.variable = "x")
+n <- 1000
+ysamp <- setSampling(Terminal = (n)^(1/3), n = n)
+yuima <- setYuima(model = ymodel, sampling = ysamp)
+set.seed(123)
+yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.3, theta2 = 0.1))
+@
+With the simulated path we can use the function \code{qmle} to estimate the parameters as follows
+<<echo=TRUE, print=FALSE,results=hide>>=
+mle1 <- qmle(yuima, start = list(theta1 = 0.8, theta2 = 0.7), 
+     lower = list(theta1=0.05, theta2=0.05), 
+     upper = list(theta1=0.5, theta2=0.5), method = "L-BFGS-B")
+@
+and the estimated coefficients are as follows
+<<echo=TRUE, print=FALSE>>=
+coef(mle1)
+summary(mle1)
+@
+Notice the interface and the output of the \code{qmle} is quite similar to the ones of the standard \code{mle} function of the \pkg{stats4} package of the base \proglang{R} system.
+
+\subsection{Adaptive Bayes estimation}
+Consider again the  diffusion process solution to
+\begin{equation} 
+\de X_t=b(X_t,\theta_2)\de t+\sigma(X_t,\theta_1)\de W_t,
+\end{equation}
+and the quasi likelihood defined in \eqref{qlik}.
+
+The adaptive Bayes type estimator is defined as follows. 
+First we choose an initial arbitrary value $\theta_2^\star\in\Theta_2$ and 
+pretend $\theta_1$ is the unknown parameter to 
+make the Bayesian type estimator $\tilde{\theta}_1$ as 
+\begin{equation} 
+\tilde{\theta}_1
+=  
+\Big[\int_{\Theta_1}\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
+\pi_1(\theta_1)d\theta_1 \Big]^{-1} \int_{\Theta_1} \theta_1 \ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
+\pi_1(\theta_1)d\theta_1
+\end{equation}
+where 
+$\pi_1$ is a prior density on $\Theta_1$. 
+According to the asymptotic theory, 
+if $\pi_1$ is positive on $\Theta_1$, any function can be used. 
+For estimation of $\theta_2$, we use $\tilde{\theta}_1$ 
+to reform the quasi-likelihood function. That is,
+the Bayes type estimator for $\theta_2$ is defined by 
+\begin{equation} 
+\tilde{\theta}_2
+= 
+\Big[\int_{\Theta_2}\ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))
+\pi_2(\theta_2)d\theta_2 \Big]^{-1} \int_{\Theta_2} \theta_2 \ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))
+\pi_2(\theta_2)d\theta_2
+\end{equation}
+where 
+$\pi_2$ is a prior density on $\Theta_2$. 
+In this way, we obtain the adaptive Bayes type estimator 
+$\tilde{\theta}=(\tilde{\theta}_1,\tilde{\theta}_2)$ 
+for $\theta=(\theta_1,\theta_2)$.
+
+Adaptive Bayes estimation is developed in \pkg{yuima} via the method $\tt adaBayes$.
+Consider again the model \eqref{eq:model1} with the same values for the parameters, i.e.
+ $\theta_1=0.3$ and $\theta_2=0.1$
+In order to perform Bayesian estimation, we need to prepare the prior densities for the parameters.
+For simplicity we use uniform distributions in $[0,1]$
+<<echo=TRUE>>=
+prior <- list(theta2=list(measure.type="code",df="dunif(z,0,1)"), theta1=list(measure.type="code",df="dunif(z,0,1)"))
+@
+Then we call $\tt adaBayes$ as follows
+<<eval=TRUE,echo=TRUE, results=hide>>=
+param.init <- list(theta2=0.5,theta1=0.5)
+bayes1 <- adaBayes(yuima, start=param.init, prior=prior, method="nomcmc")
+@
+and we can compare the adaptive Bayes estimates  with the QMLE estimates
+<<eval=TRUE,echo=TRUE>>=
+bayes1@coef
+coef(mle1)
+@
+The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical integration, otherwise MCMC method is used.
+
+\subsection{Asynchronous covariance estimation}
+Suppose that two It\^o processes are observed
+only at discrete times in a nonsynchronous manner.
+We are interested in estimating the covariance of the two processes accurately
+in such a situation.
+This type of problem arises typically in high-frequency financial time series.
+
+Let $T \in (0,\infty)$ be a terminal time for possible observations. 
+We consider a two dimensional It\^o process $(X^1,X^2)$  
+satisfying the stochastic differential equations
+\begin{eqnarray*}
+\mbox{d}X_t^l &=&\mu_t^l \mbox{d}t + \sigma_t^l \mbox{d}W_t^l ,\quad
+t\in[0,T]\\
+X_0^l &=& x_0^l        \notag
+\end{eqnarray*}
+for $l=1,2$. 
+Here $W^l$ denote standard Wiener processes with a 
+progressively measurable correlation process 
+$\mbox{d}\langle W_1 , W_2 \rangle_t = \rho_t \mbox{dt}$,  
+$\mu^l_t$ and $\sigma_t^l$ are progressively measurable processes, 
+and $x^l_0$ are initial random variables independent of $(W^1,W^2)$.  
+Diffusion type processes are in the scope but this model 
+can express more sophisticated stochastic structures.
+
+The process $X^l$ is supposed to be observed at 
+over the increasing sequence of times 
+$T^{l,i}$ $(i\in\mathbb{Z}_{\geq0})$ starting at $0$, up to time T. 
+Thus, the observables are $(T^{l,i},X^{l,i})$ with $T^{l,i}\leq T$. 
+Each $T^{l,j}$ may be a stopping time, so possibly depends on 
+the history of $(X^1,X^2)$ as well as the precedent stopping times.  
+Two sequences of stopping times 
+$T^{1,i}$ and $T^{2,j}$ are \textit{nonsynchronous},  and 
+irregularly spaced, in general. 
+In particular, cce can apply to estimation of the quadratic variation of a single 
+stochastic process sampled regularly/irregularly.
+
+The parameter of interest is the quadratic covariation between $X^1$ and $X^2$:
+
+\begin{equation}
+\theta=
+\langle X^1 , X^2 \rangle_T = \int_0^T \sigma_t^1 \sigma_t^2 \rho_t \mbox{d}t.
+\end{equation}
+The target variable $\theta$ is random in general.
+
+It can be estimated with the nonsynchronous covariance estimator 
+(Hayashi-Yoshida estimator) 
+\begin{equation}
+U_n= \sum_{i,j:T^{1,i}\leq T, T^{2,j}\leq T} (X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)
+1_{\{ (T^{1,i-1},T^{1,i}] 
+\bigcap (T^{2,j-1},T^{2,j}]  \neq \emptyset  \}}.
+\end{equation}
+That is, the product of any pair of increments $(X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)$ and
+$(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)$ will make a contribution 
+to the sum only when 
+the respective observation intervals $(T^{1,i-1},T^{1,i}] $ and $ (T^{2,j-1},T^{2,j}] $ are overlapping 
+with each other. 
+%The estimator $U_n$ was proposed by Hayashi and Yoshida  and 
+It is known that $U_n$ is consist and has asymptotically mixed normal 
+distribution 
+as $n\to\infty$ if the maximum length between two consecutive observing times 
+tends to $0$. 
+See \citet{hay-yos05,hay-yos04,hay-yos06,hay-yos08} for details.
+
+\subsubsection{Example: data generation and estimation by \pkg{yuima} package}
+We will demonstrate how to apply cce function to 
+nonsynchronous high-frequency data by simulation. 
+As an example, consider a two dimensional stochastic process 
+$(X^1_t,X^2_t)$ satisfying the stochastic differential equation
+\begin{equation}
+\begin{split}
+\mbox{d}X^1_t = \sigma_{1,t} \mbox{d}B^1_t, \\
+\mbox{d}X^2_t = \sigma_{2,t} \mbox{d}B^2_t. 
+\end{split}
+\end{equation}
+Here $B^1_t$ and $B^2_t$ denote two standard Wiener processes, 
+however they are correlated as 
+\begin{eqnarray}
+B^1_t &=& W^1_t,\\
+B^2_t &=&  \int_0^t \rho_s \mbox{d} W^1_s +
+  \int_0^t \sqrt{ 1- \rho_s^2} \mbox{d} W^2_s,
+\end{eqnarray}
+where  $W^1_t$ and $W^2_t$ are independent Wiener processes,  and 
+$\rho_t$ is the correlation function between $B^1_t$ and $B^2_t$. 
+We consider $\sigma_{l,t},l=1,2$ and $\rho_t$ 
+of the following form in this example:
+\begin{eqnarray*}
+ \sigma_{1,t} &=& \sqrt{1+t}, \\
+ \sigma_{2,t} &=& \sqrt{1+t^2}, \\
+ \rho_t &=& \frac{1}{\sqrt{2}}.
+\end{eqnarray*}
+To simulate the stochastic process $(X^1_t,X^2_t)$, we first build the model 
+by $\tt setModel$ as before. 
+It should be noted that the method of generating 
+nonsynchronous data can be replaced by a simpler one but 
+we will take a general approach here to demonstrate 
+a usage of 
+the \pkg{yuima} comprehensive package for simulation and estimation of 
+stochastic processes. 
+<<print=FALSE,echo=TRUE>>=
+# diffusion coefficient for process 1
+diff.coef.1 <- function(t,x1=0, x2=0) sqrt(1+t)
+# diffusion coefficient for process 2
+diff.coef.2 <- function(t,x1=0, x2=0) sqrt(1+t^2)
+# correlation
+cor.rho <- function(t,x1=0, x2=0) sqrt(1/2)
+# coefficient matrix for diffusion term
+diff.coef.matrix <- matrix( c( "diff.coef.1(t,x1,x2)", 
+"diff.coef.2(t,x1,x2) * cor.rho(t,x1,x2)", "", 
+"diff.coef.2(t,x1,x2) * sqrt(1-cor.rho(t,x1,x2)^2)"),2,2)
+# Model sde using yuima.model
+cor.mod <- setModel(drift = c("",""), diffusion = diff.coef.matrix, solve.variable=c("x1","x2"))
+@
+
+The parameter we want to estimate is the quadratic covariation between $X_1$ and $X_2$: 
+\begin{equation}
+\theta = \langle X_1, X_2 \rangle_T =
+ \int_0^T \sigma_{1,t} \sigma_{2,t} \rho_t \mbox{d} t.
+\end{equation}
+Later, we will compare estimated values with 
+the true value of  $\theta$ given by 
+<<echo=TRUE>>=
+CC.theta <- function( T, sigma1, sigma2, rho)
+{
+ 	tmp <- function(t) return( sigma1(t) * sigma2(t) * rho(t) )
+ 	integrate(tmp,0,T)
+}
+@ 
+For the sampling scheme, we will consider the independent Poisson sampling. 
+That is, each configuration of the sampling times $T^{l,i}$ is realized 
+as the Poisson random measure with intensity $np_l$, 
+and the two random measures are independent each other as well as 
+the stochastic processes. 
+Then it is known from today's lecture that
+\begin{equation}
+n^{1/2} ( U_n -\theta) \rightarrow N(0,c),
+\end{equation}
+as $n\to\infty$, 
+where 
+\begin{equation}
+ c = \left( \frac{2}{p_1} + \frac{2}{p_2} \right)
+\int_0^T \left( \sigma_{1,t} \sigma_{2,t}  \right)^2 \mbox{d}t +
+\left(
+\frac{2}{p_1} + \frac{2}{p_2} - \frac{2}{p_1 + p_2}
+\right)
+\int_0^T
+\left(
+\sigma_{1,t} \sigma_{2,t} \rho_t
+\right)^2 \mbox{d} t.
+\label{eq:vc}
+\end{equation}
+
+<<>>=
+set.seed(123)
+
+Terminal <- 1
+n <- 1000
+# Cumulative Covariance
+theta <- CC.theta(T=Terminal, sigma1=diff.coef.1, sigma2=diff.coef.2, rho=cor.rho)$value
+cat(sprintf("theta=%5.3f\n",theta))
+@
+so in our case $\theta=\Sexpr{round(theta,2)}$.
+<<results=hide>>=
+yuima.samp <- setSampling(Terminal=Terminal,n=n)
+yuima <- setYuima(model=cor.mod, sampling=yuima.samp)
+X <- simulate(yuima)
+@
+$\tt cce$ takes the sample and returns 
+an estimate of the quadratic covariation. For example, for the complete data
+<<>>=
+cce(X)
+@
+and we now apply random sampling
+<<>>=
+p1 <- 0.2
+p2 <- 0.3
+newsamp <- setSampling(
+ random=list(rdist=c( function(x) rexp(x, rate=p1*n/Terminal), 
+  function(x) rexp(x, rate=p1*n/Terminal))) )
+Y <- subsampling(X, sampling=newsamp)
+cce(Y)
+@ 
+Now we calculate the asymptotic variance of the estimator using \eqref{eq:vc}
+<<>>=
+# asymptotic variance
+var.c <- function(T, p1,p2, sigma1, sigma2, rho)
+{
+  tmp_integrand1 <- function(t) (sigma1(t) * sigma2(t))^2
+  i1 <- integrate(tmp_integrand1,0,T)
+  tmp_integrand2 <- function(t) (sigma1(t) * sigma2(t) * rho(t))^2
+  i2 <- integrate(tmp_integrand2,0,T)
+  2*(1/p1 + 1/p2)* i1$value + 2*(1/p1+1/p2 - 1/(p1+p2)) * i2$value
+}
+vc <- var.c(T=Terminal, p1, p2, diff.coef.1, diff.coef.2, cor.rho)
+sqrt(vc/n)
+@
+
+\subsection{Change point analysis}
+Consider a multidimensional stochastic differential equation of the form
+$$
+\de Y_t = b_t \de t + \sigma(X_t,\theta) \de W_t,\ \ t\in[0,T], 
+$$
+where $W_t$ a $r$-dimensional Wiener process and $b_t$ and $X_t$ are multidimensional processes and $\sigma$ is the diffusion coefficient (volatility) matrix.
+When $Y=X$ the problem is a diffusion model.
+The process $b_t$ may have jumps but should not explode and it is treated as a nuisance in this model.
+The change-point problem for the volatility is formalized as follows
+$$
+Y_t=
+\Bigg\{
+\begin{array}{ll}
+Y_0+\int_0^t b_s d s+\int_0^t \sigma(X_s,\theta_1^*) d W_s
+& \mbox{ for } t\in[0,\tau^*)
+\\
+Y_{\tau^*}+\int_{\tau^*}^t b_s d s+\int_{\tau^*}^t \sigma(X_s,\theta_2^*) d W_s
+& \mbox{ for } t\in[\tau^*,T]. 
+\end{array}
+$$
+The change point $\tau^*$ instant is unknown and 
+is  to be estimated, along with $\theta_1^*$ and $\theta_2^*$, from the observations sampled 
+from the path of $(X,Y)$. The \pkg{yuima} implements the quasi-maximum likelihood approach as in \citet{iacyos09} described in the following.
+Let  $\Delta_iY=Y_{t_i}-Y_{t_{i-1}}$ and 
+define 
+\begin{equation} 
+\Phi_n(t;\theta_1,\theta_2)
+=
+\sum_{i=1}^{[nt{/T}]}G_i(\theta_1)+\sum_{i={[nt{/T}]+1}}^nG_i(\theta_2),
+\label{eq:phiCH10}
+\end{equation}
+with 
+\begin{equation} 
+G_i(\theta) =
+\log\det S(X_{t_{i-1}},\theta)
++ \Delta_n^{-1}(\Delta_iY)'S(X_{t_{i-1}},\theta)^{-1}(\Delta_iY). 
+\label{eq:GiCH10}
+\end{equation}
+Suppose that there exists an estimator 
+$\hat{\theta}_k$ for each $\theta_k$, $k=1,2$. 
+In case $\theta_k^*$ are known, we define $\hat{\theta}_k$ 
+just as $\hat{\theta}_k=\theta_k^*$. 
+The change point estimator  of $\tau^*$ is
+\begin{eqnarray*} 
+\hat{\tau}_n&=&\arg\!\!\min\limits_{t\in[0,T]} 
+\Phi_n(t;\hat{\theta}_1,\hat{\theta}_2).
+\end{eqnarray*}
+
+\subsection{Example  of Volatility Change-Point Estimation}
+Consider the 2-dimensional stochastic differential equation
+$$
+\left(\begin{array}{c}
+ \de X_t^1\\ \de X_t^2 
+ \end{array}\right)
+ = 
+\left(\begin{array}{c}
+ 1 - X^1_t \\ 
+ 3 - X^2_t \\ 
+ \end{array}\right) \de t
+ +
+\left[\begin{array}{cc}
+ \theta_{1.1}  \cdot  X_t^1&0  \cdot  X_t^1 \\ 
+ 0  \cdot  X_t^2&\theta_{1.2}  \cdot  X_t^2 \\ 
+ \end{array}\right]
+'
+\left(\begin{array}{c}
+ \de W_t^1\\ \de W_t^2 
+ \end{array}\right)
+$$
+$$
+X^1_0=1.0,\quad
+X^2_0=1.0,\quad
+$$
+with change point instant at time $\tau=0.4$.  Some code is needed to simulate such a process. First we define the model
+<<cpoint1,eval=TRUE, echo=TRUE,results=hide>>=
+diff.matrix <- matrix(c("theta1.1*x1","0*x2","0*x1","theta1.2*x2"), 2, 2)
+drift.c <- c("1-x1", "3-x2")
+drift.matrix <- matrix(drift.c, 2, 1)
+ymodel <- setModel(drift=drift.matrix, diffusion=diff.matrix, time.variable="t",
+state.variable=c("x1", "x2"), solve.variable=c("x1", "x2"))
+@
+and then simulate two trajectories. One up to the change point $\tau=4$ with parameters $\theta_{1.1}=0.1$ and  $\theta_{1.1}=0.2$, and a second trajectory with parameters
+ $\theta_{1.1}=0.6$ and  $\theta_{1.2}=0.6$. For the second trajectory, the initial value is set to the last value of the first trajectory.
+<<cpoint3,results=hide>>=
+n <- 1000
+
+set.seed(123)
+
+t1 <- list(theta1.1=.1, theta1.2=0.2)
+t2 <- list(theta1.1=.6, theta1.2=.6)
+
+tau <- 0.4
+ysamp1 <- setSampling(n=tau*n, Initial=0, delta=0.01)
+yuima1 <- setYuima(model=ymodel, sampling=ysamp1)
+yuima1 <- simulate(yuima1, xinit=c(1, 1), true.parameter=t1)
+
+x1 <- yuima1@data@zoo.data[[1]]
+x1 <- as.numeric(x1[length(x1)])
+x2 <- yuima1@data@zoo.data[[2]]
+x2 <- as.numeric(x2[length(x2)])
+
+ysamp2 <- setSampling(Initial=n*tau*0.01, n=n*(1-tau), delta=0.01)
+yuima2 <- setYuima(model=ymodel, sampling=ysamp2)
+yuima2 <- simulate(yuima2, xinit=c(x1, x2), true.parameter=t2)
+yuima <- yuima1
+yuima@data@zoo.data[[1]] <- c(yuima1@data@zoo.data[[1]], yuima2@data@zoo.data[[1]][-1])
+yuima@data@zoo.data[[2]] <- c(yuima1@data@zoo.data[[2]], yuima2@data@zoo.data[[2]][-1])
+@ 
+The composed trajectory appears as follows
+<<cpoint4,fig=TRUE,width=9,height=5>>=
+plot(yuima)
+@
+
+\noindent
+Just as an example, we test the ability of the change point estimator to identify $\tau$ when for given true values of the parameters $\theta_{1.1}$ and $\theta_{1.2}$
+<<cpoint5>>=
+t.est <- CPoint(yuima,param1=t1,param2=t2, plot=TRUE)
+t.est$tau
+@
+A two stage change point estimation approach is available as explained in  \citet{iacyos09}.
+
+\subsection{LASSO model selection}
+Let $X_t$ be a  diffusion process solution to
+$$
+\de X_t = b(\alpha, X_t) \de t + \sigma(\beta,X_t)  \de W_t
+$$
+$$\alpha=(\alpha_1,...,\alpha_{p})'\in\Theta_p\subset \mathbb{R}^p, \quad p\geq1$$
+$$\beta=(\beta_1,...,\beta_q)'\in\Theta_q\subset \mathbb{R}^q, \quad q\geq1$$
+with $b:\Theta_p\times\mathbb{R}^d\to \mathbb{R}^d$, $\sigma:\Theta_q\times \mathbb{R}^d\to \mathbb{R}^d\times \mathbb{R}^m$ and $W_t,t\in [0,T]$, is a standard Brownian motion in $\mathbb{R}^m$. 
+We assume that the functions $b$ and $\sigma$ are known up to   $\alpha$ and $\beta$.
+We denote by $\theta=(\alpha,\beta)\in\Theta_p\times \Theta_q=\Theta$ the parametric vector and with $\theta_0=(\alpha_0,\beta_0)$ its unknown true value. 
+Let $\mathbb{H}_n({\bf X}_n,\theta) = \ell_n({\bf X}_n,\theta)$ from equation \eqref{qlik}.
+The quasi-MLE $\tilde{\theta}_n$ for this model is the solution of the following problem
+$$
+\tilde{\theta}_n=(\tilde\alpha_n,\tilde\beta_n)'=\arg\min_\theta \mathbb{H}_n({\bf X}_n,\theta)
+$$ 
+The adaptive LASSO estimator is defined as the solution to the quadratic problem under $L_1$ constraints
+$$
+\hat{\theta}_n=(\hat\alpha_n,\hat\beta_n)=\arg\min_\theta\mathcal{F}(\theta).
+$$
+with
+$$
+\mathcal{F}(\theta)=(\theta-\tilde{\theta}_n)\ddot{\mathbb{H}}_n({\bf X}_n, \tilde\theta_n)(\theta-\tilde{\theta}_n)'+\sum_{j=1}^p\lambda_{n,j}|\alpha_j| +\sum_{k=1}^q\gamma_{n,k}|\beta_k|
+$$
+For more details see \citet{DegIac10b}.
+The tuning parameters should be chosen as in \citet{Zou06} in the following way
+\begin{equation}
+\label{eq:penalty}
+\lambda_{n,j} = \lambda_0 |\tilde \alpha_{n,j}|^{-\delta_1}, \qquad
+\gamma_{n,k} = \gamma_0 |\tilde \beta_{n,j}|^{-\delta_2}
+\end{equation}
+where $\tilde \alpha_{n,j}$ and  $\tilde \beta_{n,k}$ are the unpenalized QML estimator of $\alpha_j$ and $\beta_k$ respectively, $\delta_1, \delta_2>0$ and usually taken unitary.
+
+\subsection{An example of use}
+The \code{lasso} method is implemented in the \pkg{yuima} package.
+Let us consider the full CKLS model
+$$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
+and let us try to estimate the parameter on  the U.S. Interest Rates monthly data from 06/1964 to 12/1989. We prepare the data, the model and the constraints for optimization
+<<results=hide>>=
+library(Ecdat)
+data(Irates)
+rates <- Irates[,"r1"]
+plot(rates)
+X <- window(rates, start=1964.471, end=1989.333)
+mod <- setModel(drift="alpha+beta*x", diffusion=matrix("sigma*x^gamma",1,1))
+yuima <- setYuima(data=setData(X), model=mod)
+lambda10 <- list(alpha=10, beta =10, sigma =10, gamma =10)
+start <- list(alpha=1, beta =-.1, sigma =.1, gamma =1)
+low <- list(alpha=-5, beta =-5, sigma =-5, gamma =-5)
+upp <- list(alpha=8, beta =8, sigma =8, gamma =8)
+@
+and now we apply the \code{lasso} function
+<<echo=TRUE,results=hide>>=
+lasso10 <- lasso(yuima, lambda10, start=start, lower=low, upper=upp,
+   method="L-BFGS-B")
+@
+From which we see that, instead of the general model
+$$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
+<<>>=
+round(lasso10$mle, 2)
+round(lasso10$lasso, 2)
+@
+the LASSO method selects the reduced model
+$$\de X_t =  0.6 \de t + 0.12 X_t^{\frac32}\de W_t$$
+
+\section{The future of YUIMA}
+some words
+
+\section*{Acknowledgements}
+some thanks
+
+%\bibliographystyle{natbib}
+
+\bibliography{bibliography}
+
 \end{document}

### 2011-09-13 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@183 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Alexandre Brouste\\University of Le Mans \And 
-        Stefano M. Iacus\\University of Milan \And 	
-        Hiroki Masuda\\University of Kyushu \And
-        Masaaki Fukasawa\\University of ???\AND
+        Masaaki Fukasawa\\University of Osaka\And
+        Hideitsu Hino\\Waseda University\And
+        Stefano M. Iacus\\University of Milan \AND 	
         Kengo Kamatani\\University of Tokyo \And
+        Koike Yuta\\University of Tokyo\And
+        Hiroki Masuda\\University of Kyushu \AND
+        Ryosuke Nomura\\University of Tokyo\And
         Yusutaka Shimuzu\\University of Osaka \And
@@
 \begin{equation}
-\de X_t = -\theta_2  X_t  \de t + \theta_1  \de W_t
+\de X_t = (2-\theta_2  X_t)  \de t + (1+X_t^2)^{\theta_1}  \de W_t
 \label{eq:model1}
 \end{equation}
-with $\theta_1=0.3$ and $\theta_2=0.1$
+with $\theta_1=0.3$ and $\theta_2=0.2$
 <<echo=TRUE, print=FALSE,results=hide>>=
-ymodel <- setModel(drift = "-x*theta2", diffusion = "theta1",
-     time.variable = "t", state.variable = "x", solve.variable = "x")
-n <- 1000
-ysamp <- setSampling(Terminal = (n)^(1/3), n = n)
+ymodel <- setModel(drift="(2-theta2*x)", diffusion="(1+x^2)^theta1")
+n <- 500
+ysamp <- setSampling(Terminal = n^(1/3), n = n)
 yuima <- setYuima(model = ymodel, sampling = ysamp)
 set.seed(123)
-yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.3, theta2 = 0.1))
+yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.2, theta2 = 0.3))
 @
@@
 <<echo=TRUE, print=FALSE,results=hide>>=
-mle1 <- qmle(yuima, start = list(theta1 = 0.8, theta2 = 0.7), 
-     lower = list(theta1=0.05, theta2=0.05), 
-     upper = list(theta1=0.5, theta2=0.5), method = "L-BFGS-B")
+param.init <- list(theta2=0.5,theta1=0.5)
+mle1 <- qmle(yuima, start =param.init , 
+     lower = list(theta1=0, theta2=0), 
+     upper = list(theta1=1, theta2=1))
 @
@@
 <<echo=TRUE, print=FALSE>>=
-coef(mle1)
 summary(mle1)
@@
 <<echo=TRUE>>=
-prior <- list(theta2=list(measure.type="code",df="dunif(z,0,1)"), theta1=list(measure.type="code",df="dunif(z,0,1)"))
+prior <- list(theta2=list(measure.type="code",df="dunif(theta2,0,1)"), theta1=list(measure.type="code",df="dunif(theta1,0,1)"))
 @
@@
 <<eval=TRUE,echo=TRUE, results=hide>>=
-param.init <- list(theta2=0.5,theta1=0.5)
 bayes1 <- adaBayes(yuima, start=param.init, prior=prior, method="nomcmc")
@@
 <<eval=TRUE,echo=TRUE>>=
-bayes1@coef
+coef(bayes1)
 coef(mle1)

### 2011-09-13 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@184 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
         Kengo Kamatani\\University of Tokyo \And
-        Koike Yuta\\University of Tokyo\And
-        Hiroki Masuda\\University of Kyushu \AND
-        Ryosuke Nomura\\University of Tokyo\And
+        Yuta Koike\\University of Tokyo\And
+        Hiroki Masuda\\University of Kyushu \And
+        Ryosuke Nomura\\University of Tokyo\AND
         Yusutaka Shimuzu\\University of Osaka \And
@@
 \section{Introduction}
-The YUIMA Project\footnote{The Project has been  funded up to 2010 by the Japan Science Technology (JST) Basic Research Programs PRESTO, Grants-in-Aid for Scientific Research No.
-19340021.} is an open source\footnote{All code in the \pkg{yuima} package is subject to the GNU General Public License, Version 2, see \url{http://www.gnu.org/licenses/gpl-2.0.html}.}
- academic project aimed at developing the \proglang{R} package named ``\pkg{yuima}'' for simulation and inference of stochastic differential equations. 
-The YUIMA Project is mainly developed by mathematicians and 
-statisticians who actively publish in the field of inference and simulation for stochastic
-differential equations.
-%The YUIMA Project Core Team, currently 
-%consists of the following people: A. Brouste, M. Fukasawa, H. Hino, S.M. Iacus, K. Kamatani, H.Masuda, Y. Shimizu, M. Uchida, N. Yoshida.
-
- The \pkg{yuima} package  provides 
+The plan of the YUIMA Project is to define the bases for a complete environment for simulation and inference for stochastic processes via an \proglang{R} package called \pkg{yuima}.
+ The package \pkg{yuima} provides 
 an object-oriented programming environment 
@@
 for stochastic processes by \proglang{R}.
-The \pkg{yuima} package adopts the  $\tt S4$ system of classes and methods \citep{chambers98}.
+The \pkg{yuima} package adopts the  \texttt{S4} system of classes and methods \citep{chambers98}.
 
@@
 The formulas produced by the theory are usually very sophisticated, 
-which makes it difficult for practitioners not necessarily 
-familiar with this field to enjoy their utility. 
+which makes it difficult for standard users not necessarily 
+familiar with this field to enjoy utilities. 
 For example, the asymptotic expansion method for computing 
-asian option prices (i.e., expectation of a functional of 
+option prices (i.e., expectation of an irregular functional of 
 a stochastic process) provides precise approximation values 
 instantaneously, taking advantage of the analytic approach, 
-but the formula, based on Malliavin calculus, has a long expression like more than one page!
+but the formula has a long expression like more than one page!
 
@@
 with simulation and/or statistics for stochastic differential equations. 
-
-Sampled data from a continuous-time process features 
-the time stamps as well as the positions of the object. 
-It is requiring a new theory of estimation. 
-The \pkg{yuima} framework can apply multi-dimensional time stamps 
-of tick data and provides diverse functions handling such kind data 
-to support statistical analysis.
+In the \pkg{yuima} package stochastic differential equations can be of very abstract type, 
+multidimensional, driven by Wiener process or fractional Brownian motion 
+with general Hurst parameter, with or without jumps specified as L\'evy noise.
+
+The \pkg{yuima} package is intended to offer the basic infrastructure on which complex models and inference procedures can be built on.
+This paper explains the design of the \pkg{yuima} package and provides some examples of applications.
+The paper is organised as follows.
+Section \ref{sec2} is an overview about the package.
+Section \ref{sec3} describe the way models are specified in \pkg{yuima}.
+Section \ref{sec4} explains asymptotic expansion methods.
+Section \ref{sec5} is a review of basic inference procedures.
+Finally, Section \ref{sec6} explains additional details and the roadmap of the YUIMA Project.
 
@@
 
-\section{The \pkg{yuima} package}
+\section{The \pkg{yuima} package}\label{sec2}
 The package \pkg{yuima} depends on some other packages, like  \pkg{zoo}, which can be installed separately.
@@
 
-\section{Model specification}
+\section{Model specification}\label{sec3}
 In order to show how general is the approach in the \pkg{yuima} package we present some examples.
@@
 
-\section{Miscellanea}
-The code \pkg{yuima} package offers several other utility and extensions from the main core classes. We will review some of these.
-
-\subsection{Asymptotic expansion}
+\section{Asymptotic expansion}\label{sec4}
 The \pkg{yuima} package can handle asymptotic expansion of functionals of  $d$-dimensional diffusion process 
@@
 
-\subsection{Export of a \code{yuima} model}
-The \pkg{yuima} implements the function \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}. A simple writing like
-\code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper.
-
-\section{Inference for stochastic processes}
+\section{Inference for stochastic processes}\label{sec5}
 The \pkg{yuima} implements several optimal techniques for parametric and nonparametric estimation of multidimensional stochastic differential equations.
@@
 and the two random measures are independent each other as well as 
-the stochastic processes. 
-Then it is known from today's lecture that
+the stochastic processes. It is known that 
 \begin{equation}
@@
 cce(Y)
-@ 
-Now we calculate the asymptotic variance of the estimator using \eqref{eq:vc}
-<<>>=
-# asymptotic variance
-var.c <- function(T, p1,p2, sigma1, sigma2, rho)
-{
-  tmp_integrand1 <- function(t) (sigma1(t) * sigma2(t))^2
-  i1 <- integrate(tmp_integrand1,0,T)
-  tmp_integrand2 <- function(t) (sigma1(t) * sigma2(t) * rho(t))^2
-  i2 <- integrate(tmp_integrand2,0,T)
-  2*(1/p1 + 1/p2)* i1$value + 2*(1/p1+1/p2 - 1/(p1+p2)) * i2$value
-}
-vc <- var.c(T=Terminal, p1, p2, diff.coef.1, diff.coef.2, cor.rho)
-sqrt(vc/n)
 @
@@
  \end{array}\right]
-'
 \left(\begin{array}{c}
@@
 
-\section{The future of YUIMA}
-some words
+\section{Miscellanea and Roadmap of YUIMA Project}\label{sec6}
+Other statistical techniques are already implemented or will be shortly releases in the \pkg{yuima}.
+For example, a nice utility is the function \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}.
+A simple writing like
+\code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper.
 
 \section*{Acknowledgements}
-some thanks
+The YUIMA Project was initially started by Nakahiro Yoshida as principal 
+investigator and then Stefano M. Iacus joined as coordinator of the \proglang{R}
+implementation. \textcolor{red}{PLEASE ADD SOME HISTORICAL FACTS AND ACKNOWLEDGMENTS OF DIFFERENT FUNDS}.
 
@@
 \end{document}
+
+The YUIMA Project\footnote{The Project has been  funded up to 2010 by the Japan Science Technology (JST) Basic Research Programs PRESTO, Grants-in-Aid for Scientific Research No.
+19340021.} is an open source\footnote{All code in the \pkg{yuima} package is subject to the GNU General Public License, Version 2, see \url{http://www.gnu.org/licenses/gpl-2.0.html}.}
+ academic project aimed at developing the \proglang{R} package named ``\pkg{yuima}'' for simulation and inference of stochastic differential equations. 
+The YUIMA Project is mainly developed by mathematicians and 
+statisticians who actively publish in the field of inference and simulation for stochastic
+differential equations.
+%The YUIMA Project Core Team, currently 
+%consists of the following people: A. Brouste, M. Fukasawa, H. Hino, S.M. Iacus, K. Kamatani, H.Masuda, Y. Shimizu, M. Uchida, N. Yoshida.
+
+ The \pkg{yuima} package  provides 
+an object-oriented programming environment 
+for simulation and statistical inference 
+for stochastic processes by \proglang{R}.
+The \pkg{yuima} package adopts the  $\tt S4$ system of classes and methods \citep{chambers98}.
+
+Under this framework, 
+the \pkg{yuima} package also supplies various functions 
+to execute simulation and statistical analysis. 
+Both categories of procedures may depend each other. 
+Statistical inference often requires a simulation technique 
+as a subroutine, and a certain simulation method 
+needs to fix a tuning parameter by applying 
+a statistical methodology. 
+It is especially the case of stochastic processes 
+because most of expected values involved 
+do not admit an explicit expression. 
+The \pkg{yuima} package facilitates comprehensive, systematic approaches 
+to the solution.
+
+Stochastic differential equations are 
+commonly used 
+to model random evolution along continuous or 
+practically continuous time, such as 
+the random movements of a stock price. 
+Theory of statistical inference for 
+stochastic differential equations already has a fairly long history, 
+more than three decades, but it is still developing quickly new 
+methodologies and expanding the area. 
+The formulas produced by the theory are usually very sophisticated, 
+which makes it difficult for practitioners not necessarily 
+familiar with this field to enjoy their utility. 
+For example, the asymptotic expansion method for computing 
+asian option prices (i.e., expectation of a functional of 
+a stochastic process) provides precise approximation values 
+instantaneously, taking advantage of the analytic approach, 
+but the formula, based on Malliavin calculus, has a long expression like more than one page!
+
+The \pkg{yuima} package delivers up-to-date methods as a package 
+onto the desk of the user working 
+with simulation and/or statistics for stochastic differential equations.
+
+Sampled data from a continuous-time process features 
+the time stamps as well as the positions of the object. 
+It is requiring a new theory of estimation. 
+The \pkg{yuima} framework can apply multi-dimensional time stamps 
+of tick data and provides diverse functions handling such kind data 
+to support statistical analysis.

### 2011-09-16 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update cpoint and cce sections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@185 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
         Ryosuke Nomura\\University of Tokyo\AND
-        Yusutaka Shimuzu\\University of Osaka \And
+        Yasutaka Shimuzu\\University of Osaka \And
         Masayuki Uchida\\University of Osaka \And
@@
 and the two random measures are independent each other as well as 
-the stochastic processes. It is known that 
+the stochastic processes. Under this scheme data become asynchronous. It is known that 
 \begin{equation}
@@
 @
-and we now apply random sampling
+<<cceplot1,fig=TRUE,width=9,height=5>>=
+plot(X,main="complete data")
+@
+
+We now apply random sampling in the following way: we define a new sampling structure via \code{setSampling} specifying in the argument \code{random} 
+a list which contains a vector of random distributions.
+For the $i$-th component of $X$ we specificy an exponential distribution with rate $n \cdot p_i/T$ for the random times.
+This will generate Poisson random times with the corresponding rate.
 <<>>=
@@
   function(x) rexp(x, rate=p1*n/Terminal))) )
+@
+Now we use the \code{subsampling} function to subsample the original data \code{X} into new asynchronous data \code{Y}
+<<>>=
 Y <- subsampling(X, sampling=newsamp)
 cce(Y)
+@ 
+<<cceplot2,fig=TRUE,width=9,height=5>>=
+plot(Y,main="asynchronous data")
 @
@@
 \begin{array}{ll}
-Y_0+\int_0^t b_s d s+\int_0^t \sigma(X_s,\theta_1^*) d W_s
+Y_0+\int_0^t b_s d s+\int_0^t \sigma(X_s,\theta_0^*) d W_s
 & \mbox{ for } t\in[0,\tau^*)
 \\
-Y_{\tau^*}+\int_{\tau^*}^t b_s d s+\int_{\tau^*}^t \sigma(X_s,\theta_2^*) d W_s
+Y_{\tau^*}+\int_{\tau^*}^t b_s d s+\int_{\tau^*}^t \sigma(X_s,\theta_1^*) d W_s
 & \mbox{ for } t\in[\tau^*,T]. 
@@
 The change point $\tau^*$ instant is unknown and 
-is  to be estimated, along with $\theta_1^*$ and $\theta_2^*$, from the observations sampled 
+is  to be estimated, along with $\theta_0^*$ and $\theta_1^*$, from the observations sampled 
 from the path of $(X,Y)$. The \pkg{yuima} implements the quasi-maximum likelihood approach as in \citet{iacyos09} described in the following.
@@
 \begin{equation} 
-\Phi_n(t;\theta_1,\theta_2)
+\Phi_n(t;\theta_0,\theta_1)
 =
-\sum_{i=1}^{[nt{/T}]}G_i(\theta_1)+\sum_{i={[nt{/T}]+1}}^nG_i(\theta_2),
+\sum_{i=1}^{[nt{/T}]}G_i(\theta_0)+\sum_{i={[nt{/T}]+1}}^nG_i(\theta_1),
 \label{eq:phiCH10}
@@
 Suppose that there exists an estimator 
-$\hat{\theta}_k$ for each $\theta_k$, $k=1,2$. 
+$\hat{\theta}_k$ for each $\theta_k$, $k=0,1$. 
 In case $\theta_k^*$ are known, we define $\hat{\theta}_k$ 
@@
 \hat{\tau}_n&=&\arg\!\!\min\limits_{t\in[0,T]} 
-\Phi_n(t;\hat{\theta}_1,\hat{\theta}_2).
+\Phi_n(t;\hat{\theta}_0,\hat{\theta}_1).
 \end{eqnarray*}
@@
 \subsection{Example  of Volatility Change-Point Estimation}
-Consider the 2-dimensional stochastic differential equation
+One example of model that can be analyzed by this technique is, for example, the 2-dimensional stochastic differential equation
 $$
@@
 \left(\begin{array}{c}
- 1 - X^1_t \\ 
+b_1(X^1_t) \\ 
+b_2(X^2_t) \\ 
+\end{array}\right) \de t
++
+\left[\begin{array}{cc}
+\theta_{1.k}  \cdot  X_t^1&0  \cdot  X_t^1 \\ 
+0  \cdot  X_t^2&\theta_{2.k}  \cdot  X_t^2 \\ 
+\end{array}\right]
+\left(\begin{array}{c}
+\de W_t^1\\ \de W_t^2 
+\end{array}\right)
+$$
+where $b_1(\cdot)$ and $b_2(\cdot)$ are any functions and $\theta_{1.k}$ and $\theta_{2.k}$ the value of the parameters before ($k=0$) and after $k=1$) the change point. 
+Just for simplicity and in order to simulate some data, we specify some specific form for $b_1(\cdot)$ and $b_2(\cdot)$ but this information will not be used in the change point analysis.
+For example, we will simulate
+the following 2-dimensional stochastic differential equation
+$$
+\left(\begin{array}{c}
+ \de X_t^1\\ \de X_t^2 
+ \end{array}\right)
+ = 
+\left(\begin{array}{c}
+ \sin(X^1_t) \\ 
  3 - X^2_t \\ 
@@
 \left[\begin{array}{cc}
- \theta_{1.1}  \cdot  X_t^1&0  \cdot  X_t^1 \\ 
- 0  \cdot  X_t^2&\theta_{1.2}  \cdot  X_t^2 \\ 
+ \theta_{1.k}  \cdot  X_t^1&0  \cdot  X_t^1 \\ 
+ 0  \cdot  X_t^2&\theta_{2.k}  \cdot  X_t^2 \\ 
  \end{array}\right]
@@
 $$
-with change point instant at time $\tau=0.4$.  Some code is needed to simulate such a process. First we define the model
+with change point instant at time $\tau=0.4$.   First, we describe the model to be simulated
 <<cpoint1,eval=TRUE, echo=TRUE,results=hide>>=
-diff.matrix <- matrix(c("theta1.1*x1","0*x2","0*x1","theta1.2*x2"), 2, 2)
-drift.c <- c("1-x1", "3-x2")
+diff.matrix <- matrix(c("theta1.k*x1","0*x2","0*x1","theta2.k*x2"), 2, 2)
+drift.c <- c("sin(x1)", "3-x2")
 drift.matrix <- matrix(drift.c, 2, 1)
@@
 @
-and then simulate two trajectories. One up to the change point $\tau=4$ with parameters $\theta_{1.1}=0.1$ and  $\theta_{1.1}=0.2$, and a second trajectory with parameters
- $\theta_{1.1}=0.6$ and  $\theta_{1.2}=0.6$. For the second trajectory, the initial value is set to the last value of the first trajectory.
+and then simulate two trajectories. One up to the change point $\tau=4$ with parameters $\theta_{1.0}=0.1$ and  $\theta_{2.0}=0.2$
 <<cpoint3,results=hide>>=
@@
 
-t1 <- list(theta1.1=.1, theta1.2=0.2)
-t2 <- list(theta1.1=.6, theta1.2=.6)
+t0 <- list(theta1.k=0.1, theta2.k=0.2)
 
@@
 yuima1 <- setYuima(model=ymodel, sampling=ysamp1)
-yuima1 <- simulate(yuima1, xinit=c(1, 1), true.parameter=t1)
+yuima1 <- simulate(yuima1, xinit=c(1, 1), true.parameter=t0)
 
@@
 x2 <- as.numeric(x2[length(x2)])
-
+@
+now we generate the second trajectory
+ with parameters
+$\theta_{1.1}=0.6$ and  $\theta_{2.1}=0.6$.
+For this trajectory, the initial value is set to the last value of the first trajectory stored in \code{x1} and \code{x2} for the two component of the process.
+<<cpoint3b,results=hide>>=
+t1 <- list(theta1.k=0.6, theta2.k=0.6)
 ysamp2 <- setSampling(Initial=n*tau*0.01, n=n*(1-tau), delta=0.01)
 yuima2 <- setYuima(model=ymodel, sampling=ysamp2)
-yuima2 <- simulate(yuima2, xinit=c(x1, x2), true.parameter=t2)
+yuima2 <- simulate(yuima2, xinit=c(x1, x2), true.parameter=t1)
+@
+finaly we collate the two trajectories
+<<cpoint3c,results=hide>>=
 yuima <- yuima1
@@
 @
+
 The composed trajectory appears as follows
@@
 
-\noindent
-Just as an example, we test the ability of the change point estimator to identify $\tau$ when for given true values of the parameters $\theta_{1.1}$ and $\theta_{1.2}$
+As said, the change point analysis do not consider the information coming from the drift part of the model and \pkg{yuima} ignores this internally.
+Just to make clear that the information on the drift term is not considered by the function \code{CPoint}, we redefine the \code{yuima} model removing the information coming from the drift and then adding back the data.
+<<cpoint4b>>=
+noDriftModel <- setModel(drift=c("0", "0"), diffusion=diff.matrix, time.variable="t",
+state.variable=c("x1", "x2"), solve.variable=c("x1", "x2"))
+noDriftModel <- setYuima(noDriftModel, data=yuima@data)
+noDriftModel@model@drift
+@
+First we show that there is no difference in using the complete model or the model without drift. For simplicity, we assume the true values of the parameters for $\theta_{1.k}$ and $\theta_{1.k}$
 <<cpoint5>>=
-t.est <- CPoint(yuima,param1=t1,param2=t2, plot=TRUE)
+t.est <- CPoint(yuima,param1=t0,param2=t1, plot=TRUE)
 t.est$tau
-@
-A two stage change point estimation approach is available as explained in  \citet{iacyos09}.
+t.est2 <- CPoint(noDriftModel,param1=t0,param2=t1, plot=TRUE)
+t.est2$tau
+@
+Now we proceed first with the estimation of the parameters before and after the change point. The \pkg{yuima} package contains two  functions
+which are useful in the framework of change point or sequential analysis.
+The function \code{qmleL} estimates a model by quasi maximum likelihood using observations in the time interval $[0,t]$ where $t$ cam be specificed by the user.
+In our example, we set \code{t=0.2}.
+Similary for \code{qmleR}, which uses only observations in the time interval $[t, T]$.
+In our example, we take \code{t=0.8}.
+<<>>=
+qmleL(noDriftModel, t=0.2, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estL
+qmleR(noDriftModel, t=0.8, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estR
+t0.est <- coef(estL)
+t1.est <- coef(estR)
+@
+and now we proceed with change point estimation
+<<>>=
+t.est3 <- CPoint(noDriftModel,param1=t0.est,param2=t1.est, plot=TRUE)
+t.est3
+@
+Notice that, even if the estimated parameters are not too accurate because we use a small subsets of observations, the change point estimate remains good.
+A two stage change point estimation approach is also possible as explained in  \citet{iacyos09}.
 

### 2011-09-16 pkg/yuimadocs/inst/doc/JSS/article.Rnw
subsampling
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@186 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\subsection{Simulation of models and sampling}
-The \code{simulate} function simulates \code{yuima} models according to Euler-Maruyama scheme in the presence of non-fractional diffusion noise and L\'eve jumps and uses the Cholesky or the Wood and Chan methods for the fractional Gaussian noise.
+\subsection{Simulation, sampling and subsampling}
+The \code{simulate} function simulates \code{yuima} models according to Euler-Maruyama scheme in the presence of non-fractional diffusion noise and L\'evy jumps and uses the Cholesky or the Wood and Chan methods for the fractional Gaussian noise.
 The \code{simulate} function accepts several arguments including the description of the sampling structure, which is an object of type \code{yuima.sampling}.
@@
 This is convenient if one wants to simulate data at very high frequency but then return only low frequency data for inference or other applications.
+We now go through few examples just to describe the use of standard arguments to these functions but the reader is invited to go thorough the man pages of the \code{yuima} packages for complete details.
+
+Assume that we want to simulate this model
+$$
+\begin{aligned}
+\de X^1_t &= -\theta X^1_t \de t + \de W^1_t + X^2_t \de W^3_t\\
+\de X^2_t &= -(X^1_t + \gamma X^2_t) \de t + X^1_t \de W^1_t + \delta \de W^2_t
+\end{aligned}
+$$
+Now we prepare the model using the \code{setModel} constructor function
+<<setModel1>>=
+sol <- c("x1","x2") # variable for numerical solution
+b <- c("-theta*x1","-x1-gamma*x2")   # drift vector 
+s <- matrix(c("1","x1","0","delta","x2","0"),2,3)  #  diffusion matrix
+mymod <- setModel(drift = b, diffusion = s, solve.variable = sol)
+@
+Suppose now that we want to simulate the process on a regular grid on the interval $[0,3]$ and $n=3000$ observations. We can prepare the sampling structure as follows
+<<>>=
+samp <- setSampling(Terminal=3, n=3000)
+@
+and look inside it
+<<>>=
+str(samp)
+@
+As seen from the output, the sampling structure is quite reach and we will show how to specify few of the slots later one.
+We now simulate this process specifying the \code{sampling} argument to \code{simulate}
+<<>>=
+set.seed(123)
+X2 <- simulate(mymod, sampling=samp)
+@
+now the sampling structure is recorded along with the data in the \code{yuima} object \code{X2}
+<<>>=
+str(X2@sampling)
+@
+\subsubsection{Subsampling}
+The sampling structure can be used to operate subsampling. Next example shows how to perform Poisson random sampling, with two independent Poisson processes, one per coordinate of \code{X2}.
+<<sub1>>=
+newsamp <- setSampling(
+random=list(rdist=c( function(x) rexp(x, rate=10), 
+function(x) rexp(x, rate=20))) )
+str(newsamp)
+@
+In the above we have specified two independent exponential  distributions to represent Poissoninan random times. Notice that the slot \code{regular} is now set to \code{FALSE}.
+Now we subsample the original trajectory of \code{X2} using the \code{subsampling} function
+<<sub2,fig=TRUE,width=9,height=5>>=
+newdata <- subsampling(X2, sampling=newsamp)
+plot(X2,plot.type="single", lty=c(1,3),ylab="X2")
+points(get.zoo.data(newdata)[[1]],col="red")
+points(get.zoo.data(newdata)[[2]],col="green",pch=18)
+@
+
+Or we can operate a deterministic sampling specifying two different regular frequencies
+<<sub3,fig=TRUE,width=9,height=5>>=
+newsamp <- setSampling(delta=c(0.1,0.2))
+newdata <- subsampling(X2, sampling=newsamp)
+plot(X2,plot.type="single", lty=c(1,3),ylab="X2")
+points(get.zoo.data(newdata)[[1]],col="red")
+points(get.zoo.data(newdata)[[2]],col="green", pch=18)
+@
+
+Again one can look at the structure of the sampling structure.
+
+Subsampling can be used within the \code{simulate} function.
+What is usually done in simulation studies, is to simulate the process at very high frequency but then use data for estimation at a lower frequency.
+This can be done in a single step in the following way.
+<<sub4,fig=TRUE,width=9,height=5>>=
+set.seed(123)
+Y.sub <- simulate(mymod,sampling=setSampling(delta=0.001), subsampling=setSampling(delta=0.01))
+set.seed(123)
+Y <- simulate(mymod, sampling=setSampling(delta=0.001))
+plot(Y, plot.type="single")
+points(get.zoo.data(Y.sub)[[1]],col="red")
+points(get.zoo.data(Y.sub)[[2]],col="green",pch=18)
+@
+
+In the previous code we have simulated the process twice just to show the effect of the subsampling but the reader should use only the line which outputs the simulation to \code{Y.sub} as seen in the next plot.
+<<sub5,fig=TRUE,width=9,height=5>>=
+plot(Y.sub, plot.type="single")
+@
 

### 2011-09-16 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update cpoint example and fixed subsample example
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@187 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 set.seed(123)
-Y.sub <- simulate(mymod,sampling=setSampling(delta=0.001), subsampling=setSampling(delta=0.01))
+Y.sub <- simulate(mymod,sampling=setSampling(delta=0.001,n=1000), subsampling=setSampling(delta=0.01,n=100))
 set.seed(123)
-Y <- simulate(mymod, sampling=setSampling(delta=0.001))
+Y <- simulate(mymod, sampling=setSampling(delta=0.001,n=1000))
 plot(Y, plot.type="single")
@@
 
-t0 <- list(theta1.k=0.1, theta2.k=0.2)
+t0 <- list(theta1.k=0.5, theta2.k=0.3)
 
@@
 yuima1 <- setYuima(model=ymodel, sampling=ysamp1)
-yuima1 <- simulate(yuima1, xinit=c(1, 1), true.parameter=t0)
+yuima1 <- simulate(yuima1, xinit=c(3, 3), true.parameter=t0)
 
@@
 <<cpoint3b,results=hide>>=
-t1 <- list(theta1.k=0.6, theta2.k=0.6)
+t1 <- list(theta1.k=0.2, theta2.k=0.4)
 ysamp2 <- setSampling(Initial=n*tau*0.01, n=n*(1-tau), delta=0.01)
@@
 <<cpoint5>>=
-t.est <- CPoint(yuima,param1=t0,param2=t1, plot=TRUE)
+t.est <- CPoint(yuima,param1=t0,param2=t1)
 t.est$tau
-t.est2 <- CPoint(noDriftModel,param1=t0,param2=t1, plot=TRUE)
+t.est2 <- CPoint(noDriftModel,param1=t0,param2=t1)
 t.est2$tau
@@
 <<>>=
-qmleL(noDriftModel, t=0.2, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estL
-qmleR(noDriftModel, t=0.8, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estR
+qmleL(noDriftModel, t=2, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estL
+qmleR(noDriftModel, t=8, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estR
 t0.est <- coef(estL)
@@
 <<>>=
-t.est3 <- CPoint(noDriftModel,param1=t0.est,param2=t1.est, plot=TRUE)
+t.est3 <- CPoint(noDriftModel,param1=t0.est,param2=t1.est)
 t.est3

### 2011-09-27 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@189 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-In the \pkg{yuima} package stochastic differential equations can be of very abstract type, multidimensional, driven by Wiener process of fractional Brownian motion with general Hurst parameter, with or without jumps specified as L\'evy noise.
+In the \pkg{yuima} package stochastic differential equations can be of very abstract type, multidimensional, driven by Wiener process or fractional Brownian motion with general Hurst parameter, with or without jumps specified as L\'evy noise.
 
@@
 instantaneously, taking advantage of the analytic approach, 
-but the formula has a long expression like more than one page!
+but the formula has a long expression like more than 900 terms!
 
@@
 
-The \code{yuima.model} class contains informations about the stochastic differential equation of interest. The constructor function \code{setModel} is
+The \code{yuima.model} class contains information about the stochastic differential equation of interest. The constructor function \code{setModel} is
 used to give a mathematical description of the stochastic differential equation. 
@@
   in the \code{parameter} slot.
-\item \code{jump.variable} the name of the variable used in the description of the L\'eve component;
+\item \code{jump.variable} the name of the variable used in the description of the L\'evy component;
 \item \code{solve.variable} contains a vector of variable names, each element corresponds to the
@@
 \end{itemize}
-As seen in the above, the parameters space is accurately described internally in a \code{yuima} object because in inference for stochastic differential equations, estimators of different parameters have different properties.
+As seen in the above, the parameter space is accurately described internally in a \code{yuima} object because in inference for stochastic differential equations, estimators of different parameters have different properties.
 Usually, the rate of convergence for estimators in the diffusion coefficient are similar to the ones in the i.i.d.
@@
 The functional is expressed in the following abstract form
-$$
-F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve), \qquad W_t^0=t
-$$ 
+\begin{equation}
+\label{yuima:func1}
+F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve), \qquad W_t^0=t.
+\end{equation} 
 A typical example of application is the case of Asian option pricing. For example, in the Black \& Scholes model
-$$\de X_t^\ve = \mu X_t^\ve\de t + \ve X_t^\ve \de W_t$$ 
+\begin{equation}
+\label{eqyuima:gbm1}
+\de X_t^\ve = \mu X_t^\ve\de t + \ve X_t^\ve \de W_t
+\end{equation}
 the price of the option is of the form
@@
 $$
-in 
-$$
-F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve)
-$$ 
+in \eqref{yuima:func1}.
 So, the call option price requires the composition  of a smooth functional
@@
 calculation of the price, while asymptotic expansion of $F^\ve$ provides very accurate approximations.
-The \pkg{yuima} package provides functions to construct the functional $F^\ve$, and automatic asymptotic expansion based on Malliavin calculus starting from a \pkg{yuima} object.
-Next is an example
+The \pkg{yuima} package provides functions to construct the functional $F^\ve$, and automatic asymptotic expansion based on Malliavin calculus \citep{Yoshida92b} starting from a \pkg{yuima} object.
+Here is an example. Consider a simple geometric Brownian motion of equation \eqref{eqyuima:gbm1} with $\mu=1$ and $X_0=1$. We define the model and the functional below:
 <<echo=TRUE, print=FALSE, results=hide>>=
-diff.matrix <- matrix( c("x*e"), 1,1)
-model <- setModel(drift = c("x"), diffusion = diff.matrix)
+model <- setModel(drift = "x", diffusion = matrix( "x*e", 1,1))
 T <- 1
@@
 F <- 0
-e <- .3
+e <- 0.5
 yuima <- setYuima(model = model, sampling = setSampling(Terminal=T, n=1000))
@@
 @
-this time the \code{setFunctional} command fills the appropriate slots
+this time the \code{setFunctional} command fills the appropriate slots inside the \code{yuima} object
 <<echo=TRUE, print=FALSE>>=
@@
 @
-Then, it is as easy as
+Then, to obtain the first term in the asymptotic expansion (i.e. for $\ve=0$), it is as easy as calling the function \code{F0} on the \code{yuima} object:
 <<echo=TRUE, print=FALSE>>=
@@
 F0
+@
+so the option price approximation is
+<<>>=
 max(F0-K,0)  # asian call option price
 @
-to obtain the zero order approximation of the value of the functional.
 We can go up to the first order approximation adding one term to the expansion
@@
 rho <- expression(0)
-get_ge <- function(x,epsilon,K,F0){
-  tmp <- (F0 - K) + (epsilon * x) 
-  tmp[(epsilon * x) < (K-F0)] <- 0
-  return( tmp )
-}
 epsilon <- e  # noise level
@@
 \section{Inference for stochastic processes}\label{sec5}
-The \pkg{yuima} implements several optimal techniques for parametric and nonparametric estimation of multidimensional stochastic differential equations.
+The \pkg{yuima} implements several optimal techniques for parametric, semi- and non-parametric estimation of (multidimensional) stochastic differential equations.
 Although most of the examples in this section are given on simulated data, the main way to fill up the \code{data} slot of a \code{yuima} object is to use the function \code{setYuima}.
@@
 where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, $A^{\otimes 2}= A^T A$ and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
-Then, \cite{Yoshida92}, the QML estimator of $\theta$ is
+Then, \citep[see e.g.][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
 $$\tilde\theta_n=\arg\min_\theta \ell_n({\bf X}_n,\theta)$$
@@
 \end{equation}
-with $\theta_1=0.3$ and $\theta_2=0.2$
+with $\theta_1=0.2$ and $\theta_2=0.3$. We generate sampled data $X_{t_i}$, with $t_i = i \cdot n^{-\frac23}$.
 <<echo=TRUE, print=FALSE,results=hide>>=
 ymodel <- setModel(drift="(2-theta2*x)", diffusion="(1+x^2)^theta1")
-n <- 500
+n <- 750
 ysamp <- setSampling(Terminal = n^(1/3), n = n)
@@
 @
-With the simulated path we can use the function \code{qmle} to estimate the parameters as follows
+With the sampled data we can use the function \code{qmle} to estimate the parameters as follows
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 @
-and the estimated coefficients are as follows
+and the estimated coefficients are extracted from the output object \code{mle1} as follows
 <<echo=TRUE, print=FALSE>>=
@@
 Adaptive Bayes estimation is developed in \pkg{yuima} via the method $\tt adaBayes$.
-Consider again the model \eqref{eq:model1} with the same values for the parameters, i.e.
- $\theta_1=0.3$ and $\theta_2=0.1$
+Consider again the model \eqref{eq:model1} with the same values for the parameters.
 In order to perform Bayesian estimation, we need to prepare the prior densities for the parameters.
@@
 @
-Then we call $\tt adaBayes$ as follows
+Then we call $\tt adaBayes$, on the same sample data we used for the \code{qmle} function, as follows
 <<eval=TRUE,echo=TRUE, results=hide>>=
-bayes1 <- adaBayes(yuima, start=param.init, prior=prior, method="nomcmc")
+bayes1 <- adaBayes(yuima, start=param.init, prior=prior)
 @
@@
 @
-The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical integration, otherwise MCMC method is used.
+%The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical
+%integration, otherwise MCMC method is used.
 
@@
 
+\subsubsection{Asynchronous estimation for nonlinear systems}
+Consider now the two-dimensional system with nonlinear feedback
+$$
+\begin{aligned}
+\de X_t &= Y_t\de t + \sigma_1(t,X_t,Y_t) \de W_t\\
+\de Y_t &= -X_t\de t + \rho(t,X_t,Y_t)\sigma_2(t,X_t,Y_t) \de W_t+
+\sigma_2(t,X_t,Y_t)\sqrt{1-\rho^2(t,X_t,Y_t)}\de B_t
+\end{aligned}
+$$
+with $ \sigma_1(t,X_t,Y_t)=\sqrt{|X_t|(1+t)}$,
+$ \sigma_2(t,X_t,Y_t)=\sqrt{|Y_t|}$,
+$\rho(t,X_t,Y_t) = \frac{1}{1+X_t^2}$ and $W_t$, $B_t$, two independent
+Brownian motions.
+We construct the model and we generate data from it
+<<>>=
+b1 <- function(x,y) y
+b2 <- function(x,y) -x
+s1 <- function(t,x,y) sqrt(abs(x)*(1+t))
+s2 <- function(t,x,y) sqrt(abs(y))
+cor.rho <- function(t,x,y) 1/(1+x^2)
+diff.mat <- matrix(c("s1(t,x,y)", "s2(t,x,y) * cor.rho(t,x,y)","", "s2(t,x,y) * sqrt(1-cor.rho(t,x,y)^2)"), 2, 2) 
+cor.mod <- setModel(drift = c("b1","b2"), diffusion = diff.mat,solve.variable = c("x", "y"),state.var=c("x","y"))
+
+## Generate a path of the process
+set.seed(111) 
+Terminal <- 1
+n <- 10000
+yuima.samp <- setSampling(Terminal = Terminal, n = n) 
+yuima <- setYuima(model = cor.mod, sampling = yuima.samp) 
+yuima <- simulate(yuima, xinit=c(2,3)) 
+@
+We apply the same Poisson random sampling so that the object \code{Y} will contain  asynchronous data
+<<>>=
+p1 <- 0.2
+p2 <- 0.3
+newsamp <- setSampling(
+random=list(rdist=c( function(x) rexp(x, rate=p1*n/Terminal), 
+function(x) rexp(x, rate=p1*n/Terminal))) )
+Y <- subsampling(yuima, sampling = newsamp)
+@
+<<cceplot3,fig=TRUE,width=9,height=5>>=
+plot(Y,main="asynchronous data (non linear case)")
+@
+
+The estimated covariance for the complete trajectory \code{yuima}
+is now compared with the one obtained on asyncronous data \code{Y}
+<<>>=
+cce(yuima)
+cce(Y)
+@
+
 \subsection{Change point analysis}
@@
 the LASSO method selects the reduced model
-$$\de X_t =  0.6 \de t + 0.12 X_t^{\frac32}\de W_t$$
+$$\de X_t =  0.6 \de t + 0.12 X_t^{\frac32}\de W_t.$$
+Notice that this model is not an ergodic one, indicating that the LASSO method shows that the real data are indeed not stationary, but still in the family of CKLS models.
 

### 2011-09-29 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@191 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %integration, otherwise MCMC method is used.
+\subsubsection{Small sample size}
+It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
+In our example above, we took $T=n^(1/3)$, with $n = \Sexpr{n}$, which is approximatively  \Sexpr{round(n^(1/3),2)}.
+Now we reduce the sample size to $n=500$ and the value of $T$ is then $T=\Sexpr{round(500^(1/3),2)}$.
+We then apply both quasi-maximum likelihood and adaptive Bayes type estimators to these data
+<<>>=
+n <- 500
+ysamp <- setSampling(Terminal = n^(1/3), n = n)
+yuima <- setYuima(model = ymodel, sampling = ysamp)
+set.seed(123)
+yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.2, theta2 = 0.3))
+param.init <- list(theta2=0.5,theta1=0.5)
+mle2 <- qmle(yuima, start =param.init , 
+lower = list(theta1=0, theta2=0), 
+upper = list(theta1=1, theta2=1))
+bayes2 <- adaBayes(yuima, start=param.init, prior=prior)
+@
+and we look at the estimates
+<<>>=
+coef(bayes2)
+coef(mle2)
+@
+Compared to the results above, we see that the parameters in the diffusion coefficients are estimated with good quality while for the parameters in the drift the quality of estimation deteriorates.
+The adaptive Bayes estimator seems to perform a little better though.
 

### 2011-10-15 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@192 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@193 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \Plainauthor{Alexandre Brouste, Stefano M. Iacus, Hiroki Masuda, Masayuki Uchida, Nakahiro Yoshida} %% comma-separated
-\Plaintitle{A Capitalized Title: Something About a Package foo} %% without formatting
-\Shorttitle{A Capitalized Title} %% a short title (if necessary)
+\Plaintitle{YUIMA: Simulation and Inference for SDE} %% without formatting
+\Shorttitle{The YUIMA Project} %% a short title (if necessary)
 
@@
 }
+The user can choose between the two simulation schemes, namely the Cholesky method and \citet{WoodChan} method.
 
@@
 
-\section*{Acknowledgements}
-The YUIMA Project was initially started by Nakahiro Yoshida as principal 
-investigator and then Stefano M. Iacus joined as coordinator of the \proglang{R}
-implementation. \textcolor{red}{PLEASE ADD SOME HISTORICAL FACTS AND ACKNOWLEDGMENTS OF DIFFERENT FUNDS}.
+%\section*{Acknowledgements}
+%The YUIMA Project was initially started by Nakahiro Yoshida as principal 
+%investigator and then Stefano M. Iacus joined as coordinator of the \proglang{R}
+%implementation. \textcolor{red}{PLEASE ADD SOME HISTORICAL FACTS AND ACKNOWLEDGMENTS OF DIFFERENT FUNDS}.
 

### 2011-10-18 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@194 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Alexandre Brouste\\University of Le Mans \And 
-        Masaaki Fukasawa\\University of Osaka\And
+        Masaaki Fukasawa\\Osaka University\And
         Hideitsu Hino\\Waseda University\And
         Stefano M. Iacus\\University of Milan \AND 	
-        Kengo Kamatani\\University of Tokyo \And
-        Yuta Koike\\University of Tokyo\And
-        Hiroki Masuda\\University of Kyushu \And
-        Ryosuke Nomura\\University of Tokyo\AND
-        Yasutaka Shimuzu\\University of Osaka \And
-        Masayuki Uchida\\University of Osaka \And
-        Nakahiro Yoshida\\University of Tokyo 
+        Kengo Kamatani\\The University of Tokyo \And
+        Yuta Koike\\The University of Tokyo\And
+        Hiroki Masuda\\Kyushu University \And
+        Ryosuke Nomura\\The University of Tokyo\AND
+        Yasutaka Shimuzu\\Osaka University \And
+        Masayuki Uchida\\Osaka University \And
+        Nakahiro Yoshida\\The University of Tokyo 
         }
@@
 \begin{aligned}
-\de X_t = & \,\,\, a(X_t)dt + b(X_t)\de W_t + \int\limits_{|z|>1}\!\!\! c(X_{t-},z)\mu(\de t,\de z) \\
+\de X_t = & \,\,\, a(X_t)\de t + b(X_t)\de W_t + \int\limits_{|z|>1}\!\!\! c(X_{t-},z)\mu(\de t,\de z) \\
 &{}+\!\! \int\limits_{0<|z|\le 1}\!\!\!  c(X_{t-},z)\{\mu(\de t,\de z) - \nu(\de z)\de t\}.
@@
 Then it is possible to consider the following SDE which involves jumps
-$$\de X_t =  a(X_t)dt + b(X_t)\de W_t + \de Z_t$$
+$$\de X_t =  a(X_t)\de t + b(X_t)\de W_t + \de Z_t$$
 In the next example we consider a compound Poisson process with intensity $\lambda=10$ with Gaussian jumps.
@@
 A simple Ornstein-Uhlembeck process with Gaussian jumps 
-$$\de X_t = -\theta X_t \de t + \sigma \de W_t + Z_t$$
+$$\de X_t = -\theta X_t \de t + \sigma \de W_t + \de Z_t$$
 is specified as
@@
 Without going into too much details, here is an example of specification of a simple Ornstein-Uhlembeck  process with IG (Inverse Gaussian) L\'evy measure
-$$\de X_t = -x dt + dZ_t$$
+$$\de X_t = -x \de t + \de Z_t$$
 <<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
@@
 The \pkg{yuima} package implements many multivariate Random Numbers Generators (RNG) which are needed to simulate L\'evy paths including
-\code{rIG} (Inverse Gaussian), \code{rNIG} (Normal Inverse Gaussian), \code{rbgamma} (Bilateral Gamma), \code{rngamma} (Gamma) and \texttt{rstable} (Stable Laws).
+\code{rIG} (Inverse Gaussian), \code{rNIG} (Normal Inverse Gaussian), \code{rbgamma} (Bilateral Gamma), \code{rngamma} (Normal Gamma) and \texttt{rstable} (Stable Laws).
 Other user-defined RNG can be used freely.
@@
 It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
-In our example above, we took $T=n^(1/3)$, with $n = \Sexpr{n}$, which is approximatively  \Sexpr{round(n^(1/3),2)}.
+In our example above, we took $T=n^\frac13$, with $n = \Sexpr{n}$, which is approximatively  \Sexpr{round(n^(1/3),2)}.
 Now we reduce the sample size to $n=500$ and the value of $T$ is then $T=\Sexpr{round(500^(1/3),2)}$.
@@
 \begin{array}{ll}
-Y_0+\int_0^t b_s d s+\int_0^t \sigma(X_s,\theta_0^*) d W_s
+Y_0+\int_0^t b_s \de s+\int_0^t \sigma(X_s,\theta_0^*) \de W_s
 & \mbox{ for } t\in[0,\tau^*)
 \\
-Y_{\tau^*}+\int_{\tau^*}^t b_s d s+\int_{\tau^*}^t \sigma(X_s,\theta_1^*) d W_s
+Y_{\tau^*}+\int_{\tau^*}^t b_s \de s+\int_{\tau^*}^t \sigma(X_s,\theta_1^*) \de W_s
 & \mbox{ for } t\in[\tau^*,T]. 

### 2011-11-11 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@195 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{amsmath,amssymb}
+
+\usepackage[utf8x]{inputenc}
 
@@
 options(prompt="R> ")
-options(width=70)
+options(width=60)
 @
@@
 \section{Introduction}
-The plan of the YUIMA Project is to define the bases for a complete environment for simulation and inference for stochastic processes via an \proglang{R} package called \pkg{yuima}.
+The plan of the YUIMA Project is to define the bases for a complete environment for simulation and inference for stochastic processes via an \proglang{R} \citep{ERRE} package called \pkg{yuima}.
  The package \pkg{yuima} provides 
@@
 for stochastic processes by \proglang{R}.
-The \pkg{yuima} package adopts the  \texttt{S4} system of classes and methods \citep{chambers98}.
+The \pkg{yuima} package adopts the  S4 system of classes and methods \citep{chambers98}.
 
@@
 
-\section{The \pkg{yuima} package}\label{sec2}
-The package \pkg{yuima} depends on some other packages, like  \pkg{zoo}, which can be installed separately.
+\section{The \pkg{yuima} Package}\label{sec2}
+The package \pkg{yuima} depends on some other packages, like  \pkg{zoo} \citep{zoo}, which can be installed separately.
 The package \pkg{zoo} is used internally to store time series data. This dependence may change in the future adopting a more flexible class for internal storage of time series.
 
-\subsection{How to obtain the package}
-The \pkg{yuima} package is hosted on \textsf{R-Forge} and the web page
+\subsection{How to Obtain the Package}
+The \pkg{yuima} package is hosted on R-Forge and the web page
 of the Project is  \url{http://r-forge.r-project.org/projects/yuima}.
-The  \textsf{R-Forge} page contains the latest development version, and stable
-version of the package will also be available through \textsf{CRAN}.
+The  R-Forge page contains the latest development version, and stable
+version of the package will also be available through CRAN.
 Development versions of the package are not supposed to be stable or functional, thus
 the occasional user should consider to install the stable version first.
-The package can be installed from \textsf{R-Forge} using
+The package can be installed from R-Forge using
 \code{install.packages("yuima",repos="http://R-Forge.R-project.org")}.
 
-\subsection{The main object and classes}
+\subsection{The Main Object and Classes}
 Before discussing the methods for simulation and inference for stochastic processes solutions to stochastic differential equations, here we discuss the main classes in the package.
@@
 
-\subsection{The $\tt yuima.model$ class}
+\subsection{The \code{yuima.model} Class}
 At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified. All multidimensional and eventually as parametric models.
@@
 \item \code{xinit} initial value of the stochastic differential equation;
-\item \code{equation.number} represents the number of equations, i.e. the number of one dimensional
+\item \code{equation.number} represents the number of equations, i.e., the number of one dimensional
   stochastic differential equations.
@@
 
-\section{Model specification}\label{sec3}
+\section{Model Specification}\label{sec3}
 In order to show how general is the approach in the \pkg{yuima} package we present some examples.
 
-\subsection{Diffusion processes}
+\subsection{Diffusion Processes}
 Assume that we want to describe the following stochastic differential equation
@@
 At this point, the package fills the proper slots of the \code{yuima} object
-{\scriptsize
 <<>>=
@@
 @
-}
 
@@
 The \code{simulate} function fills in addition the two slots \code{data} and \code{sampling} of the \code{yuima} object.
-{\scriptsize
 <<>>=
@@
 @
-}
-
-\subsection{Specification of parametric models}
+
+\subsection{Specification of Parametric Models}
 When a parametric model like
@@
 @ 
-{\scriptsize
 <<>>=
@@
 @
-}
 In order to simulate the parametric model it is necessary to specify the values of the parameters as the next code shows
@@
 
-\subsection{Multidimensional processes}
+\subsection{Multidimensional Processes}
 Next is an example with two stochastic differential equations driven by three independent Brownian motions
@@
 
-\subsubsection{Fractional Gaussian noise}
+\subsubsection{Fractional Gaussian Noise}
 In order to specify a stochastic differential equation driven by fractional Gaussian noise it is necessary to specify the value of the Hurst parameter.
@@
 @
-}
 The user can choose between the two simulation schemes, namely the Cholesky method and \citet{WoodChan} method.
 
-\subsection{L\'evy processes}
+\subsection{L\'evy Processes}
 Jump processes can be specified in different ways in mathematics and hence in \pkg{yuima} package. 
-Let $Z_t$ be a  Compound Poisson Process (i.e. jumps size follow some distribution, like the Gaussian law and jumps occur at Poisson times).
+Let $Z_t$ be a  Compound Poisson Process (i.e., jumps size follow some distribution, like the Gaussian law and jumps occur at Poisson times).
 Then it is possible to consider the following SDE which involves jumps
@@
 This model can be specified in \code{setModel} using the argument  \code{measure.type="CP"} 
-A simple Ornstein-Uhlembeck process with Gaussian jumps 
+A simple Ornstein-Uhlenbeck process with Gaussian jumps 
 $$\de X_t = -\theta X_t \de t + \sigma \de W_t + \de Z_t$$
@@
 Another possibility is to specify the jump component via the L\'evy measure.
-Without going into too much details, here is an example of specification of a simple Ornstein-Uhlembeck  process with IG (Inverse Gaussian) L\'evy measure
+Without going into too much details, here is an example of specification of a simple Ornstein-Uhlenbeck  process with IG (Inverse Gaussian) L\'evy measure
 $$\de X_t = -x \de t + \de Z_t$$
@@
 
-\subsection{Specification of generic models}
+\subsection{Specification of Generic Models}
 In general, the \pkg{yuima} package allows to specify a large family of models solutions to
@@
 
-\subsection{Simulation, sampling and subsampling}
+\subsection{Simulation, Sampling and Subsampling}
 The \code{simulate} function simulates \code{yuima} models according to Euler-Maruyama scheme in the presence of non-fractional diffusion noise and L\'evy jumps and uses the Cholesky or the Wood and Chan methods for the fractional Gaussian noise.
@@
 
-\section{Asymptotic expansion}\label{sec4}
+\section{Asymptotic Expansion}\label{sec4}
 The \pkg{yuima} package can handle asymptotic expansion of functionals of  $d$-dimensional diffusion process 
 $$\de X_t^\ve = a(X_t^\ve,\ve)\de t + b(X_t^\ve,\ve)\de W_t, \qquad \ve \in(0,1]$$
-with $W_t$ and $r$-dimensional Wiener process, i.e. $W_t=(W_t^1, \ldots, W_t^r)$.
+with $W_t$ and $r$-dimensional Wiener process, i.e., $W_t=(W_t^1, \ldots, W_t^r)$.
 The functional is expressed in the following abstract form
@@
 @
-Then, to obtain the first term in the asymptotic expansion (i.e. for $\ve=0$), it is as easy as calling the function \code{F0} on the \code{yuima} object:
+Then, to obtain the first term in the asymptotic expansion (i.e., for $\ve=0$), it is as easy as calling the function \code{F0} on the \code{yuima} object:
 <<echo=TRUE, print=FALSE>>=
@@
 
-\section{Inference for stochastic processes}\label{sec5}
+\section{Inference for Stochastic Processes}\label{sec5}
 The \pkg{yuima} implements several optimal techniques for parametric, semi- and non-parametric estimation of (multidimensional) stochastic differential equations.
@@
 
-\subsection{Quasi Maximum Likelihood estimation}
+\subsection{Quasi Maximum Likelihood Estimation}
 Consider the multidimensional diffusion process
@@
 where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, $A^{\otimes 2}= A^T A$ and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
-Then, \citep[see e.g.][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
+Then, \citep[see e.g.,][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
 $$\tilde\theta_n=\arg\min_\theta \ell_n({\bf X}_n,\theta)$$
@@
 
-\subsection{Adaptive Bayes estimation}
+\subsection{Adaptive Bayes Estimation}
 Consider again the  diffusion process solution to
@@
 %integration, otherwise MCMC method is used.
-\subsubsection{Small sample size}
+\subsubsection{Small Sample Size}
 It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
@@
 
-\subsection{Asynchronous covariance estimation}
+\subsection{Asynchronous Covariance Estimation}
 Suppose that two It\^o processes are observed
@@
 
-\subsubsection{Example: data generation and estimation by \pkg{yuima} package}
+\subsubsection{Example: Data Generation and Estimation by \pkg{yuima} Package}
 We will demonstrate how to apply cce function to 
@@
 
-\subsubsection{Asynchronous estimation for nonlinear systems}
+\subsubsection{Asynchronous Estimation for Nonlinear Systems}
 Consider now the two-dimensional system with nonlinear feedback
@@
 
-\subsection{Change point analysis}
+\subsection{Change-Point Analysis}
 Consider a multidimensional stochastic differential equation of the form
@@
 $$
-The change point $\tau^*$ instant is unknown and 
+The change-point $\tau^*$ instant is unknown and 
 is  to be estimated, along with $\theta_0^*$ and $\theta_1^*$, from the observations sampled 
@@
 just as $\hat{\theta}_k=\theta_k^*$. 
-The change point estimator  of $\tau^*$ is
+The change-point estimator  of $\tau^*$ is
 \begin{eqnarray*} 
@@
 $$
-where $b_1(\cdot)$ and $b_2(\cdot)$ are any functions and $\theta_{1.k}$ and $\theta_{2.k}$ the value of the parameters before ($k=0$) and after $k=1$) the change point. 
-Just for simplicity and in order to simulate some data, we specify some specific form for $b_1(\cdot)$ and $b_2(\cdot)$ but this information will not be used in the change point analysis.
+where $b_1(\cdot)$ and $b_2(\cdot)$ are any functions and $\theta_{1.k}$ and $\theta_{2.k}$ the value of the parameters before ($k=0$) and after $k=1$) the change-point. 
+Just for simplicity and in order to simulate some data, we specify some specific form for $b_1(\cdot)$ and $b_2(\cdot)$ but this information will not be used in the change-point analysis.
 For example, we will simulate
@@
 $$
-with change point instant at time $\tau=0.4$.   First, we describe the model to be simulated
+with change-point instant at time $\tau=0.4$.   First, we describe the model to be simulated
 <<cpoint1,eval=TRUE, echo=TRUE,results=hide>>=
@@
 @
-and then simulate two trajectories. One up to the change point $\tau=4$ with parameters $\theta_{1.0}=0.1$ and  $\theta_{2.0}=0.2$
+and then simulate two trajectories. One up to the change-point $\tau=4$ with parameters $\theta_{1.0}=0.1$ and  $\theta_{2.0}=0.2$
 <<cpoint3,results=hide>>=
@@
 
-As said, the change point analysis do not consider the information coming from the drift part of the model and \pkg{yuima} ignores this internally.
+As said, the change-point analysis do not consider the information coming from the drift part of the model and \pkg{yuima} ignores this internally.
 Just to make clear that the information on the drift term is not considered by the function \code{CPoint}, we redefine the \code{yuima} model removing the information coming from the drift and then adding back the data.
@@
 @
-Now we proceed first with the estimation of the parameters before and after the change point. The \pkg{yuima} package contains two  functions
-which are useful in the framework of change point or sequential analysis.
+Now we proceed first with the estimation of the parameters before and after the change-point. The \pkg{yuima} package contains two  functions
+which are useful in the framework of change-point or sequential analysis.
 The function \code{qmleL} estimates a model by quasi maximum likelihood using observations in the time interval $[0,t]$ where $t$ cam be specificed by the user.
@@
 @
-and now we proceed with change point estimation
+and now we proceed with change-point estimation
 <<>>=
@@
 @
-Notice that, even if the estimated parameters are not too accurate because we use a small subsets of observations, the change point estimate remains good.
-A two stage change point estimation approach is also possible as explained in  \citet{iacyos09}.
-
-\subsection{LASSO model selection}
+Notice that, even if the estimated parameters are not too accurate because we use a small subsets of observations, the change-point estimate remains good.
+A two stage change-point estimation approach is also possible as explained in  \citet{iacyos09}.
+
+\subsection{LASSO Model Selection}
 Let $X_t$ be a  diffusion process solution to
@@
 
-\subsection{An example of use}
+\subsection{An Example of Use}
 The \code{lasso} method is implemented in the \pkg{yuima} package.
@@
 \end{document}
-
-The YUIMA Project\footnote{The Project has been  funded up to 2010 by the Japan Science Technology (JST) Basic Research Programs PRESTO, Grants-in-Aid for Scientific Research No.
-19340021.} is an open source\footnote{All code in the \pkg{yuima} package is subject to the GNU General Public License, Version 2, see \url{http://www.gnu.org/licenses/gpl-2.0.html}.}
- academic project aimed at developing the \proglang{R} package named ``\pkg{yuima}'' for simulation and inference of stochastic differential equations. 
-The YUIMA Project is mainly developed by mathematicians and 
-statisticians who actively publish in the field of inference and simulation for stochastic
-differential equations.
-%The YUIMA Project Core Team, currently 
-%consists of the following people: A. Brouste, M. Fukasawa, H. Hino, S.M. Iacus, K. Kamatani, H.Masuda, Y. Shimizu, M. Uchida, N. Yoshida.
-
- The \pkg{yuima} package  provides 
-an object-oriented programming environment 
-for simulation and statistical inference 
-for stochastic processes by \proglang{R}.
-The \pkg{yuima} package adopts the  $\tt S4$ system of classes and methods \citep{chambers98}.
-
-Under this framework, 
-the \pkg{yuima} package also supplies various functions 
-to execute simulation and statistical analysis. 
-Both categories of procedures may depend each other. 
-Statistical inference often requires a simulation technique 
-as a subroutine, and a certain simulation method 
-needs to fix a tuning parameter by applying 
-a statistical methodology. 
-It is especially the case of stochastic processes 
-because most of expected values involved 
-do not admit an explicit expression. 
-The \pkg{yuima} package facilitates comprehensive, systematic approaches 
-to the solution.
-
-Stochastic differential equations are 
-commonly used 
-to model random evolution along continuous or 
-practically continuous time, such as 
-the random movements of a stock price. 
-Theory of statistical inference for 
-stochastic differential equations already has a fairly long history, 
-more than three decades, but it is still developing quickly new 
-methodologies and expanding the area. 
-The formulas produced by the theory are usually very sophisticated, 
-which makes it difficult for practitioners not necessarily 
-familiar with this field to enjoy their utility. 
-For example, the asymptotic expansion method for computing 
-asian option prices (i.e., expectation of a functional of 
-a stochastic process) provides precise approximation values 
-instantaneously, taking advantage of the analytic approach, 
-but the formula, based on Malliavin calculus, has a long expression like more than one page!
-
-The \pkg{yuima} package delivers up-to-date methods as a package 
-onto the desk of the user working 
-with simulation and/or statistics for stochastic differential equations.
-
-Sampled data from a continuous-time process features 
-the time stamps as well as the positions of the object. 
-It is requiring a new theory of estimation. 
-The \pkg{yuima} framework can apply multi-dimensional time stamps 
-of tick data and provides diverse functions handling such kind data 
-to support statistical analysis.

### 2012-07-25 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
revision
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@200 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2011-11-11 pkg/yuimadocs/inst/doc/JSS/article.Rnw
@@
 to execute simulation and statistical analysis. 
-Both categories of procedures may depend each other. 
+Both categories of procedures may depend \textcolor{red}{on} each other.
 Statistical inference often requires a simulation technique 
@@
 familiar with this field to enjoy utilities. 
-For example, the asymptotic expansion method for computing 
+\textcolor{red}{For example, by taking advantage of the analytic approach,  the asymptotic expansion method for computing 
 option prices (i.e., expectation of an irregular functional of 
 a stochastic process) provides precise approximation values 
-instantaneously, taking advantage of the analytic approach, 
-but the formula has a long expression like more than 900 terms!
+instantaneously. The expansion formula, which has a long expression involving more than 900 terms of multiple integrals, is already coded in the \pkg{yuima} package for generic diffusion processes.}
 
@@
 For example, in case one wants to simulate a stochastic process, only the slots \code{model} and \code{sampling} should be present, while the slot \code{data} will be filled by the simulator.
-We now discuss in details the different object separately.
+We now discuss in details the \textcolor{red}{different objects} separately.
 
@@
 \item \code{jump} contains the parameters belonging to the  coefficient of the L\'evy noise;
-\item \code{measure} contains the parameters describing to the L\'evy measure;
+\item \code{measure} contains the parameters \textcolor{red}{describing the L\'evy} measure;
 \end{itemize}
@@
   are assumed to be \code{x} and \code{t} but the user can freely choose them.
-  The \code{yuima.model} class assumes that the user either use default
+  The \code{yuima.model} class assumes that the user either \textcolor{red}{uses} default
   names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
@@
 \section{Model Specification}\label{sec3}
-In order to show how general is the approach in the \pkg{yuima} package we present some examples.
+In order to show \textcolor{red}{how general the approach is} in the \pkg{yuima} package we present some examples.
 
@@
 $$\de X_t = -3 X_t \de t + \frac{1}{1+X_t^2}\de W_t$$
-This is done in \pkg{yuima} specifying the drift and diffusion coefficients as plain mathematical expressions
+This is done in \pkg{yuima} \textcolor{red}{by} specifying the drift and diffusion coefficients as plain mathematical expressions
 <<print=FALSE,echo=FALSE,results=hide>>=
@@
 
-\subsubsection{Fractional Gaussian Noise}
+\subsection{Fractional Gaussian Noise}
 In order to specify a stochastic differential equation driven by fractional Gaussian noise it is necessary to specify the value of the Hurst parameter.
@@
 @
-The user can choose between the two simulation schemes, namely the Cholesky method and \citet{WoodChan} method.
+The user can choose \textcolor{red}{between  two} simulation schemes, namely the Cholesky method and \citet{WoodChan} method.
 
@@
 The \code{simulate} function accepts several arguments including the description of the sampling structure, which is an object of type \code{yuima.sampling}.
-The \code{setSampling} allow for the specification of different sampling parameters including random sampling.
-Further, the \code{subsampling} allow to subsample a trajectory of a simulated stochastic differential equation or a given time series in the \code{yuima.data} slot of a \code{yuima} object.
+The \code{setSampling} \textcolor{red}{allows} for the specification of different sampling parameters including random sampling.
+Further, the \code{subsampling} \textcolor{red}{allows us} to subsample a trajectory of a simulated stochastic differential equation or a given time series in the \code{yuima.data} slot of a \code{yuima} object.
 Sampling and subsampling can be specified jointly as arguments to the \code{simulate} function.
 This is convenient if one wants to simulate data at very high frequency but then return only low frequency data for inference or other applications.
-We now go through few examples just to describe the use of standard arguments to these functions but the reader is invited to go thorough the man pages of the \code{yuima} packages for complete details.
+\textcolor{red}{In what follows we explain how to specify  arguments of these \pkg{yuima} functions.
+Complete details can be found in  the man pages of the \pkg{yuima} package.}
 
@@
 @
-As seen from the output, the sampling structure is quite reach and we will show how to specify few of the slots later one.
+As seen from the output, the sampling structure is \textcolor{red}{quite rich} and we will show how to specify \textcolor{red}{some} of the slots \textcolor{red}{later on}.
 We now simulate this process specifying the \code{sampling} argument to \code{simulate}
@@
 @
-In the above we have specified two independent exponential  distributions to represent Poissoninan random times. Notice that the slot \code{regular} is now set to \code{FALSE}.
+In the above we have specified two independent exponential  distributions to represent \textcolor{red}{Poisson arrival times}. Notice that the slot \code{regular} is now set to \code{FALSE}.
 Now we subsample the original trajectory of \code{X2} using the \code{subsampling} function
@@
 $$\de X_t^\ve = a(X_t^\ve,\ve)\de t + b(X_t^\ve,\ve)\de W_t, \qquad \ve \in(0,1]$$
-with $W_t$ and $r$-dimensional Wiener process, i.e., $W_t=(W_t^1, \ldots, W_t^r)$.
+with $W_t$ \textcolor{red}{an} $r$-dimensional Wiener process, i.e., $W_t=(W_t^1, \ldots, W_t^r)$.
 The functional is expressed in the following abstract form
@@
 \end{eqnarray}}
-where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, $A^{\otimes 2}= A^T A$ and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
+where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, \textcolor{red}{$A^{\otimes 2}= A A^T$} and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
 Then, \citep[see e.g.,][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
-$$\tilde\theta_n=\arg\min_\theta \ell_n({\bf X}_n,\theta)$$
+\textcolor{red}{$$\hat\theta_n=\arg\max_\theta \ell_n({\bf X}_n,\theta)$$}
+
+\textcolor{red}{The \pkg{yuima} package implements QML estimation via the  \code{qmle} function.
+The interface and the output of the \code{qmle} function are made as similar as possible to those of the standard \code{mle} function in the \pkg{stats4} package of the basic \proglang{R} system.
+The main arguments to  \code{qmle} consist of a \code{yuima} object and initial values (\code{start}) for the optimizer.
+The \code{yuima} object must contain the slots \code{model} and \code{data}.
+The \code{start} argument must be specified as a named list, where the  names of the  elements of the list  correspond to the names of the parameters as they appear in the \code{yuima} object.
+Optionally, one can specify named lists of \code{upper}  and \code{lower} bounds
+to identify the search region of the optimizer.
+The standard optimizer is \code{BFGS} when no bounds are specified.
+If bounds are specified then \code{L-BFGS-B} is used.
+More optimizers can be added in the  future.}
+
 As an example, we consider the simple model
@@
 @
-With the sampled data we can use the function \code{qmle} to estimate the parameters as follows
+\textcolor{red}{Now the \code{yuima} object contains both the \code{model} and \code{data} slots.
+We set the initial values for the optimizer as $\theta_1=\theta_2=0.5$ and we specify them as a named list called \code{param.init}.
+We can now use the function \code{qmle} to estimate the parameters $\theta_1$ and $\theta_2$ as follows}
 <<echo=TRUE, print=FALSE,results=hide>>=
 param.init <- list(theta2=0.5,theta1=0.5)
-mle1 <- qmle(yuima, start =param.init , 
-     lower = list(theta1=0, theta2=0), 
-     upper = list(theta1=1, theta2=1))
-@
-and the estimated coefficients are extracted from the output object \code{mle1} as follows
+upp.par <-  list(theta1=0, theta2=0)
+low.par <-  list(theta1=1, theta2=1)
+mle1 <- qmle(yuima, start = param.init,  lower = upp.par, upper = low.par)
+@
+\textcolor{red}{where \code{upp.par} and \code{low.par} are the upper and lower bounds of the search region used by the optimizer (\code{L-BFGS-B} in this case).} 
+\textcolor{red}{The} estimated coefficients are extracted from the output object \code{mle1} as follows
 <<echo=TRUE, print=FALSE>>=
@@
 @
-Notice the interface and the output of the \code{qmle} is quite similar to the ones of the standard \code{mle} function of the \pkg{stats4} package of the base \proglang{R} system.
+%Notice the interface and the output of the \code{qmle} is quite similar to the ones of the standard \code{mle} function of the \pkg{stats4} package of the base \proglang{R} system.
 
@@
 =  
-\Big[\int_{\Theta_1}\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
-\pi_1(\theta_1)d\theta_1 \Big]^{-1} \int_{\Theta_1} \theta_1 \ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
-\pi_1(\theta_1)d\theta_1
+\Big[\int_{\Theta_1}\textcolor{red}{\exp\{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))\}}
+\pi_1(\theta_1)d\theta_1 \Big]^{-1} \int_{\Theta_1} \theta_1 \textcolor{red}{\exp \{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
+\}}\pi_1(\theta_1)d\theta_1
 \end{equation}
@@
 = 
-\Big[\int_{\Theta_2}\ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))
-\pi_2(\theta_2)d\theta_2 \Big]^{-1} \int_{\Theta_2} \theta_2 \ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))
+\Big[\int_{\Theta_2}\textcolor{red}{\exp\{ \ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}}
+\pi_2(\theta_2)d\theta_2 \Big]^{-1} \int_{\Theta_2} \theta_2 \textcolor{red}{\exp \{\ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}}
 \pi_2(\theta_2)d\theta_2
@@
 irregularly spaced, in general. 
-In particular, cce can apply to estimation of the quadratic variation of a single 
+In particular, \textcolor{red}{\code{cce}} can apply to estimation of the quadratic variation of a single 
 stochastic process sampled regularly/irregularly.
@@
 \subsubsection{Example: Data Generation and Estimation by \pkg{yuima} Package}
-We will demonstrate how to apply cce function to 
+We will demonstrate how to apply \textcolor{red}{\code{cce}} function to 
 nonsynchronous high-frequency data by simulation. 
@@
 @
-$\tt cce$ takes the sample and returns 
+\code{cce} takes the sample and returns 
 an estimate of the quadratic covariation. For example, for the complete data
@@
 $$
-where $W_t$ a $r$-dimensional Wiener process and $b_t$ and $X_t$ are multidimensional processes and $\sigma$ is the diffusion coefficient (volatility) matrix.
+where $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process and $b_t$ and $X_t$ are multidimensional processes and $\sigma$ is the diffusion coefficient (volatility) matrix.
 When $Y=X$ the problem is a diffusion model.
@@
 \log\det S(X_{t_{i-1}},\theta)
-+ \Delta_n^{-1}(\Delta_iY)'S(X_{t_{i-1}},\theta)^{-1}(\Delta_iY). 
++ \Delta_n^{-1}(\Delta_iY)'S(X_{t_{i-1}},\theta)^{-1}(\Delta_iY)
 \label{eq:GiCH10}
 \end{equation}
+\textcolor{red}{and $S=\sigma^{\otimes 2}$.}
 Suppose that there exists an estimator 
@@
  \de W_t^1\\ \de W_t^2 
- \end{array}\right)
+ \end{array}\right),\quad \textcolor{red}{ t\in[0,T],}
 $$
@@
 $$
-with change-point instant at time $\tau=0.4$.   First, we describe the model to be simulated
+with change-point instant at time $\tau=0.4$ \textcolor{red}{and $T=10$}.   First, we describe the model to be simulated
 <<cpoint1,eval=TRUE, echo=TRUE,results=hide>>=
@@
 @
-and then simulate two trajectories. One up to the change-point $\tau=4$ with parameters $\theta_{1.0}=0.1$ and  $\theta_{2.0}=0.2$
+and then simulate two trajectories. One up to the change-point $\tau=4$ with parameters \textcolor{red}{$\theta_{1.0}=0.5$ and  $\theta_{2.0}=0.3$}
 <<cpoint3,results=hide>>=
@@
 t0 <- list(theta1.k=0.5, theta2.k=0.3)
-
-tau <- 0.4
-ysamp1 <- setSampling(n=tau*n, Initial=0, delta=0.01)
+T <- 10
+tau <- 4
+pobs <- tau/T
+ysamp1 <- setSampling(n=n*pobs, Initial=0, delta=0.01)
 yuima1 <- setYuima(model=ymodel, sampling=ysamp1)
@@
  with parameters
-$\theta_{1.1}=0.6$ and  $\theta_{2.1}=0.6$.
+\textcolor{red}{$\theta_{1.1}=0.2$ and  $\theta_{2.1}=0.4$}.
 For this trajectory, the initial value is set to the last value of the first trajectory stored in \code{x1} and \code{x2} for the two component of the process.
@@
 t1 <- list(theta1.k=0.2, theta2.k=0.4)
-ysamp2 <- setSampling(Initial=n*tau*0.01, n=n*(1-tau), delta=0.01)
+ysamp2 <- setSampling(Initial=n*pobs*0.01, n=n*(1-pobs), delta=0.01)
 yuima2 <- setYuima(model=ymodel, sampling=ysamp2)
@@
 which are useful in the framework of change-point or sequential analysis.
-The function \code{qmleL} estimates a model by quasi maximum likelihood using observations in the time interval $[0,t]$ where $t$ cam be specificed by the user.
-In our example, we set \code{t=0.2}.
-Similary for \code{qmleR}, which uses only observations in the time interval $[t, T]$.
-In our example, we take \code{t=0.8}.
+The function \code{qmleL} estimates a model by quasi maximum likelihood using observations in the time interval $[0,t]$ where $t$ can be specificed by the user.
+In our example, we set \code{t=2}.
+Similarly for \code{qmleR}, which uses only observations in the time interval $[t, T]$.
+In our example, we take \code{t=8}.
 <<>>=
@@
 $$
-\mathcal{F}(\theta)=(\theta-\tilde{\theta}_n)\ddot{\mathbb{H}}_n({\bf X}_n, \tilde\theta_n)(\theta-\tilde{\theta}_n)'+\sum_{j=1}^p\lambda_{n,j}|\alpha_j| +\sum_{k=1}^q\gamma_{n,k}|\beta_k|
+\mathcal{F}(\theta)=(\theta-\tilde{\theta}_n)^T\ddot{\mathbb{H}}_n({\bf X}_n, \tilde\theta_n)(\theta-\tilde{\theta}_n)+\sum_{j=1}^p\lambda_{n,j}|\alpha_j| +\sum_{k=1}^q\gamma_{n,k}|\beta_k|
 $$
@@
 \section{Miscellanea and Roadmap of YUIMA Project}\label{sec6}
-Other statistical techniques are already implemented or will be shortly releases in the \pkg{yuima}.
+Other statistical techniques are already implemented or will be shortly released in the \pkg{yuima}.
 For example, a nice utility is the function \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}.

### 2012-07-26 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
updated
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@201 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% declarations for jss.cls %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\usepackage{Rd}
+
+%%% INFOS FOR YUIMA CORE TEAM %%%
+
+% USE THIS IF YOU DON'T WANT TO EXECUTE R CODE
+\SweaveOpts{prefix.string=yuima, echo=TRUE, eval=FALSE}
+
+% USE THIS INSTEAD IF YOU WANT TO EXECUTE R CODE
+%\SweaveOpts{prefix.string=yuima, echo=TRUE, eval=TRUE}
+
+%% HOW TO BUILD THIS FILE
+%% step 1) Sweave will produce LaTeX file
+%% step 2) remove all out file (don't ask why)
+%% step 3) pdflatex the .tex file
+%% all the above steps should be done VIA R CMD as
+%% explained in the next lines:
+
+%% FROM A TERMINAL SHELL, IN THE SAME DIRECTORY OF THE .RNW FILE 
+%% TYPE THESE THREE COMMANDS IN SEQUENCE
+%% step 1)   R CMD Sweave article-new
+%% step 2)   rm *.out
+%% step 3)   R CMD pdflatex article-new
+
+%% you can repeat steps 2) and 3) to update references etc.
 
@@
 %% Fax: +43/1/31336-734
-
-\SweaveOpts{prefix.string=yuima} 
-\SweaveOpts{echo=TRUE}
 
@@
 Then we call $\tt adaBayes$, on the same sample data we used for the \code{qmle} function, as follows
-<<eval=TRUE,echo=TRUE, results=hide>>=
+<<echo=TRUE, results=hide>>=
 bayes1 <- adaBayes(yuima, start=param.init, prior=prior)
@@
 and we can compare the adaptive Bayes estimates  with the QMLE estimates
-<<eval=TRUE,echo=TRUE>>=
+<<echo=TRUE>>=
 coef(bayes1)
@@
 with change-point instant at time $\tau=0.4$ \textcolor{red}{and $T=10$}.   First, we describe the model to be simulated
-<<cpoint1,eval=TRUE, echo=TRUE,results=hide>>=
+<<cpoint1, echo=TRUE,results=hide>>=
 diff.matrix <- matrix(c("theta1.k*x1","0*x2","0*x1","theta2.k*x2"), 2, 2)

### 2012-07-26 pkg/yuimadocs/inst/doc/JSS/script.sh
sun this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@203 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,24 @@
+#!/bin/bash
+date
+before="$(date +%s)"
+echo 'SWEAVE | PDFLATEX | BIBTEX | PDFLATEX | PDFLATEX. Custom engine--Akasurak-16Nov2009'
+#Updated 20Jul2010 for auto including Sweave.sty
+
+
+export PATH=$PATH:/usr/texbin:/usr/local/bin
+
+R CMD Sweave "$1"
+rm *.out
+R CMD pdflatex "${1%.*}"
+bibtex  "${1%.*}.aux"
+rm *.out
+R CMD pdflatex "${1%.*}"
+rm *.out
+R CMD pdflatex "${1%.*}"
+
+open "$1.pdf"
+
+after="$(date +%s)"
+elapsed_seconds="$(expr $after - $before)"
+date
+echo Elapsed time '(m:s)': $(date -r $elapsed_seconds +%M:%S)

### 2013-02-05 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
updated
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@202 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
sun this
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@204 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
my changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@205 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
my update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@219 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+%% before editing this file get the new version, type this command on Terminal
+%% in the same directory where this Rnw-file lives
+%% svn up
+
 %% HOW TO BUILD THIS FILE
-%% step 1) Sweave will produce LaTeX file
-%% step 2) remove all out file (don't ask why)
-%% step 3) pdflatex the .tex file
-%% all the above steps should be done VIA R CMD as
-%% explained in the next lines:
-
-%% FROM A TERMINAL SHELL, IN THE SAME DIRECTORY OF THE .RNW FILE 
-%% TYPE THESE THREE COMMANDS IN SEQUENCE
-%% step 1)   R CMD Sweave article-new
-%% step 2)   rm *.out
-%% step 3)   R CMD pdflatex article-new
-
-%% you can repeat steps 2) and 3) to update references etc.
+%% from inside the directory where this Rnw-file lives
+%% ./script.sh article-new
+
+%% to commit to svn the changes type
+%% from inside the directory where this Rnw-file lives
+%% svn commit -m "my changes"
 
@@
 For example, in case one wants to simulate a stochastic process, only the slots \code{model} and \code{sampling} should be present, while the slot \code{data} will be filled by the simulator.
-We now discuss in details the \textcolor{red}{different objects} separately.
-
-\subsection{The \code{yuima.model} Class}
-At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified. All multidimensional and eventually as parametric models.
+We discuss in details the \textcolor{red}{different objects} separately in the next sections.
+
+\textcolor{blue}{The general idea of the \pkg{yuima} package is to separate into different subclass objects the statistical model, the data and the statistical methods.
+As will be explained with several examples, the user may give a mathematical description of the statistical model with \code{setModel} which prepares a \code{yuima.model} object by filling the appropriate slots.
+If the aim is the simulation of the solution of the stochastic  differential equation specified in the \code{yuima.model} object then, using the method \code{simulate}, it is possible to obtain one trajectory of the process.
+As an output, a \code{yuima} object is created which contains the original model specified in the \code{yuima.model} object in the slot named \code{model} and two additional slots named \code{data}, for the simulated data, and \code{sampling} which contains the description of the simulation scheme used as well as other informations.
+The details of \code{simulate} will be explained in Section \ref{sec:simul} along with the use of method \code{setSampling} which allows to specify the type of sampling scheme to be used by the \code{simulate} method.
+But \code{yuima} object may contain the slot \code{data} non only as the outcome of \code{simulate} but also because the user decides to analyse its own data.
+In this case the method \code{setData} is used to transform most types  of \proglang{R} time series object into a a proper \code{yuima.data} object.
+When the slots \code{data} and \code{model} are available, many other methods can be used to perform statistical analysis on these sde's models. These methods will be discussed in Section \ref{sec5}.
+Further, functionals of stochastic differential equations can be defined using the \code{setFunctional} method and evaluated using asymptotc expansion methods as explained in Section \ref{sec4}.
+The \code{setFunctional} method creates a \code{yuima.functional} object which is included along with a \code{yuima.model} object into a \code{yuima} object in order to evaluate the asymptotic expansion.}
+
+\subsection{The \code{yuima.model} Class}\label{sec:model}
+At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified. \textcolor{blue}{Here we present  a brief
+overview of these models as they will be described in details in Section \ref{sec3}, but this allow to introduce an overall view of the slots of the \code{yuima.model} class.}. 
+\textcolor{blue}{In \pkg{yuima} one can describe three main families of stochastic processes at present.
+These models can be one or multidimensional and eventually described as parametric models.
+Let $X_0=x_0$ be the initial value of the process, then, the main classes are as follows:}
 \begin{itemize}
-\item diffusions $\displaystyle  \de X_t=a(t,X_t)dt + b(t,X_t)\de W_t $, where $W_t$ is a standard Brownian motion;
-\item fractional Gaussian noise, with $H$ the Hurst parameter
-$$ \de X_t=a(t,X_t)dt + b(t,X_t)\de W_t^{H}; $$ 
-\item diffusions with jumps and  L\'evy processes solution to
+\item \textcolor{blue}{diffusion models described as} 
+$$  \de X_t=a(t,X_t,\theta)dt + b(t,X_t,\theta)\de W_t, \quad X_0=x_0$$
+ where $W_t$ is a standard Brownian motion;
+\item \textcolor{blue}{sde's driven by fractional Gaussian noise, with $H$ the Hurst parameter, described as}
+$$ \de X_t=a(t,X_t,\theta)dt + b(t,X_t,\theta)\de W_t^{H}, \quad X_0=x_0;$$ 
+\item \textcolor{blue}{diffusion process with jumps} and  L\'evy processes solution to
 $$
 \begin{aligned}
-\de X_t = & \,\,\, a(X_t)\de t + b(X_t)\de W_t + \int\limits_{|z|>1}\!\!\! c(X_{t-},z)\mu(\de t,\de z) \\
-&{}+\!\! \int\limits_{0<|z|\le 1}\!\!\!  c(X_{t-},z)\{\mu(\de t,\de z) - \nu(\de z)\de t\}.
-\end{aligned}
+\de X_t = & \,\,\, a(t,X_t,\theta)\de t + b(t,X_t,\theta)\de W_t + \int\limits_{|z|>1}\!\!\! c(X_{t-},z)\mu(\de t,\de z) \\
+&{}+\!\! \int\limits_{0<|z|\le 1}\!\!\!  c(X_{t-},z)\{\mu(\de t,\de z) - \nu(\de z)\de t\}
+\end{aligned}, \quad X_0=x_0;
 $$
 \end{itemize}
+\textcolor{blue}{The functions $a(\cdot)$, $b(\cdot)$ and $c(\cdot)$ may have a different number of arguments.
+For example, if the model is homogeneous in time and drift and diffusion coefficients are entirely specified, then we will use the notaion $a(x)$ and $b(x)$ and describe the diffusion model simply as $\displaystyle \de X_t=a(X_t)dt + b(X_t)\de W_t$.
+And so forth.
+Detailed hypotheses and regularity conditions on the coefficients $a(\cdot)$, $b(\cdot)$ and $c(\cdot)$ for each class of models will be given in the next sections.
+Nevertheless, it is important to remark that these notations only matter the mathematical description of the model while each coefficient is passed to \pkg{yuima} methods as \proglang{R} mathematical expressions.
+It means that, for example, $a(t, X_t, \theta) = t\cdot \sqrt{\theta X_t}$ will be passed as \code{t*sqrt(x*theta)}, therefore, the order of the arguments is not relevant to \proglang{R} as well as its mathematical description as long as it is consistent through each specific section.
+Furhter, \pkg{yuima} is able to accept any user-specified notation for the state variable $x$ (for $X_t$) and the time variable $t$ and the remaining term of an \proglang{R} expression will be interpreted as parameter as will explained in Section \ref{sec:diff}.
+We are now able to give an overview of the mian slots of the most important class of the \pkg{yuima} package.}
 
@@
   coefficient matrix relative to first noise;
-\item \code{hurst} is the Hurst index of the fractional Brownian motion, by default \code{0.5}  meaning a standard Brownian motion;
-\item \code{parameter} which is a short name for ``parameters'' which 
-  is a list with the following entries:
+\item \code{hurst} is the Hurst index of the fractional Brownian motion, by default \code{0.5}  meaning a standard Brownian motion.
+\textcolor{blue}{More details will be given in Section \ref{sec:fgn}};
+\item \code{parameter},  which is a short name for ``parameters'',  
+  is a list with the following entries \textcolor{blue}{(explained details in Section \ref{sec:par})}:
 \begin{itemize}
@@
 \item \code{jump} contains the parameters belonging to the  coefficient of the L\'evy noise;
-\item \code{measure} contains the parameters \textcolor{red}{describing the L\'evy} measure;
+\item \code{measure} contains the parameters \textcolor{red}{describing the L\'evy} measure (\textcolor{blue}{(explained details in Section \ref{sec:jump})};
 \end{itemize}
-\item \code{measure} specifies the measure of the L\'evy noise;
-\item \code{measure.type} a switch to specify how the L\'evy measure is described;
+\item \code{measure} specifies the measure of the L\'evy noise (\textcolor{blue}{(see Section \ref{sec:jump})};
+\item \code{measure.type} a switch to specify how the L\'evy measure is described (\textcolor{blue}{(see Section \ref{sec:jump})};
 \item \code{state.variable} and \code{time.variable}, by default,
-  are assumed to be \code{x} and \code{t} but the user can freely choose them.
+  are assumed to be \code{x} and \code{t} but the user can freely choose them \textcolor{blue}{and they matter the right hand-side of the equation of the sde}.
   The \code{yuima.model} class assumes that the user either \textcolor{red}{uses} default
@@
   the rest of the symbols are considered parameters and distributed accordingly
-  in the \code{parameter} slot.
-\item \code{jump.variable} the name of the variable used in the description of the L\'evy component;
+  in the \code{parameter} slot. \textcolor{blue}{Example of use will be given in Section \ref{sec:diff};}
+\item \code{jump.variable} the name of the variable used in the description of the L\'evy component (\textcolor{blue}{(see Section \ref{sec:jump})};
 \item \code{solve.variable} contains a vector of variable names, each element corresponds to the
-   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order.
+   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order. \textcolor{blue}{An example of use can be found in \Section \ref{sec:multi}.}
 \item \code{noise.number} indicates the number of sources of noise.
@@
 
-\subsection{Diffusion Processes}
+\subsection{Diffusion Processes}\label{sec:diff}
 Assume that we want to describe the following stochastic differential equation
-$$\de X_t = -3 X_t \de t + \frac{1}{1+X_t^2}\de W_t$$
-This is done in \pkg{yuima} \textcolor{red}{by} specifying the drift and diffusion coefficients as plain mathematical expressions
+$$\de X_t = -3 X_t \de t + \frac{1}{1+X_t^2}\de W_t.$$
+\textcolor{blue}{In the above $a(x) = -3 x$ and $b(x) = \frac{1}{1+x^2}$ according to the notaion of previous section}
+This \textcolor{blue}{can be described} in \pkg{yuima} \textcolor{red}{by} specifying the drift and diffusion coefficients as plain 
+\textcolor{blue}{\proglang{R} expressions passed as strings}
 <<print=FALSE,echo=FALSE,results=hide>>=
@@
 @ 
+\textcolor{blue}{By default, \pkg{yuima} assumes that the state variable is \code{x} and the time variable is \code{t} and the solve variable is again \code{x}.
+Notice that the left hand-side of the equation is implicit, this is why \code{yuima.model} has the slot \code{solve.variable}.
+The user should not we worried about the warning raised by \pkg{yuima} at this stage, as this is just to inform the user about the implicit assumption on the solution variable.}
 At this point, the package fills the proper slots of the \code{yuima} object
@@
 
-For the above, it is possible to see that the jump coefficient is void and the Hurst parameter is set to 0.5, because this corresponds to the standard Brownian motion.
-Now, with \code{mod1} in hands, it is very easy to simulate a trajectory of the process as follows
+From the above, it is possible to see that the jump coefficient is void and the Hurst parameter is set to 0.5, \textcolor{blue}{because this is a model where the driving process is the standard Brownian motion, i.e.
+a fractional Brownian motion if Hurst index $H=\frac12$.}
+Now, with \code{mod1} in hands, it is very easy to simulate a trajectory \textcolor{blue}{Euler-Maruyama scheme} of the process as follows
 <<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
@@
 @
-
-\subsection{Specification of Parametric Models}
-When a parametric model like
+\textcolor{More details on how to change the default sampling scheme for the \code{simulate} method and how to perform subsampling will be given in Section \ref{sec:simul}.}
+
+\subsection{User specified state and time variables: NEW SECTION}
+Suppose now that the user wants to specify her own model using a prescribed notation, e.g., some sdes like
+$$\de Y_s = -3 s Y_s \de s + \frac{1}{1+Y_s^2}\de W_s,$$
+then this can be done in \pkg{yuima} as follows
+<<echo=TRUE, print=FALSE,results=hide>>=
+mod1b <- setModel(drift = "-3*s*y", diffusion = "1/(1+y^2)", state.var="y", time.var="s")
+@
+In this case the solution variable is the same as the state variable. Indeed, the \code{yuima.model} object appears as follows
+<<>>=
+str(mod1b)
+@
+Once again, the user may use the \code{simulate} method to perform simulation.
+
+\subsection{Specification of Parametric Models}\label{sec:par}
+\textcolor{blue}{Assume now that we want to specify a parametric model like this}
 $$\de X_t = -\theta X_t \de t + \frac{1}{1+X_t^\gamma}\de W_t$$
-is specified, \pkg{yuima} attempts to distinguish the parameters' names from the ones of the state and time variables
+\textcolor{blue}{where $a(x,\theta) = -\theta x$ and $b(x,\gamma) = 1/(1+x^\gamma)$.}
+Then, \pkg{yuima} attempts to distinguish the parameters' names from the ones of the state and time variables
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 @
+\textcolor{blue}{so, in this case, \code{theta} and \code{gamma}, which are different form \code{x} and \code{t}, are assumed to be parameters.
+Notice that in the above notation $\theta$ and $\gamma$ are generic names for the components of a parameters' vector $\theta$ in the notation of Section \ref{sec:model}.}
 <<>>=
@@
 @
-In order to simulate the parametric model it is necessary to specify the values of the parameters as the next code shows
+In order to simulate the parametric model it is necessary to specify the values of the \textcolor{blue}{parameters $\theta$ and $\gamma$ as shown in the next code chunk}
 <<echo=TRUE, print=FALSE,fig=TRUE,height=4,results=hide>>=
@@
 
-\subsection{Multidimensional Processes}
-Next is an example with two stochastic differential equations driven by three independent Brownian motions
+\subsection{Multidimensional Processes}\label{sec:multi}
+\textcolor{blue}{Next is an example of a system of two stochastic differential equations for the couple $(X_{t,1}, X_{t,2})$ driven by three independent Brownian motions $(W_{t,1}, W_{t,2}, W_{t,3})$}
 $$
 \begin{aligned}
-\de X^1_t &= -3 X^1_t \de t + \de W^1_t + X^2_t \de W^3_t\\
-\de X^2_t &= -(X^1_t + 2 X^2_t) \de t + X^1_t \de W^1_t + 3 \de W^2_t
+\de X_{1,t} &= -3 X_{1,t} \de t + \de W_{1,t} + X_{2,t} \de W_{3,t}\\
+\de X_{2,t} &= -(X_{1,t} + 2 X_{2,t}) \de t + X_{1,t} \de W_{1,t} + 3 \de W_{2,t}
 \end{aligned}
 $$
-but this has to be organized into matrix form
-$$
-\left(\begin{array}{c}\de X^1_t \\\de X^2_t\end{array}\right)=
-\left(\begin{array}{c} -3 X^1_t \\  -X^1_t - 2X^2_t\end{array}\right)\de t +
-\left[\begin{array}{ccc}1 & 0 & X^2_t \\X^1_t & 3 & 0\end{array}\right]
-\left(\begin{array}{c}\de W_t^1 \\ \de W_t^2 \\ \de W_t^3\end{array}\right)
-$$
+but this has to be organized into matrix form \textcolor{red}{with a vector of drift expressions and a diffusion matrix}
+$$
+\left(\begin{array}{c}\de X_{1,t} \\\de X_{2,t}\end{array}\right)=
+\left(\begin{array}{c} -3 X_{1,t} \\  -X_{1,t} - 2X_{2,t}\end{array}\right)\de t +
+\left[\begin{array}{ccc}1 & 0 & X_{2,t} \\X_{1,t} & 3 & 0\end{array}\right]
+\left(\begin{array}{c}\de W_{t,1} \\ \de W_{t,2} \\ \de W_{t,3}\end{array}\right)
+$$
+\textcolor{blue}{For this system it now necessary to instruct \pkg{yuima} about the state variable on bot the left-hand side of the equation and the right-hand side of the equation, i.e.
+the \code{solve.variable}.}
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 
-\subsection{Fractional Gaussian Noise}
+\section{Fractional Gaussian Noise}\label{sec:fgn}
 In order to specify a stochastic differential equation driven by fractional Gaussian noise it is necessary to specify the value of the Hurst parameter.
@@
 
-\subsection{L\'evy Processes}
+\subsection{L\'evy Processes}\label{sec:jump}
 Jump processes can be specified in different ways in mathematics and hence in \pkg{yuima} package. 
@@
 
-\subsection{Simulation, Sampling and Subsampling}
+\section{Simulation, Sampling and Subsampling}\label{sec:simul}
 The \code{simulate} function simulates \code{yuima} models according to Euler-Maruyama scheme in the presence of non-fractional diffusion noise and L\'evy jumps and uses the Cholesky or the Wood and Chan methods for the fractional Gaussian noise.

### 2013-02-06 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@220 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[utf8x]{inputenc}
+
+\newcommand{\colorr}{\color[rgb]{0.8,0,0}}
+\newcommand{\colorg}{\color[rgb]{0,0.5,0}}
+\newcommand{\colorb}{\color[rgb]{0,0,0.8}}
+\newcommand{\colord}{\color[rgb]{0.8,0.3,0}}
+\newcommand{\colorn}{\color[rgb]{1,1,1}}
+\newcommand{\colorred}{\color[rgb]{0.8,0,0}}
+\newcommand{\colory}{\color{yellow}}
+\newcommand{\coloro}{\color[rgb]{1,0.4,0}}%{1,0.851,0}}
+\newcommand{\coloroy}{\color[rgb]{1,0.95,0}}
+\newcommand{\colorsb}{\color[rgb]{0,0.95,1}}
+\newcommand{\colorro}{\color[rgb]{0.851,0.255,0.467}} %rose
 
@@
 
+\def\ep{{\varepsilon}}
 \def\ve{{\varepsilon}}
@@
 Further, functionals of stochastic differential equations can be defined using the \code{setFunctional} method and evaluated using asymptotc expansion methods as explained in Section \ref{sec4}.
-The \code{setFunctional} method creates a \code{yuima.functional} object which is included along with a \code{yuima.model} object into a \code{yuima} object in order to evaluate the asymptotic expansion.}
+The \code{setFunctional} method creates a \code{yuima.functional} object which is included along with a \code{yuima.model}  into a \code{yuima} object in order to be used for the evaluation of its asymptotic expansion.}
 
@@
 At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified. \textcolor{blue}{Here we present  a brief
-overview of these models as they will be described in details in Section \ref{sec3}, but this allow to introduce an overall view of the slots of the \code{yuima.model} class.}. 
+overview of these models as they will be described in details in Section \ref{sec3}, but this allow to introduce an overall view of the slots of the \code{yuima.model} class.} 
 \textcolor{blue}{In \pkg{yuima} one can describe three main families of stochastic processes at present.
@@
 \item \code{solve.variable} contains a vector of variable names, each element corresponds to the
-   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order. \textcolor{blue}{An example of use can be found in \Section \ref{sec:multi}.}
+   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order. \textcolor{blue}{An example of use can be found in Section \ref{sec:multi}.}
 \item \code{noise.number} indicates the number of sources of noise.
@@
 In order to show \textcolor{red}{how general the approach is} in the \pkg{yuima} package we present some examples.
-
-\subsection{Diffusion Processes}\label{sec:diff}
+\textcolor{blue}{Throughout this section we assume that all the stochastic differential equations exists while in Section \ref{sec5} we will give precise regularity conditions needed to have a properly defined statistical model.}
+
+\subsection{One Dimensional Diffusion Processes}\label{sec:diff}
 Assume that we want to describe the following stochastic differential equation
 $$\de X_t = -3 X_t \de t + \frac{1}{1+X_t^2}\de W_t.$$
-\textcolor{blue}{In the above $a(x) = -3 x$ and $b(x) = \frac{1}{1+x^2}$ according to the notaion of previous section}
+\textcolor{blue}{In the above $a(x) = -3 x$ and $b(x) = \frac{1}{1+x^2}$ according to the notaion of previous section and $W_t$ is a standard Wiener process. }
 This \textcolor{blue}{can be described} in \pkg{yuima} \textcolor{red}{by} specifying the drift and diffusion coefficients as plain 
@@
 @
-\textcolor{More details on how to change the default sampling scheme for the \code{simulate} method and how to perform subsampling will be given in Section \ref{sec:simul}.}
-
-\subsection{User specified state and time variables: NEW SECTION}
+\textcolor{blue}{More details on how to change the default sampling scheme for the \code{simulate} method and how to perform subsampling will be given in Section \ref{sec:simul}.}
+
+\subsection{User Specified State and Time Variables: NEW SECTION}
 Suppose now that the user wants to specify her own model using a prescribed notation, e.g., some sdes like
 $$\de Y_s = -3 s Y_s \de s + \frac{1}{1+Y_s^2}\de W_s,$$
-then this can be done in \pkg{yuima} as follows
+where $a(s,y) = -3sy$ and $b(y) = 1/(1+y^2)$.
+Then this model can be described in \pkg{yuima} as follows
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 @
+\textcolor{blue}{Looking at the structure of the \code{noise.number} slot in \code{mod3}, one can see that this is now set to 3 which is taken as the number of columns of the diffusion matrix.}
 Again, this model can be easily simulated
@@
 But it is also possible to specify more complex models like the following
-$$
+ \begin{equation} \label{sabr}
  \begin{cases}
- &\de X_t^1 = X_t^2 \left| X_t^1 \right|^{2/3} \de W_t^1, \\
- &\de X_t^2 = g(t)\de X_t^3, \\
- &\de X_t^3 = X_t^3( \mu  \de t + \sigma (\rho \de W_t^1 + \sqrt{1-\rho^2} \de W_t^2))
- \end{cases},
-$$
-where  $g(t) = 0.4 + (0.1 + 0.2t) e^{-2t}$.
+ & \mbox{d} X_{1,t} = X_{2,t} \left| X_{1,t} \right|^{2/3} \mbox{d}W_{1,t}, \\
+ & \mbox{d}X_{2,t} = g(t)\mbox{d}X_{3,t}, \\
+ & \mbox{d}X_{3,t} = X_{3,t}( \mu  \mbox{d}t + \sigma (\rho \mbox{d}W_{1,t} + \sqrt{1-\rho^2}
+   \mbox{d}W_{2,t}))
+ \end{cases}
+ \end{equation}
+$$ (X_{1,0}, X_{2,0}, X_{3,0}) = (1.0,0.1,1.0)$$
+ with $\mu = 0.1, \sigma = 0.2, \rho = -0.7 $ and $g(t) = 0.4 + (0.1 + 0.2t) e^{-2t}$,  for
+ example, where $W = (W_1, W_2)$ is a 2-dimensional standard Brownian motion. In order to pass this model to \pkg{yuima} we need to rewrite it in matrix form.
+ The solution  $X = (X_1, X_2, X_3)$ takes values on $\mathbb{R}_+^3$.
+ This is a stochastic integral equation defined as
+ \begin{equation}
+ X_t = X_0 +  \int_0^t a(s,X_s) \mbox{d}s + \int_0^t b(s,X_s) \mbox{d}W_s
+ \end{equation}
+ with 
+ \[
+  a(s,x) = \left( \begin{array}{@{\,}c@{\,}} 0 \\ g(s) \mu x_3 \\ \mu x_3 \end{array} \right),
+ \ \ 
+ b(s,x) = \left( \begin{array}{@{\,}cc@{\,}} x_2 | x_1|^{2/3} &  0 \\
+  g(s)\sigma \rho x_3 &  g(s)\sigma \sqrt{1-\rho^2} x_3 \\ 
+ \sigma \rho  x_3 & \sigma \sqrt{1-\rho^2} x_3 
+ \end{array} \right)
+ \]
+ for $ x = (x_1, x_2, x_3)$. 
+<<echo=TRUE, print=FALSE,results=hide>>=
+mu <- 0.1
+sig <- 0.2
+rho <- -0.7
+
+g <- function(t) {0.4 + (0.1 + 0.2*t)* exp(-2*t)}
+
+f1 <- function(t, x1, x2, x3) {
+    ret <- 0
+    if(x1 > 0 && x2 > 0) ret <- x2*exp(log(x1)*2/3)
+    return(ret)
+}
+
+f2 <- function(t, x1, x2, x3) {
+     ret <- 0
+     if(x3 > 0) ret <- rho*sig*x3
+     return(ret)
+}
+
+f3 <- function(t, x1, x2, x3) {
+     ret <- 0
+     if(x3 > 0) ret <- sqrt(1-rho^2)*sig*x3
+     return(ret)
+}
+
+diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3) * g(t)", 
+     "f2(t,x1,x2,x3)", "0", "f3(t,x1,x2,x3)*g(t)", "f3(t,x1,x2,x3)"),  3, 2)
+
+sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, state.variable = c("x1", "x2", "x3"), 
+    solve.variable = c("x1", "x2", "x3"))
+str(sabr.mod@parameter)
+@
+\textcolor{blue}{The functions \code{f1}, \code{f2} and \code{f3} are defined in a way that, when the trajectory of the processes crosso zero from above, the trajectory is stopped ad zero.
+Notice that in this way the only visible parameter for \pkg{yuima} is \code{mu} as \code{rho} and \code{sig} are inside the bodies of the functions \code{f2} and \code{f3}.
+If we want to instruct \pkg{yuima} about these parameters, they should appear explicitly as arguments of the functions as explained by this \proglang{R} code}
+<<echo=TRUE, print=FALSE,results=hide>>=
+ f2 <- function(t, x1, x2, x3, rho, sig) {
+     ret <- 0
+     if(x3 > 0) ret <- rho*sig*x3
+     return(ret)
+ }
+
+ f3 <- function(t, x1, x2, x3, rho, sig) {
+     ret <- 0
+     if(x3 > 0) ret <- sqrt(1-rho^2)*sig*x3
+     return(ret)
+ }
+
+ diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3,rho, sig) * g(t)", 
+     "f2(t,x1,x2,x3,rho,sig)", "0", "f3(t,x1,x2,x3,rho,sig)*g(t)", "f3(t,x1,x2,x3,rho,sig)"),  3, 2)
+
+ sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, 
+                       state.variable = c("x1", "x2", "x3"), solve.variable = c("x1", "x2", "x3"))
+str(sabr.mod@parameter)
+@
 
 \section{Fractional Gaussian Noise}\label{sec:fgn}
+The \pkg{yuima} allows for the description of stochastic differential equations driven by fractional Brownian motion of the following type
+$$\de X_t =  a(X_t) \de t + b(X_t) \de W_t^H
+$$
+where $W^H=\left( W^H_t, t\geq 0 \right)$ is a normalized fractional Brownian motion (fBM), {\it i.e.} the zero mean Gaussian processes with covariance function
+$$E( W_s^H W_t^H )= \frac{1}{2} \left( |s|^{2H}+|t|^{2H}- |t-s|^{2H}\right)$$ with Hurst exponent $H\in(0,1)$.
+The fractional Brownian motion process is  neither Markovian nor a semimartingale for $H\neq\frac{1}{2}$ but
+remains Gaussian.
+For $H > \frac12$, the solution $X_t$ above  presents the long-range dependance property that makes it useful for different applications in biology, physics, ethernet traffic or finance.
 In order to specify a stochastic differential equation driven by fractional Gaussian noise it is necessary to specify the value of the Hurst parameter.
@@
 we proceed as follows
-<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
-mod4 <- setModel(drift="3*y", diffusion=1, hurst=0.3, solve.var="y")
+<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=6,results=hide>>=
+mod4A <- setModel(drift="3*y", diffusion=1, hurst=0.3, solve.var="y")
+mod4B <- setModel(drift="3*y", diffusion=1, hurst=0.7, solve.var="y")
 set.seed(123)
-X <- simulate(mod4,sampling=setSampling(n=1000))
-plot(X)
+X1 <- simulate(mod4A,sampling=setSampling(n=1000))
+X2 <- simulate(mod4B,sampling=setSampling(n=1000))
+par(mfrow=c(2,1))
+par(mar=c(2,3,1,1))
+plot(X1,main="H=0.3")
+plot(X2,main="H=0.7")
 @
@@
 <<>>=
-str(mod4)
+str(mod4A)
 @
 The user can choose \textcolor{red}{between  two} simulation schemes, namely the Cholesky method and \citet{WoodChan} method.
+This is done via the argument \code{methodfGn} in the \code{simulate} method.
+The default simulation scheme is Wood and Chan and it is chosen by setting \code{methodfGn="WoodChan"}, the other simply by setting \code{methodfGn} to \code{Cholesky}.
 
@@
 \section{Asymptotic Expansion}\label{sec4}
-The \pkg{yuima} package can handle asymptotic expansion of functionals of  $d$-dimensional diffusion process 
-$$\de X_t^\ve = a(X_t^\ve,\ve)\de t + b(X_t^\ve,\ve)\de W_t, \qquad \ve \in(0,1]$$
-with $W_t$ \textcolor{red}{an} $r$-dimensional Wiener process, i.e., $W_t=(W_t^1, \ldots, W_t^r)$.
+{\colord For numerical computation of the expectation of a functional, 
+the Monte-Carlo methods universally gives a solution although they are time consuming and involve stochastic error 
+depending on the number of repetitions. 
+An alternative to the Monte-Carlo method is the method of asymptotic expansion, that can often give a solution with comparable accuracy, 
+besides, it is much superior in the computation time because of the approximation by an analytic formula. 
+ }
+
+The \pkg{yuima} package {\colord provides numerical approximation to the expectation of 
+a functional of a diffusion process in terms of the asymptotic expansion scheme. 
+Let us consider a family of}
+ $d$-dimensional diffusion processes {\colord $X=(X^{(\ep)}_t)_{t\in[0,T]}$ ($\ep\in(0,1]$) specified by the stochastic differential equasions}
+$$\de X_t^{(\ve)} = a(X_t^{(\ve)},\ve)\de t + b(X_t^{(\ve)},\ve)\de W_t, \qquad \ve \in(0,1]$$
+{\colord where}  $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process, i.e., $W_t=(W_{1,t}, \ldots, W_{r,t})$.
 The functional is expressed in the following abstract form
@@
 \label{yuima:func1}
-F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve), \qquad W_t^0=t.
+F^{(\ve)}  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^{(\ve)},\de)\de W_t^\alpha + F(X_t^{(\ve)},\ve), \qquad W_t^0=t.
 \end{equation} 
@@
 \label{eqyuima:gbm1}
-\de X_t^\ve = \mu X_t^\ve\de t + \ve X_t^\ve \de W_t
+\de X_t^{(\ve)} = \mu X_t^{(\ve)}\de t + \ve X_t^{(\ve)} \de W_t
 \end{equation}
 the price of the option is of the form
-$$ \dE\left\{ \max\left( \frac{1}{T}\int_0^T X_t^\ve \de t - K,0\right)\right\}.$$
+$$ \dE\left\{ \max\left( \frac{1}{T}\int_0^T X_t^{(\ve)} \de t - K,0\right)\right\}.$$
 Thus the functional of interest is
-$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$ 
+$$F^{(\ve)} = \frac{1}{T}\int_0^T X_t^{(\ve)} \de t, \qquad r=1$$ 
 with
@@
 So, the call option price requires the composition  of a smooth functional
-$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$
+$$F^\ve(X_t^{(\ve)}) = \frac{1}{T}\int_0^T X_t^{(\ve)} \de t, \qquad r=1$$
 with the irregular function
@@
 @
-Then, to obtain the first term in the asymptotic expansion (i.e., for $\ve=0$), it is as easy as calling the function \code{F0} on the \code{yuima} object:
+Then, to obtain the first term in the asymptotic expansion (i.e., for $\ve=0$), it is as easy as calling the function \code{F0} on the 
 <<echo=TRUE, print=FALSE>>=
@@
 The \pkg{yuima} implements several optimal techniques for parametric, semi- and non-parametric estimation of (multidimensional) stochastic differential equations.
+
+\subsection{How to input data into a \code{yuima} object}
 Although most of the examples in this section are given on simulated data, the main way to fill up the \code{data} slot of a \code{yuima} object is to use the function \code{setYuima}.
@@
 In particular, to estimate a \code{yuima.model} called \code{mod} on the data \code{X} one can use a code like this \code{my.yuima <- setYuima(data=setData(X), model=mod)} and then pass \code{my.yuima} to the inference functions as described in what follows.
+
+For example, assuming that an Internet connection is available, the following simple list of commands downloads data from the internet and constructs a \code{yuima} object with the \code{data} slot containing the time series.
+<<echo=TRUE, print=FALSE,results=hide,tidy=TRUE>>=
+data <- read.csv("http://chart.yahoo.com/table.csv?s=IBM&a=1&b=01&c=2012&d=1&e=01&f=2013&g=d&x=.csv")
+x <- setYuima(data=setData(data$Close))
+str(x@data)
+@
 

### 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@221 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@223 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[utf8x]{inputenc}
+
+\usepackage{comment}
+\excludecomment{en-text}
+
+\usepackage{enumerate}%\begin{enumerate}[{(}i{)}]
+\usepackage{bm}
+\usepackage{bbm} %\mathbbm{ABCabc\nabla}
+\usepackage[bbgreekl]{mathbbol} %\mathbb{abcABC\nabla\bbalpha\bbbeta\Delta  }
+\usepackage{natbib}
+\usepackage{dsfont}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{latexsym}
+\usepackage{amsmath}
 
@@
 a fractional Brownian motion if Hurst index $H=\frac12$.}
-Now, with \code{mod1} in hands, it is very easy to simulate a trajectory \textcolor{blue}{Euler-Maruyama scheme} of the process as follows
+Now, with \code{mod1} in hands, it is very easy to simulate a trajectory by  \textcolor{blue}{Euler-Maruyama scheme} of the process as follows
 <<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
@@
 \left(\begin{array}{c} -3 X_{1,t} \\  -X_{1,t} - 2X_{2,t}\end{array}\right)\de t +
-\left[\begin{array}{ccc}1 & 0 & X_{2,t} \\X_{1,t} & 3 & 0\end{array}\right]
+\left(\begin{array}{ccc}1 & 0 & X_{2,t} \\X_{1,t} & 3 & 0\end{array}\right)
 \left(\begin{array}{c}\de W_{t,1} \\ \de W_{t,2} \\ \de W_{t,3}\end{array}\right)
@@
 
-\section{Fractional Gaussian Noise}\label{sec:fgn}
+\subsection{Fractional Gaussian Noise}\label{sec:fgn}
 The \pkg{yuima} allows for the description of stochastic differential equations driven by fractional Brownian motion of the following type
@@
 where $W^H=\left( W^H_t, t\geq 0 \right)$ is a normalized fractional Brownian motion (fBM), {\it i.e.} the zero mean Gaussian processes with covariance function
-$$E( W_s^H W_t^H )= \frac{1}{2} \left( |s|^{2H}+|t|^{2H}- |t-s|^{2H}\right)$$ with Hurst exponent $H\in(0,1)$.
+$$\dE( W_s^H W_t^H )= \frac{1}{2} \left[ |s|^{2H}+|t|^{2H}- |t-s|^{2H}\right]$$ with Hurst exponent $H\in(0,1)$.
 The fractional Brownian motion process is  neither Markovian nor a semimartingale for $H\neq\frac{1}{2}$ but
@@
 \section{Asymptotic Expansion}\label{sec4}
-{\colord For numerical computation of the expectation of a functional, 
-the Monte-Carlo methods universally gives a solution although they are time consuming and involve stochastic error 
+{\colorb For numerical computation of the expectation of a random variable,
+the Monte-Carlo method gives a universal solution although it is time consuming and involves 
+ stochastic errors of certain scale 
 depending on the number of repetitions. 
-An alternative to the Monte-Carlo method is the method of asymptotic expansion, that can often give a solution with comparable accuracy, 
-besides, it is much superior in the computation time because of the approximation by an analytic formula. 
+An alternative method is the asymptotic expansion, that can often give a solution with accuracy comparable or superior to  
+Monte-Carlo approaches, 
+besides, it has advantage in the computation time because of the approximation by an analytic formula. 
  }
 
-The \pkg{yuima} package {\colord provides numerical approximation to the expectation of 
-a functional of a diffusion process in terms of the asymptotic expansion scheme. 
-Let us consider a family of}
- $d$-dimensional diffusion processes {\colord $X=(X^{(\ep)}_t)_{t\in[0,T]}$ ($\ep\in(0,1]$) specified by the stochastic differential equasions}
-$$\de X_t^{(\ve)} = a(X_t^{(\ve)},\ve)\de t + b(X_t^{(\ve)},\ve)\de W_t, \qquad \ve \in(0,1]$$
-{\colord where}  $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process, i.e., $W_t=(W_{1,t}, \ldots, W_{r,t})$.
-The functional is expressed in the following abstract form
+%The \pkg{yuima} 
+%package {\colorb provides numerical approximation to the expectation of 
+%a functional of a diffusion process in terms of the asymptotic expansion scheme. }
+Let us consider a family of 
+ $d$-dimensional diffusion processes {\colorb $X=(X^{(\ve)}_t)_{t\in[0,T]}$ ($\ve\in(0,1]$) specified 
+ by the stochastic integral equation}
+\begin{equation}\label{asy01}
+X_t^{(\ve)} = x_0+\int_0^t a(X_s^{(\ve)},\ve)\de s + \int_0^t b(X_s^{(\ve)},\ve)\de W_s, \qquad t\in[0,T]
+\end{equation}
+{\colorb for $\ve \in(0,1]$, 
+where}  $W_t=(W_{1,t}, \ldots, W_{r,t})$ \textcolor{red}{is an} $r$-dimensional Wiener process.
+The functional of interest is expressed in the following abstract form
 \begin{equation}
 \label{yuima:func1}
-F^{(\ve)}  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^{(\ve)},\de)\de W_t^\alpha + F(X_t^{(\ve)},\ve), \qquad W_t^0=t.
-\end{equation} 
-A typical example of application is the case of Asian option pricing. For example, in the Black \& Scholes model
+F^{(\ve)}  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^{(\ve)},\ve)\de W_t^\alpha + F(X_T^{(\ve)},\ve), \qquad W_t^0=t.
+\end{equation}
+
+A typical application is the Asian option pricing. For example, in the Black-Scholes model
 \begin{equation}
 \label{eqyuima:gbm1}
-\de X_t^{(\ve)} = \mu X_t^{(\ve)}\de t + \ve X_t^{(\ve)} \de W_t
-\end{equation}
-the price of the option is of the form
-$$ \dE\left\{ \max\left( \frac{1}{T}\int_0^T X_t^{(\ve)} \de t - K,0\right)\right\}.$$
+\de X_t^{(\ve)} = \mu X_t^{(\ve)}\de t + \ve X_t^{(\ve)} \de W_t, 
+\end{equation}
+the price of the option under zero interest rate is of the form
+$$ \dE\left[\max\left( \frac{1}{T}\int_0^T X_t^{(\ve)} \de t - K,0\right)\right].$$
+%
 Thus the functional of interest is
@@
 in \eqref{yuima:func1}.
-So, the call option price requires the composition  of a smooth functional
-$$F^\ve(X_t^{(\ve)}) = \frac{1}{T}\int_0^T X_t^{(\ve)} \de t, \qquad r=1$$
-with the irregular function
-$$
-\max(x-K,0)
-$$
+%
+%\begin{en-text}%%%%%%%%%%%
+%So, the call option price requires the composition  of a smooth functional
+%$$F^\ve(X_t^{(\ve)}) = \frac{1}{T}\int_0^T X_t^{(\ve)} \de t, \qquad r=1$$
+%with the irregular function
+%$$
+%\max(x-K,0)
+%$$
+%\end{en-text}%%%%%%%%%%
+%
+{\colorb Similarly, for $F(x,\ve)=x$, 
+the functional becomes $F^{(\ve)}=X^{(\ve)}_T$ and the price of the European call option is 
+$\dE[\max(X^{(\ve)}_T-K,0)]$. This value has a closed form in the Black-Sholes economy but 
+it is necessary to apply some numerical method for pricing the Asian option even in this linear case. }
+
+{\colorb 
+Returning to the general system (\ref{asy01})-(\ref{yuima:func1}), 
+we will assume that the stochastic system is deterministic in the limit as $\ep\downarrow0$, 
+that is,  
+$$ 
+b(\cdot,0)\>=\>0 \quad \mbox{and}\quad  f_\alpha(\cdot,0)=0\quad (\alpha=1,...,r).
+$$ 
+%
+Since $X^{(0)}_t$ is deterministic solution to the ordinary differential equation 
+$$
+\de X^{(0)}_t/\de t =a(X^{(0)}_t,0),\quad X^{(0)}_0=x_0,
+$$
+the functional $F^{(0)}$ becomes a constant given by 
+\begin{equation} \label{130106-1}
+F^{(0)} = \int_0^T f_0(X^{(0)}_t,0)\de t+F(X^{(0)}_T,0).
+\end{equation}
+
+Under standard regularity of $a$, $b$, $f_\alpha$ and $F$, 
+it is possible to show $F^{(\ep)}$ has a version that is smooth in $\ep\in[0,1)$ almost surely, and hence, 
+$$
+\tilde{F}^{(\ep)} := \ep^{-1}(F^{(\ep)}-F^{(0)})
+$$
+admits a stochastic expansion
+\begin{equation}
+\label{130206-8}
+\tilde{F}^{(\ep)} \quad \sim \quad
+\tilde{F}^{[0]}+\ep\tilde{F}^{[1]}+\ep^2\tilde{F}^{[2]}+\cdots
+\end{equation}
+as $\ep\downarrow0$. 
+This stochastic expansion makes sense in the Sobolev spaces of the Malliavin calculus. 
+Then the so-called Watanabe's theory (\cite{Watanabe87}) validates 
+the asymptotic expansion of the (generalized) expectation 
+\begin{equation} \label{130206-2}
+\dE[g(\tilde{F}^{(\ep)})] \quad \sim \quad d_0(g)
++\ep d_1(g)+\ep^2 d_2(g)+\cdots
+\end{equation} 
+as $\ep\downarrow0$ for measurable functions $g$ at most polynomial growth or 
+more generally for Schwartz distributions, 
+under the uniform nondegeneracy of the Malliavain covariance of $\tilde{F}^{(\ep)}$.
+\footnote{This condition ensures smoothness of the distribution of $\tilde{F}^{(\ep)}$. 
+It should be remarked that the Watanabe's theory is more general than the present use 
+for the variable $\tilde{F}^{(\ep)}$ having a Gaussian principal part $\tilde{F}^{(0)}$. }
+%
+In the present situation, each $d_i(g)$ is expresses as 
+$$
+d_i(g) = \int g(z)p_i(z)\phi(z;0,v)dz,
+$$
+where $p_i$ is a polynomial and $\phi(z;0,v)$ is the density of the normal distribution $N(0,v)$
+with $v=\mbox{Cov}[\tilde{F}^{(0)}]$. 
+Polynomials $p_i$ are given by the conditional expectation of multiple Wiener integrals. 
+%This expansion formula is implemented in \pkg{yuima}. 
+%In finance, we can apply this expansion to $g(z)=\max(z-c,0)$, for example. 
+The expansion (\ref{130206-2}) holds uniformly in a class of functions $g$.
+
+As mentioned above, 
 Monte Carlo methods require a huge number of simulations to get the desired accuracy of the
-calculation of the price, while asymptotic expansion of $F^\ve$ provides very accurate approximations.
-The \pkg{yuima} package provides functions to construct the functional $F^\ve$, and automatic asymptotic expansion based on Malliavin calculus \citep{Yoshida92b} starting from a \pkg{yuima} object.
-Here is an example. Consider a simple geometric Brownian motion of equation \eqref{eqyuima:gbm1} with $\mu=1$ and $X_0=1$. We define the model and the functional below:
+calculation of the expectation, while the asymptotic expansion of $F^{(\ve)}$ gives very fast 
+and accurate approximation by analytic formulation.
+The \pkg{yuima}  package provides functions to construct the functional $F^{(\ve)}$, and automatic asymptotic expansion based on the Malliavin calculus.  starting from a  \pkg{yuima} object. 
+This asymptotic expansion approach to option pricing was proposed in \citep{Yoshida92b} 
+and many related papers are available today.
+
+Though our method can be applied to the nonlinear system (\ref{asy01})-(\ref{yuima:func1}), 
+just for an example, we shall consider 
+the Asian call option of the geometric Brownian motion of equation \eqref{eqyuima:gbm1} with $\mu=1$ and $x_0=1$, 
+and 
+\begin{equation} \label{130206-5} 
+g(x)= %F^{(0)}-K+\ep x\>=\>\ep\left(
+\max\left(F^{(0)}-K+\ep x,0\right)%\right). 
+\end{equation} 
+Set the model  (\ref{asy01}) and the functional (\ref{yuima:func1}) as follows:
 <<echo=TRUE, print=FALSE, results=hide>>=
@@
 @
-this time the \code{setFunctional} command fills the appropriate slots inside the \code{yuima} object
+This time the \code{setFunctional} command fills the appropriate slots inside the \code{yuima} object
 <<echo=TRUE, print=FALSE>>=
@@
 @
-Then, to obtain the first term in the asymptotic expansion (i.e., for $\ve=0$), it is as easy as calling the function \code{F0} on the 
+Then the limit $F^{(0)}$ of $F^{(\ep)}$ is easily obtained by calling the function \code{F0} on the \code{yuima} object:
 <<echo=TRUE, print=FALSE>>=
@@
 @
-so the option price approximation is
-<<>>=
-max(F0-K,0)  # asian call option price
-@
-We can go up to the first order approximation adding one term to the expansion
+Set the function $g$ according to (\ref{130206-5}): 
+%<<>>=
+%max(F0-K,0)  # asian call option price
+%@
+%We can go up to the first order approximation adding one term to the expansion
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 }
+@
+Now we are at the point of computing the coefficients $d_i$ ($i=0,1,2$) in the expansion of 
+the price $\dE[\max(F^{(\ep)}-K,0)]$ 
+by applying the function \code{asymptotic_term}: 
+<<echo=TRUE, print=FALSE,results=hide>>=
 asymp <- asymptotic_term(yuima, block=10, rho, g)
-@
-and the final value is
+asymp
+@
+Then the sums 
 <<echo=TRUE, print=FALSE>>=
-asymp$d0 + e * asymp$d1  # asymp. exp. of asian call price
-@
+asymp$d0 + e * asymp$d1  # 1st order asymp. exp. of asian call price
+asymp$d0 + e * asymp$d1+ e^2* asymp$d2  # 2nd order asymp. exp. of asian call price
+@
+give the first and second order asymptotic expansions, respectively.
+
+{\coloro koko MC here!! Why don't we make asyexp = \code{asymptotic_term} + sums? 
+Also Monte-Carlo driver montecarlo = \code{simulate} with number of repetition 
++ plot + comparison of plot and theoretical curve?}
+
+We remark that the expansion of $\dE[g(\tilde{F}^{(\ep)})G^{(\ep)}]$ is also possible by the same method 
+for a functional $G^{(\ep)}$ having a stochastic expansion like (\ref{130206-8}). 
+Thus our method works even under the existence of a stochastic discount factor.
 
@@
 \subsection{Quasi Maximum Likelihood Estimation}
+{\bf[please remember: $\hat\theta$ = MLE;  $\tilde \theta$ = BAYES; $\check\theta$ = LASSO]}
+\par
 Consider the multidimensional diffusion process
 $$
-\de X_t = b(\theta_2,X_t)\de t + \sigma(\theta_1, X_t) \de W_t
+\de X_t = a(X_t,\theta_2)\de t + b(X_t,\theta_1) \de W_t
 $$
@@
 +\frac{1}{\Delta_n}
-\Sigma_{i-1}^{-1}(\theta_1)[\Delta X_i-\Delta_n b_{i-1}(\theta_2)]^{\otimes 2}\right\}
+\Sigma_{i-1}^{-1}(\theta_1)[\Delta X_i-\Delta_n a_{i-1}(\theta_2)]^{\otimes 2}\right\}
 \end{eqnarray}}
-where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, \textcolor{red}{$A^{\otimes 2}= A A^T$} and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
+where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $a_i(\theta_2)=a(X_{t_i},\theta_2)$, $\Sigma=b^{\otimes 2}$, \textcolor{red}{$A^{\otimes 2}= A A^T$} and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
 Then, \citep[see e.g.,][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
@@
 
+\subsubsection{QMLE in practice}
+
 As an example, we consider the simple model
@@
 \begin{equation} 
-\de X_t=b(X_t,\theta_2)\de t+\sigma(X_t,\theta_1)\de W_t,
+\de X_t=a(X_t,\theta_2)\de t+b(X_t,\theta_1)\de W_t,
 \end{equation}
@@
 %integration, otherwise MCMC method is used.
-\subsubsection{Small Sample Size}
+\subsubsection{The Effect of Small Sample Size in Drift Estimation}
 It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
@@
 Let $T \in (0,\infty)$ be a terminal time for possible observations. 
-We consider a two dimensional It\^o process $(X^1,X^2)$  
+We consider a two dimensional It\^o process $(X_1,X_2)$  
 satisfying the stochastic differential equations
 \begin{eqnarray*}
-\mbox{d}X_t^l &=&\mu_t^l \mbox{d}t + \sigma_t^l \mbox{d}W_t^l ,\quad
+\mbox{d}X_{l,t} &=&\mu_{l,t} \mbox{d}t + \sigma_{l,t} \mbox{d}W_{l,t} ,\quad
 t\in[0,T]\\
-X_0^l &=& x_0^l        \notag
+X_{l,0} &=& x_{l,0}        \notag
 \end{eqnarray*}
 for $l=1,2$. 
-Here $W^l$ denote standard Wiener processes with a 
+Here $W_l$ denote standard Wiener processes with a 
 progressively measurable correlation process 
 $\mbox{d}\langle W_1 , W_2 \rangle_t = \rho_t \mbox{dt}$,  
-$\mu^l_t$ and $\sigma_t^l$ are progressively measurable processes, 
-and $x^l_0$ are initial random variables independent of $(W^1,W^2)$.  
+$\mu_{l,t}$ and $\sigma_{l,t}$ are progressively measurable processes, 
+and $x_{l,0}$ are initial random variables independent of $(W_1,W_2)$.  
 Diffusion type processes are in the scope but this model 
@@
 
-The process $X^l$ is supposed to be observed at 
+The process $X_l$ is supposed to be observed at 
 over the increasing sequence of times 
 $T^{l,i}$ $(i\in\mathbb{Z}_{\geq0})$ starting at $0$, up to time T. 
-Thus, the observables are $(T^{l,i},X^{l,i})$ with $T^{l,i}\leq T$. 
+Thus, the observables are $(T^{l,i},X_{l,i})$ with $T^{l,i}\leq T$. 
 Each $T^{l,j}$ may be a stopping time, so possibly depends on 
-the history of $(X^1,X^2)$ as well as the precedent stopping times.  
+the history of $(X_1,X_2)$ as well as the precedent stopping times.  
 Two sequences of stopping times 
@@
 
-The parameter of interest is the quadratic covariation between $X^1$ and $X^2$:
+The parameter of interest is the quadratic covariation between $X_1$ and $X_2$:
 
@@
 \theta=
-\langle X^1 , X^2 \rangle_T = \int_0^T \sigma_t^1 \sigma_t^2 \rho_t \mbox{d}t.
+\langle X_1 , X_2 \rangle_T = \int_0^T \sigma_{1,t} \sigma_{2,t} \rho_t \mbox{d}t.
 \end{equation}
@@
 \begin{equation}
-U_n= \sum_{i,j:T^{1,i}\leq T, T^{2,j}\leq T} (X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)
+U_n= \sum_{i,j:T^{1,i}\leq T, T^{2,j}\leq T} (X_{1,T^{1,i}}-X_{1,T^{1,i-1}})(X_{2,T^{2,j}}-X_{2,T^{2,j-1}})
 1_{\{ (T^{1,i-1},T^{1,i}] 
@@
 \end{equation}
-That is, the product of any pair of increments $(X_{T^{1,i}}^1-X_{T^{1,i-1}}^1)$ and
-$(X_{T^{2,j}}^2-X_{T^{2,j-1}}^2)$ will make a contribution 
+That is, the product of any pair of increments $(X_{1,T^{1,i}}-X_{1,T^{1,i-1}})$ and
+$(X_{2,T^{2,j}}-X_{2,T^{2,j-1}})$ will make a contribution 
 to the sum only when 
@@
 As an example, consider a two dimensional stochastic process 
-$(X^1_t,X^2_t)$ satisfying the stochastic differential equation
+$(X_{1,t},X_{2,t})$ satisfying the stochastic differential equation
 \begin{equation}
 \begin{split}
-\mbox{d}X^1_t = \sigma_{1,t} \mbox{d}B^1_t, \\
-\mbox{d}X^2_t = \sigma_{2,t} \mbox{d}B^2_t. 
+\mbox{d}X_{1,t} = \sigma_{1,t} \mbox{d}B_{1,t}, \\
+\mbox{d}X_{2,t} = \sigma_{2,t} \mbox{d}B_{2,t}. 
 \end{split}
 \end{equation}
-Here $B^1_t$ and $B^2_t$ denote two standard Wiener processes, 
+Here $B_{1,t}$ and $B_{2,t}$ denote two standard Wiener processes, 
 however they are correlated as 
 \begin{eqnarray}
-B^1_t &=& W^1_t,\\
-B^2_t &=&  \int_0^t \rho_s \mbox{d} W^1_s +
-  \int_0^t \sqrt{ 1- \rho_s^2} \mbox{d} W^2_s,
+B_{1,t} &=& W_{1,t},\\
+B_{2,t} &=&  \int_0^t \rho_s \mbox{d} W_{1,s} +
+  \int_0^t \sqrt{ 1- \rho_s^2} \mbox{d} W_{2,s},
 \end{eqnarray}
-where  $W^1_t$ and $W^2_t$ are independent Wiener processes,  and 
-$\rho_t$ is the correlation function between $B^1_t$ and $B^2_t$. 
+where  $W_{1,t}$ and $W_{2,t}$ are independent Wiener processes,  and 
+$\rho_t$ is the correlation function between $B_{1,t}$ and $B_{2,t}$. 
 We consider $\sigma_{l,t},l=1,2$ and $\rho_t$ 
@@
 \end{eqnarray*}
-To simulate the stochastic process $(X^1_t,X^2_t)$, we first build the model 
+To simulate the stochastic process $(X_{1,t},X_{2,t})$, we first build the model 
 by $\tt setModel$ as before. 
@@
 $$
-\de Y_t = b_t \de t + \sigma(X_t,\theta) \de W_t,\ \ t\in[0,T], 
-$$
-where $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process and $b_t$ and $X_t$ are multidimensional processes and $\sigma$ is the diffusion coefficient (volatility) matrix.
+\de Y_t = a_t \de t + b(X_t,\theta) \de W_t,\ \ t\in[0,T], 
+$$
+where $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process and $a_t$ and $X_t$ are multidimensional processes and $b$ is the diffusion coefficient (volatility) matrix.
 When $Y=X$ the problem is a diffusion model.
-The process $b_t$ may have jumps but should not explode and it is treated as a nuisance in this model.
+The process $a_t$ may have jumps but should not explode and it is treated as a nuisance in this model.
 The change-point problem for the volatility is formalized as follows
@@
 \begin{array}{ll}
-Y_0+\int_0^t b_s \de s+\int_0^t \sigma(X_s,\theta_0^*) \de W_s
+Y_0+\int_0^t a_s \de s+\int_0^t b(X_s,\theta_0^*) \de W_s
 & \mbox{ for } t\in[0,\tau^*)
 \\
-Y_{\tau^*}+\int_{\tau^*}^t b_s \de s+\int_{\tau^*}^t \sigma(X_s,\theta_1^*) \de W_s
+Y_{\tau^*}+\int_{\tau^*}^t a_s \de s+\int_{\tau^*}^t b(X_s,\theta_1^*) \de W_s
 & \mbox{ for } t\in[\tau^*,T]. 
@@
 
-\subsection{Example  of Volatility Change-Point Estimation}
+\subsubsection{Example  of Volatility Change-Point Estimation for 2-Dimensional SDE's}
 One example of model that can be analyzed by this technique is, for example, the 2-dimensional stochastic differential equation
@@
 \left(\begin{array}{c}
-\de X_t^1\\ \de X_t^2 
+\de X_{1,t}\\ \de X_{2,t} 
 \end{array}\right)
@@
 \left(\begin{array}{c}
-b_1(X^1_t) \\ 
-b_2(X^2_t) \\ 
+b_1(X_{1,t}) \\
+b_2(X_{2,t}) \\ 
 \end{array}\right) \de t
 +
-\left[\begin{array}{cc}
-\theta_{1.k}  \cdot  X_t^1&0  \cdot  X_t^1 \\ 
-0  \cdot  X_t^2&\theta_{2.k}  \cdot  X_t^2 \\ 
-\end{array}\right]
+\left(\begin{array}{cc}
+\theta_{1.k}  \cdot  X_{1,t}&0  \cdot  X_{1,t} \\ 
+0  \cdot  X_{2,t}&\theta_{2.k}  \cdot  X_{2,t} \\ 
+\end{array}\right)
 \left(\begin{array}{c}
-\de W_t^1\\ \de W_t^2 
-\end{array}\right)
+\de W_{1,t}\\ \de W_{2,t} 
+\end{array}\right),\quad \textcolor{red}{ t\in[0,T],}
 $$
@@
 \left(\begin{array}{c}
- \de X_t^1\\ \de X_t^2 
+ \de X_{1,t}\\ \de X_{2,t} 
  \end{array}\right)
@@
 \left(\begin{array}{c}
- \sin(X^1_t) \\ 
- 3 - X^2_t \\ 
+ \sin(X_{1,t}) \\
+ 3 - X_{2,t} \\ 
  \end{array}\right) \de t
  +
-\left[\begin{array}{cc}
- \theta_{1.k}  \cdot  X_t^1&0  \cdot  X_t^1 \\ 
- 0  \cdot  X_t^2&\theta_{2.k}  \cdot  X_t^2 \\ 
- \end{array}\right]
+\left(\begin{array}{cc}
+ \theta_{1.k}  \cdot  X_{1,t}&0  \cdot  X_{1,t} \\ 
+ 0  \cdot  X_{2,t}&\theta_{2.k}  \cdot  X_{2,t} \\ 
+ \end{array}\right)
 \left(\begin{array}{c}
- \de W_t^1\\ \de W_t^2 
+ \de W_{1,t}\\ \de W_{2,t} 
  \end{array}\right),\quad \textcolor{red}{ t\in[0,T],}
@@
 $$
-X^1_0=1.0,\quad
-X^2_0=1.0,\quad
+X_{1,0}=1.0,\quad
+X_{2,0}=1.0,\quad
 $$
@@
 $$
-\de X_t = b(\alpha, X_t) \de t + \sigma(\beta,X_t)  \de W_t
+\de X_t = a( X_t,\alpha) \de t + b(X_t,\beta)  \de W_t
 $$
@@
 with $b:\Theta_p\times\mathbb{R}^d\to \mathbb{R}^d$, $\sigma:\Theta_q\times \mathbb{R}^d\to \mathbb{R}^d\times \mathbb{R}^m$ and $W_t,t\in [0,T]$, is a standard Brownian motion in $\mathbb{R}^m$. 
-We assume that the functions $b$ and $\sigma$ are known up to   $\alpha$ and $\beta$.
+We assume that the functions $a$ and $b$ are known up to   $\alpha$ and $\beta$.
 We denote by $\theta=(\alpha,\beta)\in\Theta_p\times \Theta_q=\Theta$ the parametric vector and with $\theta_0=(\alpha_0,\beta_0)$ its unknown true value. 
@@
 
-\subsection{An Example of Use}
+\subsubsection{An Example of Model Selection for Interest Rates Data}
 The \code{lasso} method is implemented in the \pkg{yuima} package.

### 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
various update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@224 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-{\coloro koko MC here!! Why don't we make asyexp = \code{asymptotic_term} + sums? 
-Also Monte-Carlo driver montecarlo = \code{simulate} with number of repetition 
-+ plot + comparison of plot and theoretical curve?}
-
 We remark that the expansion of $\dE[g(\tilde{F}^{(\ep)})G^{(\ep)}]$ is also possible by the same method 
@@
 \subsection{Quasi Maximum Likelihood Estimation}
-{\bf[please remember: $\hat\theta$ = MLE;  $\tilde \theta$ = BAYES; $\check\theta$ = LASSO]}
-\par
+{\bf ADD SOMETHING ON ASYMPT DISTRIBUTION \& SAMPLING SCHEME} 
 Consider the multidimensional diffusion process
-$$
+\begin{equation}
+\label{eq:sdemle}
 \de X_t = a(X_t,\theta_2)\de t + b(X_t,\theta_1) \de W_t
-$$
+\end{equation}
 where $W_t$ is an {$r$}-dimensional standard Wiener process 
 independent of the initial value $X_0=x_0$. 
+Moreover, $\theta_1\in\Theta_1\subset \mathbb{R}^p, p\geq1$,
+$\theta_2\in\Theta_2\subset \mathbb{R}^q, q\geq1$,
+$a:\mathbb{R}^d\times\Theta_2 \to \mathbb{R}^d$ and $b:\mathbb{R}^d\times\Theta_1 \to \mathbb{R}^d\times \mathbb{R}^r$.
+
 Quasi-MLE assumes the following approximation of the true log-likelihood for multidimensional diffusions
@@
 Then, \citep[see e.g.,][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
-\textcolor{red}{$$\hat\theta_n=\arg\max_\theta \ell_n({\bf X}_n,\theta)$$}
+\textcolor{red}{$$\hat\theta=\arg\max_\theta \ell_n({\bf X}_n,\theta)$$}
 
@@
 \subsection{Adaptive Bayes Estimation}
-Consider again the  diffusion process solution to
-\begin{equation} 
-\de X_t=a(X_t,\theta_2)\de t+b(X_t,\theta_1)\de W_t,
-\end{equation}
+Consider again the  diffusion process solution to \eqref{eq:sdemle}
 and the quasi likelihood defined in \eqref{qlik}.
@@
 $$
-where $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process and $a_t$ and $X_t$ are multidimensional processes and $b$ is the diffusion coefficient (volatility) matrix.
+where $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process and $a_t$ and $X_t$ are multidimensional processes, $\theta\in\Theta\subset \mathbb{R}^p$,  $b:\mathbb{R}^d\times\Theta \to \mathbb{R}^d\times \mathbb{R}^r$, is the diffusion coefficient (volatility) matrix.
+
 When $Y=X$ the problem is a diffusion model.
@@
 \end{equation}
-\textcolor{red}{and $S=\sigma^{\otimes 2}$.}
+\textcolor{red}{and $S=b^{\otimes 2}$.}
 Suppose that there exists an estimator 
@@
 \begin{eqnarray*} 
-\hat{\tau}_n&=&\arg\!\!\min\limits_{t\in[0,T]} 
+\hat{\tau}&=&\arg\!\!\min\limits_{t\in[0,T]} 
 \Phi_n(t;\hat{\theta}_0,\hat{\theta}_1).
@@
 \left(\begin{array}{c}
-b_1(X_{1,t}) \\
-b_2(X_{2,t}) \\ 
+a_1(X_{1,t}) \\
+a_2(X_{2,t}) \\
 \end{array}\right) \de t
@@
 Let $\mathbb{H}_n({\bf X}_n,\theta) = \ell_n({\bf X}_n,\theta)$ from equation \eqref{qlik}.
-The quasi-MLE $\tilde{\theta}_n$ for this model is the solution of the following problem
-$$
-\tilde{\theta}_n=(\tilde\alpha_n,\tilde\beta_n)'=\arg\min_\theta \mathbb{H}_n({\bf X}_n,\theta)
+The quasi-MLE $\hat{\theta}$ for this model is the solution of the following problem
+$$
+\hat{\theta}=(\hat\alpha,\hat\beta)'=\arg\min_\theta \mathbb{H}_n({\bf X}_n,\theta)
 $$ 
@@
 $$
-\hat{\theta}_n=(\hat\alpha_n,\hat\beta_n)=\arg\min_\theta\mathcal{F}(\theta).
+\check{\theta}=(\check\alpha,\check\beta)=\arg\min_\theta\mathcal{F}(\theta).
 $$
@@
 $$
-\mathcal{F}(\theta)=(\theta-\tilde{\theta}_n)^T\ddot{\mathbb{H}}_n({\bf X}_n, \tilde\theta_n)(\theta-\tilde{\theta}_n)+\sum_{j=1}^p\lambda_{n,j}|\alpha_j| +\sum_{k=1}^q\gamma_{n,k}|\beta_k|
+\mathcal{F}(\theta)=(\theta-\hat{\theta})^T\ddot{\mathbb{H}}_n({\bf X}_n, \hat\theta)(\theta-\hat{\theta})+\sum_{j=1}^p\lambda_{n,j}|\alpha_j| +\sum_{k=1}^q\gamma_{n,k}|\beta_k|
 $$
@@
 \label{eq:penalty}
-\lambda_{n,j} = \lambda_0 |\tilde \alpha_{n,j}|^{-\delta_1}, \qquad
-\gamma_{n,k} = \gamma_0 |\tilde \beta_{n,j}|^{-\delta_2}
-\end{equation}
-where $\tilde \alpha_{n,j}$ and  $\tilde \beta_{n,k}$ are the unpenalized QML estimator of $\alpha_j$ and $\beta_k$ respectively, $\delta_1, \delta_2>0$ and usually taken unitary.
+\lambda_{n,j} = \lambda_0 |\hat \alpha_{j}|^{-\delta_1}, \qquad
+\gamma_{n,k} = \gamma_0 |\hat \beta_{j}|^{-\delta_2}
+\end{equation}
+where $\hat \alpha_{j}$ and  $\hat \beta_{k}$ are the unpenalized QML estimator of $\alpha_j$ and $\beta_k$ respectively, $\delta_1, \delta_2>0$ and usually taken unitary.
 

### 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
various update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@225 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{Quasi Maximum Likelihood Estimation}
-{\bf ADD SOMETHING ON ASYMPT DISTRIBUTION \& SAMPLING SCHEME} 
+{\bf NAKAHIRO: ADD SOMETHING ON ASYMPT DISTRIBUTION \& SAMPLING SCHEME} \par
 Consider the multidimensional diffusion process
@@
 It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
-In our example above, we took $T=n^\frac13$, with $n = \Sexpr{n}$, which is approximatively  \Sexpr{round(n^(1/3),2)}.
-Now we reduce the sample size to $n=500$ and the value of $T$ is then $T=\Sexpr{round(500^(1/3),2)}$.
+In our example above, we took $T=n^\frac13$, with $n = \Sexpr{n}$, which is approximatively  \Sexpr{round(n^(1/3),2)}. Now we reduce the sample size to $n=500$ and then $T=\Sexpr{round(500^(1/3),2)}$.
 We then apply both quasi-maximum likelihood and adaptive Bayes type estimators to these data
@@
 \subsection{LASSO Model Selection}
+{\bf: STEFANO: EXPLAIN WHAT LASSO MEANS}\par
 Let $X_t$ be a  diffusion process solution to

### 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
various update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@226 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % USE THIS INSTEAD IF YOU WANT TO EXECUTE R CODE
-%\SweaveOpts{prefix.string=yuima, echo=TRUE, eval=TRUE}
+\SweaveOpts{prefix.string=yuima, echo=TRUE, eval=TRUE}
 
@@
 
+{\bf STEFANO: add comparison with other approx formulas}\par
+
 \section{Inference for Stochastic Processes}\label{sec5}
@@
 %integration, otherwise MCMC method is used.
-\subsubsection{The Effect of Small Sample Size in Drift Estimation}
+\subsubsection{The Effect of Small Sample Size on Drift Estimation}
 It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
@@
 \subsection{LASSO Model Selection}
-{\bf: STEFANO: EXPLAIN WHAT LASSO MEANS}\par
+{\bf STEFANO: EXPLAIN WHAT LASSO MEANS}\par
 Let $X_t$ be a  diffusion process solution to

### 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
various update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@227 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % USE THIS INSTEAD IF YOU WANT TO EXECUTE R CODE
-\SweaveOpts{prefix.string=yuima, echo=TRUE, eval=TRUE}
+%\SweaveOpts{prefix.string=yuima, echo=TRUE, eval=TRUE}
 
@@
 <<echo=TRUE, print=FALSE,results=hide>>=
-mod1b <- setModel(drift = "-3*s*y", diffusion = "1/(1+y^2)", state.var="y", time.var="s")
+mod1b <- setModel(drift = "-3*s*y", diffusion = "1/(1+y^2)", 
+state.var="y", time.var="s")
 @
@@
 
-diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3) * g(t)", 
-     "f2(t,x1,x2,x3)", "0", "f3(t,x1,x2,x3)*g(t)", "f3(t,x1,x2,x3)"),  3, 2)
-
-sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, state.variable = c("x1", "x2", "x3"), 
+diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)",
+ "f2(t,x1,x2,x3) * g(t)", "f2(t,x1,x2,x3)", "0", 
+     "f3(t,x1,x2,x3)*g(t)", "f3(t,x1,x2,x3)"),  3, 2)
+
+sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), 
+diffusion = diff.coef.matrix, state.variable = c("x1", "x2", "x3"),
     solve.variable = c("x1", "x2", "x3"))
@@
 
- diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)", "f2(t,x1,x2,x3,rho, sig) * g(t)", 
-     "f2(t,x1,x2,x3,rho,sig)", "0", "f3(t,x1,x2,x3,rho,sig)*g(t)", "f3(t,x1,x2,x3,rho,sig)"),  3, 2)
-
- sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), diffusion = diff.coef.matrix, 
-                       state.variable = c("x1", "x2", "x3"), solve.variable = c("x1", "x2", "x3"))
+ diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)", 
+ "f2(t,x1,x2,x3,rho, sig) * g(t)", "f2(t,x1,x2,x3,rho,sig)", 
+ "0", "f3(t,x1,x2,x3,rho,sig)*g(t)", "f3(t,x1,x2,x3,rho,sig)"),  3, 2)
+
+ sabr.mod <- setModel(drift = c("0", "mu*g(t)*x3", "mu*x3"), 
+ diffusion = diff.coef.matrix, state.variable = c("x1", "x2", "x3"), 
+ solve.variable = c("x1", "x2", "x3"))
 str(sabr.mod@parameter)
@@
 set.seed(123)
-Y.sub <- simulate(mymod,sampling=setSampling(delta=0.001,n=1000), subsampling=setSampling(delta=0.01,n=100))
+Y.sub <- simulate(mymod,sampling=setSampling(delta=0.001,n=1000),
+subsampling=setSampling(delta=0.01,n=100))
 set.seed(123)
@@
 <<echo=TRUE, print=FALSE,results=hide,tidy=TRUE>>=
-data <- read.csv("http://chart.yahoo.com/table.csv?s=IBM&a=1&b=01&c=2012&d=1&e=01&f=2013&g=d&x=.csv")
+data <- read.csv("http://chart.yahoo.com/table.csv?s=IBM&g=d&x=.csv")
 x <- setYuima(data=setData(data$Close))
@@
 set.seed(123)
-yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.2, theta2 = 0.3))
+yuima <- simulate(yuima, xinit = 1, 
+true.parameter = list(theta1 = 0.2, theta2 = 0.3))
 @
@@
 <<echo=TRUE>>=
-prior <- list(theta2=list(measure.type="code",df="dunif(theta2,0,1)"), theta1=list(measure.type="code",df="dunif(theta1,0,1)"))
+prior <- list(theta2=list(measure.type="code",df="dunif(theta2,0,1)"),
+ theta1=list(measure.type="code",df="dunif(theta1,0,1)"))
 @
@@
 set.seed(123)
-yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.2, theta2 = 0.3))
+yuima <- simulate(yuima, xinit = 1, 
+true.parameter = list(theta1 = 0.2, theta2 = 0.3))
 param.init <- list(theta2=0.5,theta1=0.5)
@@
 # Model sde using yuima.model
-cor.mod <- setModel(drift = c("",""), diffusion = diff.coef.matrix, solve.variable=c("x1","x2"))
+cor.mod <- setModel(drift = c("",""), diffusion = diff.coef.matrix,
+ solve.variable=c("x1","x2"))
 @
@@
 # Cumulative Covariance
-theta <- CC.theta(T=Terminal, sigma1=diff.coef.1, sigma2=diff.coef.2, rho=cor.rho)$value
+theta <- CC.theta(T=Terminal, sigma1=diff.coef.1, 
+sigma2=diff.coef.2, rho=cor.rho)$value
 cat(sprintf("theta=%5.3f\n",theta))
@@
 cor.rho <- function(t,x,y) 1/(1+x^2)
-diff.mat <- matrix(c("s1(t,x,y)", "s2(t,x,y) * cor.rho(t,x,y)","", "s2(t,x,y) * sqrt(1-cor.rho(t,x,y)^2)"), 2, 2) 
-cor.mod <- setModel(drift = c("b1","b2"), diffusion = diff.mat,solve.variable = c("x", "y"),state.var=c("x","y"))
+diff.mat <- matrix(c("s1(t,x,y)", "s2(t,x,y) * cor.rho(t,x,y)","",
+ "s2(t,x,y) * sqrt(1-cor.rho(t,x,y)^2)"), 2, 2) 
+cor.mod <- setModel(drift = c("b1","b2"), diffusion = diff.mat,
+solve.variable = c("x", "y"),state.var=c("x","y"))
 
@@
 drift.matrix <- matrix(drift.c, 2, 1)
-ymodel <- setModel(drift=drift.matrix, diffusion=diff.matrix, time.variable="t",
-state.variable=c("x1", "x2"), solve.variable=c("x1", "x2"))
+ymodel <- setModel(drift=drift.matrix, diffusion=diff.matrix, 
+time.variable="t", state.variable=c("x1", "x2"), 
+solve.variable=c("x1", "x2"))
 @
@@
 yuima <- yuima1
-yuima@data@zoo.data[[1]] <- c(yuima1@data@zoo.data[[1]], yuima2@data@zoo.data[[1]][-1])
-yuima@data@zoo.data[[2]] <- c(yuima1@data@zoo.data[[2]], yuima2@data@zoo.data[[2]][-1])
+yuima@data@zoo.data[[1]] <- c(yuima1@data@zoo.data[[1]], 
+yuima2@data@zoo.data[[1]][-1])
+yuima@data@zoo.data[[2]] <- c(yuima1@data@zoo.data[[2]], 
+yuima2@data@zoo.data[[2]][-1])
 @
@@
 <<cpoint4b>>=
-noDriftModel <- setModel(drift=c("0", "0"), diffusion=diff.matrix, time.variable="t",
-state.variable=c("x1", "x2"), solve.variable=c("x1", "x2"))
+noDriftModel <- setModel(drift=c("0", "0"), diffusion=diff.matrix,
+ time.variable="t", state.variable=c("x1", "x2"), 
+ solve.variable=c("x1", "x2"))
 noDriftModel <- setYuima(noDriftModel, data=yuima@data)
@@
 <<>>=
-qmleL(noDriftModel, t=2, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estL
-qmleR(noDriftModel, t=8, start=list(theta1.k=0.1, theta2.k=0.1),lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), method="L-BFGS-B") -> estR
+qmleL(noDriftModel, t=2, start=list(theta1.k=0.1, theta2.k=0.1),
+lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), 
+method="L-BFGS-B") -> estL
+qmleR(noDriftModel, t=8, start=list(theta1.k=0.1, theta2.k=0.1),
+lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), 
+method="L-BFGS-B") -> estR
 t0.est <- coef(estL)

### 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
various update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@228 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \subsection{LASSO Model Selection}
-{\bf STEFANO: EXPLAIN WHAT LASSO MEANS}\par
+The Least Absolute Shrinkage and Selection Operator (LASSO) is a useful and well studied approach to the problem of model selection and its major advantage is the simultaneous execution of both parameter estimation and variable selection (see \citet{Tib96}; \citet{Knight00}; \citet{Efron}).
+
+To simplify the idea: take a full specified regression model
+$$ Y = \theta_0 +  \theta_1 X_1+  \theta_2 X_2 + \cdots +  \theta_k X_k$$
+perform least squares estimation under $L_1$ constraints, i.e.
+
+$$\hat\theta = \arg\min_\theta \left\{(Y-\theta X)^T (Y-\theta X) + \sum_{i=1}^k |\theta_i|\right\}$$
+
+model selection occurs when some of the $\theta_i$ are estimated as zeros. The same idea can be applied to diffusion processes.
 Let $X_t$ be a  diffusion process solution to

### 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
various update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@229 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+<<>>=
+library(fExoticOptions)
+LevyAsianApproxOption(TypeFlag = "c", S = xinit, SA = xinit, X = K, 
+      Time = 1, time = 1, r = 0.0, b = 1, sigma = e)@price
+TurnbullWakemanAsianApproxOption(TypeFlag = "c", S = xinit, SA = xinit, 
+X = K, Time = 1, time = 1, r = 0.0, b = 1, sigma = e,tau=0)@price
+@
 {\bf STEFANO: add comparison with other approx formulas}\par

### 2013-02-08 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
last update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@230 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2013-02-07 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
@@
 T <- 1
-xinit <- 1
-K <- 1
+xinit <- 150
+K <- 100
 f <- list( expression(x/T), expression(0))
@@
 <<echo=TRUE, print=FALSE>>=
-asymp$d0 + e * asymp$d1  # 1st order asymp. exp. of asian call price
-asymp$d0 + e * asymp$d1+ e^2* asymp$d2  # 2nd order asymp. exp. of asian call price
+asy1 <- asymp$d0 + e * asymp$d1  # 1st order asymp. exp. of asian call price
+asy1
+asy2 <- asymp$d0 + e * asymp$d1+ e^2* asymp$d2  # 2nd order asymp. exp. of asian call price
+asy2
 @
@@
 Thus our method works even under the existence of a stochastic discount factor. 
-
-{\bf STEFANO: add comparison with other approx formulas}\par
+One can compare the result of the asymptotic expansion with other well known techniques like Edgeworth series expansion for the log-normal distribution as proposed, e.g.,  in \citet{Levy92}.
+This approximation is available through the package \pkg{fExoticOptions}.
+%%% https://www.rocq.inria.fr/mathfi/Premia/free-version/doc/premia-doc/pdf_html/asian_doc/index.html#x1-80005.1.1
+<<>>=
+library(fExoticOptions)
+levy <- LevyAsianApproxOption(TypeFlag = "c", S = xinit, SA = xinit, X = K, 
+      Time = 1, time = 1, r = 0.0, b = 1, sigma = e)@price
+levy
+@
+and the relative difference between the two approximations is \Sexpr{round((asy2-levy)/asy2*100,1)}\%.
+
+Of course, \pkg{yuima} approach is more general as Levy approximation only holds when the process $X_t$ is the geometric Brownian motion.
+
+\textcolor{red}{NAKAHIRO: ADD REMARK about 92 paper?}
 
@@
 The \pkg{yuima} implements several optimal techniques for parametric, semi- and non-parametric estimation of (multidimensional) stochastic differential equations.
+Most of the techniques presented below apply to high frequency data, i.e.
+when $\Delta_n$, the time lag between to consecutive observations of the process, converges to zero as  the number  $n$ of observations       increases.
 
@@
 \subsection{Quasi Maximum Likelihood Estimation}
-{\bf NAKAHIRO: ADD SOMETHING ON ASYMPT DISTRIBUTION \& SAMPLING SCHEME} \par
-Consider the multidimensional diffusion process
+Consider a multidimensional diffusion process
 \begin{equation}
 \label{eq:sdemle}
-\de X_t = a(X_t,\theta_2)\de t + b(X_t,\theta_1) \de W_t
+\de X_t = a(X_t,\theta_2)\de t + b(X_t,\theta_1) \de W_t,\quad X_0=x_0,
 \end{equation}
 where $W_t$ is an {$r$}-dimensional standard Wiener process 
-independent of the initial value $X_0=x_0$. 
-Moreover, $\theta_1\in\Theta_1\subset \mathbb{R}^p, p\geq1$,
-$\theta_2\in\Theta_2\subset \mathbb{R}^q, q\geq1$,
+independent of the initial variable $X_0$. 
+Moreover, $\theta_1\in\Theta_1\subset \mathbb{R}^p$, 
+$\theta_2\in\Theta_2\subset \mathbb{R}^q$,
 $a:\mathbb{R}^d\times\Theta_2 \to \mathbb{R}^d$ and $b:\mathbb{R}^d\times\Theta_1 \to \mathbb{R}^d\times \mathbb{R}^r$. 
-
-Quasi-MLE assumes the following approximation of the true log-likelihood for multidimensional diffusions
+The naming of $\theta_2$ and $\theta_1$ is theoretically natural in view of the optimal convergence rates of the estimators for these parameters.
+Given sample data ${\bf X}_n = (X_{t_i})_{i=0,\ldots,n}$ with $t_i = i\Delta_n$,
+QMLE (Quasi Maximum Likelihood Estimator) makes use of the following approximation of the true log-likelihood for multidimensional diffusions
 {\small
@@
 where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $a_i(\theta_2)=a(X_{t_i},\theta_2)$, $\Sigma=b^{\otimes 2}$, \textcolor{red}{$A^{\otimes 2}= A A^T$} and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
-Then, \citep[see e.g.,][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
+Then, \citep[see e.g.,][]{Yoshida92, Kessler97}, the QMLE of $\theta$ is
 \textcolor{red}{$$\hat\theta=\arg\max_\theta \ell_n({\bf X}_n,\theta)$$}
 
-\textcolor{red}{The \pkg{yuima} package implements QML estimation via the  \code{qmle} function.
+\textcolor{red}{The \pkg{yuima} package implements QMLE via the  \code{qmle} function.
 The interface and the output of the \code{qmle} function are made as similar as possible to those of the standard \code{mle} function in the \pkg{stats4} package of the basic \proglang{R} system.
@@
 \begin{equation}
-\de X_t = (2-\theta_2  X_t)  \de t + (1+X_t^2)^{\theta_1}  \de W_t
+\de X_t = (2-\theta_2  X_t)  \de t + (1+X_t^2)^{\theta_1}  \de W_t, \quad X_0=1
 \label{eq:model1}
 \end{equation}
-with $\theta_1=0.2$ and $\theta_2=0.3$. We generate sampled data $X_{t_i}$, with $t_i = i \cdot n^{-\frac23}$.
+with $\theta_1=0.2$ and $\theta_2=0.3$. Before calling \code{qmle}, we generate sampled data $X_{t_i}$, with $t_i = i \cdot n^{-\frac23}$:
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 param.init <- list(theta2=0.5,theta1=0.5)
-upp.par <-  list(theta1=0, theta2=0)
-low.par <-  list(theta1=1, theta2=1)
-mle1 <- qmle(yuima, start = param.init,  lower = upp.par, upper = low.par)
+low.par <-  list(theta1=0, theta2=0)
+upp.par <-  list(theta1=1, theta2=1)
+mle1 <- qmle(yuima, start = param.init,  lower = low.par, upper = upp.par)
 @
@@
 
+\subsubsection{Theoretical remarks on QMLE, High-freq \& ergodicity}
+\textcolor{red}{XXXXX}
+
 \subsection{Adaptive Bayes Estimation}
-Consider again the  diffusion process solution to \eqref{eq:sdemle}
+Consider again the  diffusion process in \eqref{eq:sdemle}
 and the quasi likelihood defined in \eqref{qlik}.
@@
 <<echo=TRUE>>=
-coef(bayes1)
-coef(mle1)
-@
-%The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical
-%integration, otherwise MCMC method is used.
+coef(summary(bayes1))
+coef(summary(mle1))
+@
+The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical integration, otherwise MCMC method is used.
+
+\subsubsection{Theoretical remarks Adaptive Bayes}
+\textcolor{red}{XXXXX}
+
 \subsubsection{The Effect of Small Sample Size on Drift Estimation}
@@
 <<>>=
-coef(bayes2)
-coef(mle2)
+coef(summary(bayes2))
+coef(summary(mle2))
 @
@@
 
+\subsubsection{Add example of two stage estimation}
+\textcolor{red}{XXXXX}
+
 \subsection{LASSO Model Selection}

### 2013-03-05 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
update revision
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@232 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{amsmath}
+
+\def\be{\begin{equation}}
+\def\ee{\end{equation}}
+\def\bea{\begin{eqnarray}}
+\def\eea{\end{eqnarray}}
+\def\beas{\begin{eqnarray*}}
+\def\eeas{\end{eqnarray*}}
+\def\sskip{\hspace{5mm}}
 
@@
         Ryosuke Nomura\\The University of Tokyo\AND
+        Teppei Ogihara\\Osaka University \And
         Yasutaka Shimuzu\\Osaka University \And
@@
 \section{Introduction}
-The plan of the YUIMA Project is to define the bases for a complete environment for simulation and inference for stochastic processes via an \proglang{R} \citep{ERRE} package called \pkg{yuima}.
- The package \pkg{yuima} provides 
-an object-oriented programming environment 
-for simulation and statistical inference 
-for stochastic processes by \proglang{R}.
+The plan of the YUIMA Project is to construct the bases for a complete environment for simulation and inference for stochastic processes via an \proglang{R} \citep{ERRE} package called \pkg{yuima}.
+% The package \pkg{yuima} provides an object-oriented programming environment for simulation and statistical inference for stochastic processes by \proglang{R}.
 The \pkg{yuima} package adopts the  S4 system of classes and methods \citep{chambers98}.
@@
 the \pkg{yuima} package also supplies various functions 
-to execute simulation and statistical analysis. 
-Both categories of procedures may depend \textcolor{red}{on} each other.
+to carry out simulation and statistical analysis. 
+Both categories of procedures may depend on each other.
 Statistical inference often requires a simulation technique 
@@
 practically continuous time, such as 
-the random movements of a stock price. 
+the random movements of stock prices. 
 Theory of statistical inference for 
@@
 familiar with this field to enjoy utilities. 
-\textcolor{red}{For example, by taking advantage of the analytic approach,  the asymptotic expansion method for computing 
+For example, by taking advantage of the analytic approach,  the asymptotic expansion method for computing 
 option prices (i.e., expectation of an irregular functional of 
 a stochastic process) provides precise approximation values 
-instantaneously. The expansion formula, which has a long expression involving more than 900 terms of multiple integrals, is already coded in the \pkg{yuima} package for generic diffusion processes.}
+instantaneously. The expansion formula, which has a long expression involving more than 900 terms of multiple integrals, is already coded in the \pkg{yuima} package for generic diffusion processes.
 
@@
 with simulation and/or statistics for stochastic differential equations. 
-In the \pkg{yuima} package stochastic differential equations can be of very abstract type, 
+In the \pkg{yuima} package, stochastic differential equations can be of very abstract type, 
 multidimensional, driven by Wiener process or fractional Brownian motion 
@@
 The \pkg{yuima} package is intended to offer the basic infrastructure on which complex models and inference procedures can be built on.
-This paper explains the design of the \pkg{yuima} package and provides some examples of applications.
-The paper is organised as follows.
+This paper explains the design of the \pkg{yuima} package and illustrates some examples of applications.
+The paper is organized as follows.
 Section \ref{sec2} is an overview about the package.
-Section \ref{sec3} describe the way models are specified in \pkg{yuima}.
+Section \ref{sec3} describes the way in which models are specified in \pkg{yuima}.
 Section \ref{sec4} explains asymptotic expansion methods.
 Section \ref{sec5} is a review of basic inference procedures.
-Finally, Section \ref{sec6} explains additional details and the roadmap of the YUIMA Project.
-
-Although we assume that the reader of this paper has a basic knowledge of the \proglang{R} language, most of the examples are easy to be understood by anyone.
+Finally, Section \ref{sec6} gives additional details and the roadmap of the YUIMA Project.
+
+Although we assume that the reader of this paper has a basic knowledge of the \proglang{R} language, most of the examples are easy to understand 
+if he/she knows stochastic differential equations intuitively or symbolically.
 
@@
 For example, in case one wants to simulate a stochastic process, only the slots \code{model} and \code{sampling} should be present, while the slot \code{data} will be filled by the simulator.
-We discuss in details the \textcolor{red}{different objects} separately in the next sections.
-
-\textcolor{blue}{The general idea of the \pkg{yuima} package is to separate into different subclass objects the statistical model, the data and the statistical methods.
+We discuss in details the different objects separately in the next sections.
+
+The general idea of the \pkg{yuima} package is to separate into different subclass objects the statistical model, the data and the statistical methods.
 As will be explained with several examples, the user may give a mathematical description of the statistical model with \code{setModel} which prepares a \code{yuima.model} object by filling the appropriate slots.
@@
 Further, functionals of stochastic differential equations can be defined using the \code{setFunctional} method and evaluated using asymptotc expansion methods as explained in Section \ref{sec4}.
-The \code{setFunctional} method creates a \code{yuima.functional} object which is included along with a \code{yuima.model}  into a \code{yuima} object in order to be used for the evaluation of its asymptotic expansion.}
+The \code{setFunctional} method creates a \code{yuima.functional} object which is included along with a \code{yuima.model}  into a \code{yuima} object in order to be used for the evaluation of its asymptotic expansion.
 
 \subsection{The \code{yuima.model} Class}\label{sec:model}
-At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified. \textcolor{blue}{Here we present  a brief
-overview of these models as they will be described in details in Section \ref{sec3}, but this allow to introduce an overall view of the slots of the \code{yuima.model} class.} 
-\textcolor{blue}{In \pkg{yuima} one can describe three main families of stochastic processes at present.
+At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified.  Here we present  a brief
+overview of these models as they will be described in details in Section \ref{sec3}, but this allow to introduce an overall view of the slots of the \code{yuima.model} class.
+In \pkg{yuima} one can describe three main families of stochastic processes at present.
 These models can be one or multidimensional and eventually described as parametric models.
-Let $X_0=x_0$ be the initial value of the process, then, the main classes are as follows:}
+Let $X_0=x_0$ be the initial value of the process, then, the main classes are as follows:
 \begin{itemize}
-\item \textcolor{blue}{diffusion models described as} 
+\item diffusion models described as
 $$  \de X_t=a(t,X_t,\theta)dt + b(t,X_t,\theta)\de W_t, \quad X_0=x_0$$
  where $W_t$ is a standard Brownian motion;
-\item \textcolor{blue}{sde's driven by fractional Gaussian noise, with $H$ the Hurst parameter, described as}
+\item sde's driven by fractional Gaussian noise, with $H$ the Hurst parameter, described as
 $$ \de X_t=a(t,X_t,\theta)dt + b(t,X_t,\theta)\de W_t^{H}, \quad X_0=x_0;$$ 
-\item \textcolor{blue}{diffusion process with jumps} and  L\'evy processes solution to
+\item diffusion process with jumps and  L\'evy processes solution to
 $$
@@
 \end{itemize}
-\textcolor{blue}{The functions $a(\cdot)$, $b(\cdot)$ and $c(\cdot)$ may have a different number of arguments.
+ The functions $a(\cdot)$, $b(\cdot)$ and $c(\cdot)$ may have a different number of arguments.
 For example, if the model is homogeneous in time and drift and diffusion coefficients are entirely specified, then we will use the notaion $a(x)$ and $b(x)$ and describe the diffusion model simply as $\displaystyle \de X_t=a(X_t)dt + b(X_t)\de W_t$.
@@
 Furhter, \pkg{yuima} is able to accept any user-specified notation for the state variable $x$ (for $X_t$) and the time variable $t$ and the remaining term of an \proglang{R} expression will be interpreted as parameter as will explained in Section \ref{sec:diff}.
-We are now able to give an overview of the mian slots of the most important class of the \pkg{yuima} package.}
+We are now able to give an overview of the mian slots of the most important class of the \pkg{yuima} package.
 
@@
   coefficient matrix relative to first noise;
-\item \code{hurst} is the Hurst index of the fractional Brownian motion, by default \code{0.5}  meaning a standard Brownian motion.
-\textcolor{blue}{More details will be given in Section \ref{sec:fgn}};
+\item \code{hurst} is the Hurst index of the fractional Brownian motion, by default \code{0.5}  meaning a standard Brownian motion. More details will be given in Section \ref{sec:fgn};
 \item \code{parameter},  which is a short name for ``parameters'',  
-  is a list with the following entries \textcolor{blue}{(explained details in Section \ref{sec:par})}:
+  is a list with the following entries (explained details in Section \ref{sec:par}):
 \begin{itemize}
@@
 \item \code{jump} contains the parameters belonging to the  coefficient of the L\'evy noise;
-\item \code{measure} contains the parameters \textcolor{red}{describing the L\'evy} measure (\textcolor{blue}{(explained details in Section \ref{sec:jump})};
+\item \code{measure} contains the parameters describing the L\'evy measure (explained details in Section \ref{sec:jump});
 \end{itemize}
-\item \code{measure} specifies the measure of the L\'evy noise (\textcolor{blue}{(see Section \ref{sec:jump})};
-\item \code{measure.type} a switch to specify how the L\'evy measure is described (\textcolor{blue}{(see Section \ref{sec:jump})};
+\item \code{measure} specifies the measure of the L\'evy noise (see Section \ref{sec:jump});
+\item \code{measure.type} a switch to specify how the L\'evy measure is described (see Section \ref{sec:jump});
 \item \code{state.variable} and \code{time.variable}, by default,
-  are assumed to be \code{x} and \code{t} but the user can freely choose them \textcolor{blue}{and they matter the right hand-side of the equation of the sde}.
-  The \code{yuima.model} class assumes that the user either \textcolor{red}{uses} default
+  are assumed to be \code{x} and \code{t} but the user can freely choose them and they matter the right hand-side of the equation of the sde.
+  The \code{yuima.model} class assumes that the user either uses default
   names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
   the rest of the symbols are considered parameters and distributed accordingly
-  in the \code{parameter} slot. \textcolor{blue}{Example of use will be given in Section \ref{sec:diff};}
-\item \code{jump.variable} the name of the variable used in the description of the L\'evy component (\textcolor{blue}{(see Section \ref{sec:jump})};
+  in the \code{parameter} slot. Example of use will be given in Section \ref{sec:diff};
+\item \code{jump.variable} the name of the variable used in the description of the L\'evy component (see Section \ref{sec:jump});
 \item \code{solve.variable} contains a vector of variable names, each element corresponds to the
-   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order. \textcolor{blue}{An example of use can be found in Section \ref{sec:multi}.}
+   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order. An example of use can be found in Section \ref{sec:multi}.
 \item \code{noise.number} indicates the number of sources of noise.
@@
 \section{Model Specification}\label{sec3}
-In order to show \textcolor{red}{how general the approach is} in the \pkg{yuima} package we present some examples.
-\textcolor{blue}{Throughout this section we assume that all the stochastic differential equations exists while in Section \ref{sec5} we will give precise regularity conditions needed to have a properly defined statistical model.}
+In order to show how general the approach is in the \pkg{yuima} package we present some examples.
+Throughout this section we assume that all the stochastic differential equations exists while in Section \ref{sec5} we will give precise regularity conditions needed to have a properly defined statistical model.
 
@@
 $$\de X_t = -3 X_t \de t + \frac{1}{1+X_t^2}\de W_t.$$
-\textcolor{blue}{In the above $a(x) = -3 x$ and $b(x) = \frac{1}{1+x^2}$ according to the notaion of previous section and $W_t$ is a standard Wiener process. }
-This \textcolor{blue}{can be described} in \pkg{yuima} \textcolor{red}{by} specifying the drift and diffusion coefficients as plain 
-\textcolor{blue}{\proglang{R} expressions passed as strings}
+In the above $a(x) = -3 x$ and $b(x) = \frac{1}{1+x^2}$ according to the notaion of previous section and $W_t$ is a standard Wiener process.
+This can be described in \pkg{yuima} by specifying the drift and diffusion coefficients as plain
+\proglang{R} expressions passed as strings
 <<print=FALSE,echo=FALSE,results=hide>>=
@@
 @ 
-\textcolor{blue}{By default, \pkg{yuima} assumes that the state variable is \code{x} and the time variable is \code{t} and the solve variable is again \code{x}.
-Notice that the left hand-side of the equation is implicit, this is why \code{yuima.model} has the slot \code{solve.variable}.
+By default, \pkg{yuima} assumes that the state variable is \code{x} and the time variable is \code{t} and the solve variable is again \code{x}.
+Notice that the left hand-side of the equation is implicit, this is why \code{yuima.model} has the slot \code{solve.variable.
 The user should not we worried about the warning raised by \pkg{yuima} at this stage, as this is just to inform the user about the implicit assumption on the solution variable.}
@@
 
-From the above, it is possible to see that the jump coefficient is void and the Hurst parameter is set to 0.5, \textcolor{blue}{because this is a model where the driving process is the standard Brownian motion, i.e.
-a fractional Brownian motion if Hurst index $H=\frac12$.}
-Now, with \code{mod1} in hands, it is very easy to simulate a trajectory by  \textcolor{blue}{Euler-Maruyama scheme} of the process as follows
+From the above, it is possible to see that the jump coefficient is void and the Hurst parameter is set to 0.5, because this is a model where the driving process is the standard Brownian motion, i.e.
+a fractional Brownian motion if Hurst index $H=\frac12$.
+Now, with \code{mod1} in hands, it is very easy to simulate a trajectory by  Euler-Maruyama scheme of the process as follows
 <<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
@@
 @
-\textcolor{blue}{More details on how to change the default sampling scheme for the \code{simulate} method and how to perform subsampling will be given in Section \ref{sec:simul}.}
+More details on how to change the default sampling scheme for the \code{simulate} method and how to perform subsampling will be given in Section \ref{sec:simul}.
 
@@
 \subsection{Specification of Parametric Models}\label{sec:par}
-\textcolor{blue}{Assume now that we want to specify a parametric model like this}
+Assume now that we want to specify a parametric model like this
 $$\de X_t = -\theta X_t \de t + \frac{1}{1+X_t^\gamma}\de W_t$$
-\textcolor{blue}{where $a(x,\theta) = -\theta x$ and $b(x,\gamma) = 1/(1+x^\gamma)$.}
+where $a(x,\theta) = -\theta x$ and $b(x,\gamma) = 1/(1+x^\gamma)$.
 Then, \pkg{yuima} attempts to distinguish the parameters' names from the ones of the state and time variables
@@
 @
-\textcolor{blue}{so, in this case, \code{theta} and \code{gamma}, which are different form \code{x} and \code{t}, are assumed to be parameters.
-Notice that in the above notation $\theta$ and $\gamma$ are generic names for the components of a parameters' vector $\theta$ in the notation of Section \ref{sec:model}.}
+so, in this case, \code{theta} and \code{gamma}, which are different form \code{x} and \code{t}, are assumed to be parameters.
+Notice that in the above notation $\theta$ and $\gamma$ are generic names for the components of a parameters' vector $\theta$ in the notation of Section \ref{sec:model}.
 <<>>=
@@
 @
-In order to simulate the parametric model it is necessary to specify the values of the \textcolor{blue}{parameters $\theta$ and $\gamma$ as shown in the next code chunk}
+In order to simulate the parametric model it is necessary to specify the values of the parameters $\theta$ and $\gamma$ as shown in the next code chunk
 <<echo=TRUE, print=FALSE,fig=TRUE,height=4,results=hide>>=
@@
 \subsection{Multidimensional Processes}\label{sec:multi}
-\textcolor{blue}{Next is an example of a system of two stochastic differential equations for the couple $(X_{t,1}, X_{t,2})$ driven by three independent Brownian motions $(W_{t,1}, W_{t,2}, W_{t,3})$}
+Next is an example of a system of two stochastic differential equations for the couple $(X_{t,1}, X_{t,2})$ driven by three independent Brownian motions $(W_{t,1}, W_{t,2}, W_{t,3})$
 $$
@@
 $$
-but this has to be organized into matrix form \textcolor{red}{with a vector of drift expressions and a diffusion matrix}
+but this has to be organized into matrix form with a vector of drift expressions and a diffusion matrix
 $$
@@
 $$
-\textcolor{blue}{For this system it now necessary to instruct \pkg{yuima} about the state variable on bot the left-hand side of the equation and the right-hand side of the equation, i.e.
-the \code{solve.variable}.}
+For this system it now necessary to instruct \pkg{yuima} about the state variable on bot the left-hand side of the equation and the right-hand side of the equation, i.e. the \code{solve.variable}.
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 @
-\textcolor{blue}{Looking at the structure of the \code{noise.number} slot in \code{mod3}, one can see that this is now set to 3 which is taken as the number of columns of the diffusion matrix.}
+Looking at the structure of the \code{noise.number} slot in \code{mod3}, one can see that this is now set to 3 which is taken as the number of columns of the diffusion matrix.
 Again, this model can be easily simulated
@@
 @
-\textcolor{blue}{The functions \code{f1}, \code{f2} and \code{f3} are defined in a way that, when the trajectory of the processes crosso zero from above, the trajectory is stopped ad zero.
+The functions \code{f1}, \code{f2} and \code{f3} are defined in a way that, when the trajectory of the processes crosso zero from above, the trajectory is stopped ad zero.
 Notice that in this way the only visible parameter for \pkg{yuima} is \code{mu} as \code{rho} and \code{sig} are inside the bodies of the functions \code{f2} and \code{f3}.
-If we want to instruct \pkg{yuima} about these parameters, they should appear explicitly as arguments of the functions as explained by this \proglang{R} code}
+If we want to instruct \pkg{yuima} about these parameters, they should appear explicitly as arguments of the functions as explained by this \proglang{R} code
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 @
-The user can choose \textcolor{red}{between  two} simulation schemes, namely the Cholesky method and \citet{WoodChan} method.
+The user can choose between  two simulation schemes, namely the Cholesky method and \citet{WoodChan} method.
 This is done via the argument \code{methodfGn} in the \code{simulate} method.
@@
 The \code{simulate} function accepts several arguments including the description of the sampling structure, which is an object of type \code{yuima.sampling}.
-The \code{setSampling} \textcolor{red}{allows} for the specification of different sampling parameters including random sampling.
-Further, the \code{subsampling} \textcolor{red}{allows us} to subsample a trajectory of a simulated stochastic differential equation or a given time series in the \code{yuima.data} slot of a \code{yuima} object.
+The \code{setSampling} allows for the specification of different sampling parameters including random sampling.
+Further, the \code{subsampling} allows us to subsample a trajectory of a simulated stochastic differential equation or a given time series in the \code{yuima.data} slot of a \code{yuima} object.
 Sampling and subsampling can be specified jointly as arguments to the \code{simulate} function.
 This is convenient if one wants to simulate data at very high frequency but then return only low frequency data for inference or other applications.
-\textcolor{red}{In what follows we explain how to specify  arguments of these \pkg{yuima} functions.
-Complete details can be found in  the man pages of the \pkg{yuima} package.}
+In what follows we explain how to specify  arguments of these \pkg{yuima} functions.
+Complete details can be found in  the man pages of the \pkg{yuima} package.
 
@@
 @
-As seen from the output, the sampling structure is \textcolor{red}{quite rich} and we will show how to specify \textcolor{red}{some} of the slots \textcolor{red}{later on}.
+As seen from the output, the sampling structure is quite rich and we will show how to specify some of the slots later on.
 We now simulate this process specifying the \code{sampling} argument to \code{simulate}
@@
 @
-In the above we have specified two independent exponential  distributions to represent \textcolor{red}{Poisson arrival times}. Notice that the slot \code{regular} is now set to \code{FALSE}.
+In the above we have specified two independent exponential  distributions to represent Poisson arrival times. Notice that the slot \code{regular} is now set to \code{FALSE}.
 Now we subsample the original trajectory of \code{X2} using the \code{subsampling} function
@@
 \section{Asymptotic Expansion}\label{sec4}
-{\colorb For numerical computation of the expectation of a random variable,
+For numerical computation of the expectation of a random variable,
 the Monte-Carlo method gives a universal solution although it is time consuming and involves 
@@
 besides, it has advantage in the computation time because of the approximation by an analytic formula.
- }
 
 %The \pkg{yuima} 
-%package {\colorb provides numerical approximation to the expectation of 
-%a functional of a diffusion process in terms of the asymptotic expansion scheme. }
+%package  provides numerical approximation to the expectation of 
+%a functional of a diffusion process in terms of the asymptotic expansion scheme.
 Let us consider a family of 
- $d$-dimensional diffusion processes {\colorb $X=(X^{(\ve)}_t)_{t\in[0,T]}$ ($\ve\in(0,1]$) specified 
- by the stochastic integral equation}
+ $d$-dimensional diffusion processes  $X=(X^{(\ve)}_t)_{t\in[0,T]}$ ($\ve\in(0,1]$) specified 
+ by the stochastic integral equation
 \begin{equation}\label{asy01}
@@
 \end{equation}
-{\colorb for $\ve \in(0,1]$, 
-where}  $W_t=(W_{1,t}, \ldots, W_{r,t})$ \textcolor{red}{is an} $r$-dimensional Wiener process.
+ for $\ve \in(0,1]$, 
+where  $W_t=(W_{1,t}, \ldots, W_{r,t})$ is an $r$-dimensional Wiener process.
 The functional of interest is expressed in the following abstract form
@@
 %
-{\colorb Similarly, for $F(x,\ve)=x$, 
+ Similarly, for $F(x,\ve)=x$, 
 the functional becomes $F^{(\ve)}=X^{(\ve)}_T$ and the price of the European call option is 
 $\dE[\max(X^{(\ve)}_T-K,0)]$. This value has a closed form in the Black-Sholes economy but 
-it is necessary to apply some numerical method for pricing the Asian option even in this linear case. }
-
-{\colorb 
+it is necessary to apply some numerical method for pricing the Asian option even in this linear case.
+
 Returning to the general system (\ref{asy01})-(\ref{yuima:func1}), 
@@
 that is,  
-$$ 
-b(\cdot,0)\>=\>0 \quad \mbox{and}\quad  f_\alpha(\cdot,0)=0\quad (\alpha=1,...,r).
-$$ 
+\beas
+b(\cdot,0)\>=\>0 \sskip\mbox{and}\sskip f_\alpha(\cdot,0)=0\sskip(\alpha=1,...,r).
+\eeas
 %
 Since $X^{(0)}_t$ is deterministic solution to the ordinary differential equation 
-$$
-\de X^{(0)}_t/\de t =a(X^{(0)}_t,0),\quad X^{(0)}_0=x_0,
-$$
+\beas 
+\de X^{(0)}_t/\de t =a(X^{(0)}_t,0),&& X^{(0)}_0=x_0,
+\eeas
 the functional $F^{(0)}$ becomes a constant given by 
-\begin{equation} \label{130106-1}
-F^{(0)} = \int_0^T f_0(X^{(0)}_t,0)\de t+F(X^{(0)}_T,0).
-\end{equation}
+\bea\label{130106-1}
+F^{(0)} &=& \int_0^T f_0(X^{(0)}_t,0)\de t+F(X^{(0)}_T,0).
+\eea
 
@@
 it is possible to show $F^{(\ep)}$ has a version that is smooth in $\ep\in[0,1)$ almost surely, and hence, 
-$$
-\tilde{F}^{(\ep)} := \ep^{-1}(F^{(\ep)}-F^{(0)})
-$$
+\beas 
+\tilde{F}^{(\ep)} &:=& \ep^{-1}(F^{(\ep)}-F^{(0)})
+\eeas
 admits a stochastic expansion 
-\begin{equation}
-\label{130206-8}
-\tilde{F}^{(\ep)} \quad \sim \quad
+\bea\label{130206-8} 
+\tilde{F}^{(\ep)} &\sim& 
 \tilde{F}^{[0]}+\ep\tilde{F}^{[1]}+\ep^2\tilde{F}^{[2]}+\cdots
-\end{equation}
+\eea
 as $\ep\downarrow0$. 
@@
 the asymptotic expansion of the (generalized) expectation 
-\begin{equation} \label{130206-2}
-\dE[g(\tilde{F}^{(\ep)})] \quad \sim \quad d_0(g)
+\bea\label{130206-2}
+\dE[g(\tilde{F}^{(\ep)})] &\sim& d_0(g)
 +\ep d_1(g)+\ep^2 d_2(g)+\cdots
-\end{equation} 
+\eea
 as $\ep\downarrow0$ for measurable functions $g$ at most polynomial growth or 
@@
 In the present situation, each $d_i(g)$ is expresses as 
-$$
-d_i(g) = \int g(z)p_i(z)\phi(z;0,v)dz,
-$$
+\beas 
+d_i(g) &=& \int g(z)p_i(z)\phi(z;0,v)dz,
+\eeas
 where $p_i$ is a polynomial and $\phi(z;0,v)$ is the density of the normal distribution $N(0,v)$ 
@@
 and 
-\begin{equation} \label{130206-5} 
-g(x)= %F^{(0)}-K+\ep x\>=\>\ep\left(
+\bea\label{130206-5} 
+g(x)&=& %F^{(0)}-K+\ep x\>=\>\ep\left(
 \max\left(F^{(0)}-K+\ep x,0\right)%\right). 
-\end{equation} 
+\eea
 Set the model  (\ref{asy01}) and the functional (\ref{yuima:func1}) as follows:
@@
 Thus our method works even under the existence of a stochastic discount factor.
+
 One can compare the result of the asymptotic expansion with other well known techniques like Edgeworth series expansion for the log-normal distribution as proposed, e.g.,  in \citet{Levy92}.
@@
 
-Of course, \pkg{yuima} approach is more general as Levy approximation only holds when the process $X_t$ is the geometric Brownian motion.
-
-\textcolor{red}{NAKAHIRO: ADD REMARK about 92 paper?}
+Of course, \pkg{yuima} approach is more general in that the above Levy approximation only holds when the process $X_t$ is the geometric Brownian motion.
+We now give an example when the underlying process $X_t$ is the following CIR model
+$$
+\de X_t =  0.9 X_t \de t + \ep \sqrt{X_t} \de W_t, X0 = 1
+$$
+and we calculate the asymptotic expansion of a european call option with strike price $K=10$ for $\ep=0.4$.
+
+<<>>=
+a <- 0.9
+e <- 0.4
+Terminal <- 3
+
+xinit <- 1
+K <- 10
+
+drift <- "a * x"
+diffusion <- "e * sqrt(x)"
+
+model <- setModel(drift=drift,diffusion=diffusion)
+
+n <- 1000*Terminal
+yuima <- setYuima(model = model,sampling = setSampling(Terminal=Terminal,n=n))
+
+f <- list(c(expression(0)),c(expression(0)))
+F <- expression(x)
+
+yuima.ae <- setFunctional(yuima,f=f,F=F,xinit=xinit,e=e)
+rho <- expression(0)
+F1 <- F0(yuima.ae)
+
+get_ge <- function(x,epsilon,K,F0){
+	tmp <- (F0 - K) + (epsilon * x[1])
+	tmp[(epsilon * x[1]) > (K - F0)] <- 0
+	return( - tmp )
+}
+
+g <- function(x){
+	return(get_ge(x,e,K,F1))
+}
+
+time1 <- proc.time()
+asymp <- asymptotic_term(yuima.ae,block=100,rho,g)
+time2 <- proc.time()
+@
+We now extract the first and second order terms of the asymptotic expansion from the \code{asymp} object
+<<>>=
+ae.value0 <- asymp$d0
+ae.value0
+ae.value1 <- asymp$d0 + e * asymp$d1
+ae.value1
+ae.value2 <- as.numeric(asymp$d0 + e * asymp$d1 + e^2 * asymp$d2)
+ae.value2
+ae.time <- time2 - time1
+ae.time
+@
+As it can be seen, the contribution of the term corresponding to the second order of the asymptotic expansion give a real contribution to the approximation and the final approximated value \Sexpr{round(ae.value2,5)} can be compared with a Monte Carlo experiment with 1000000 replications
+<<>>=
+time3 <- proc.time()
+
+time4 <- proc.time()
+mc.time <- time4 - time3
+mc.time/ae.time*100
+@
 
@@
 The naming of $\theta_2$ and $\theta_1$ is theoretically natural in view of the optimal convergence rates of the estimators for these parameters.
-Given sample data ${\bf X}_n = (X_{t_i})_{i=0,\ldots,n}$ with $t_i = i\Delta_n$,
+Given sampled data ${\bf X}_n = (X_{t_i})_{i=0,\ldots,n}$ with $t_i = i\Delta_n$,
 QMLE (Quasi Maximum Likelihood Estimator) makes use of the following approximation of the true log-likelihood for multidimensional diffusions
@@
 +\frac{1}{\Delta_n}
-\Sigma_{i-1}^{-1}(\theta_1)[\Delta X_i-\Delta_n a_{i-1}(\theta_2)]^{\otimes 2}\right\}
+\Sigma_{i-1}^{-1}(\theta_1)[(\Delta X_i-\Delta_n a_{i-1}(\theta_2))^{\otimes 2}]\right\}
 \end{eqnarray}}
-where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $a_i(\theta_2)=a(X_{t_i},\theta_2)$, $\Sigma=b^{\otimes 2}$, \textcolor{red}{$A^{\otimes 2}= A A^T$} and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.
-Then, \citep[see e.g.,][]{Yoshida92, Kessler97}, the QMLE of $\theta$ is
-\textcolor{red}{$$\hat\theta=\arg\max_\theta \ell_n({\bf X}_n,\theta)$$}
-
-\textcolor{red}{The \pkg{yuima} package implements QMLE via the  \code{qmle} function.
+\noindent
+where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $a_i(\theta_2)=a(X_{t_i},\theta_2)$, $\Sigma=b^{\otimes 2}$, $A^{\otimes 2}= A A^T$ and $A^{-1}$ the inverse of $A$, $A[B] = \mbox{tr}(A B)$.
+Then 
+%\citep[see e.g.,][]{Yoshida92, Kessler97, Yoshida11}, 
+the QMLE of $\theta$ is an estimator that satisfies  
+$$\hat\theta=\arg\max_\theta \ell_n({\bf X}_n,\theta)$$
+exactly or approximately.
+
+The \pkg{yuima} package implements QMLE via the  \code{qmle} function.
 The interface and the output of the \code{qmle} function are made as similar as possible to those of the standard \code{mle} function in the \pkg{stats4} package of the basic \proglang{R} system.
@@
 If bounds are specified then \code{L-BFGS-B} is used.
-More optimizers can be added in the  future.}
+More optimizers can be added in the  future.
 
@@
 @
-\textcolor{red}{Now the \code{yuima} object contains both the \code{model} and \code{data} slots.
+Now the \code{yuima} object contains both the \code{model} and \code{data} slots.
 We set the initial values for the optimizer as $\theta_1=\theta_2=0.5$ and we specify them as a named list called \code{param.init}.
-We can now use the function \code{qmle} to estimate the parameters $\theta_1$ and $\theta_2$ as follows}
+We can now use the function \code{qmle} to estimate the parameters $\theta_1$ and $\theta_2$ as follows
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 @
-\textcolor{red}{where \code{upp.par} and \code{low.par} are the upper and lower bounds of the search region used by the optimizer (\code{L-BFGS-B} in this case).} 
-\textcolor{red}{The} estimated coefficients are extracted from the output object \code{mle1} as follows
+where \code{upp.par} and \code{low.par} are the upper and lower bounds of the search region used by the optimizer (\code{L-BFGS-B} in this case).
+The estimated coefficients are extracted from the output object \code{mle1} as follows
 <<echo=TRUE, print=FALSE>>=
@@
 
-\subsubsection{Theoretical remarks on QMLE, High-freq \& ergodicity}
-\textcolor{red}{XXXXX}
+\subsubsection{Theoretical remarks on QMLE}%, High-freq \& ergodicity}
+Here are some theoretical remarks. 
+Consistency must be required when an estimator is applied; otherwise 
+the estimation would lose mathematical backing because 
+the more data the observer obtains, the worse the estimator behaves. 
+For consistency of $\hat{\theta}_1$, we are assuming 
+$\Delta_n\to0$ as $n\to\infty$. Indeed, under this condition, $\hat{\theta}_1$ 
+has asymptotically (mixed) normality 
+\citep{GenonCatalotJacod93, UchidaYoshida12}. 
+On the other hand, one needs $T=n\Delta_n\to\infty$ for consistency of 
+$\hat{\theta}_2$ because 
+the Fisher information for $\theta_2$ is finite for a finite $T$ 
+and consistent estimation of $\theta_2$ is theoretically impossible. 
+When $T\to\infty$, usually ergodicity is assumed to ensure a law of large numbers and as a result 
+we obtain 
+consistency of $\hat{\theta}_2$, moreover asymptotic normality. 
+We assume the condition $n\Delta_n^p\to0$ for $p=2$ 
+while applying the quasi log-likelihood (\ref{qlik}) based on the local Gaussian approximation. 
+In practical applications, reduction of the parameter's dimension and relaxing 
+the above condition to $n\Delta_n^p\to0$ for $p$ bigger than $2$ are very important. 
+Adaptive estimation methods were proposed for $p=3$ and for any $p$ 
+in \citep{Yoshida92} and \citep{UchidaYoshida11}, respectively, with the convergence of moments 
+by a large deviation argument. 
+When $T$ is regarded to be not large, the small sample effect on estimation of $\theta_2$ appears, which 
+will be discussed in Section \ref{small.sample}. 
+Modules for QMLE and Bayes estimators are going to be available for jump-diffusions. 
+See Shimizu and Yoshida \citep{ShimizuYoshida2006} 
+and Ogihara and Yoshida \citep{OgiharaYoshida2011}.
 
@@
 =  
-\Big[\int_{\Theta_1}\textcolor{red}{\exp\{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))\}}
-\pi_1(\theta_1)d\theta_1 \Big]^{-1} \int_{\Theta_1} \theta_1 \textcolor{red}{\exp \{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
-\}}\pi_1(\theta_1)d\theta_1
+\Big[\int_{\Theta_1}\exp\{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))\}
+\pi_1(\theta_1)d\theta_1 \Big]^{-1} \int_{\Theta_1} \theta_1 \exp \{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
+\}\pi_1(\theta_1)d\theta_1
 \end{equation}
@@
 $\pi_1$ is a prior density on $\Theta_1$. 
-According to the asymptotic theory, 
-if $\pi_1$ is positive on $\Theta_1$, any function can be used. 
+According to the asymptotic theory 
+under hight-frequent samplings, any function $\pi_1$ can be used
+if it is positive on $\Theta_1$. 
 For estimation of $\theta_2$, we use $\tilde{\theta}_1$ 
@@
 = 
-\Big[\int_{\Theta_2}\textcolor{red}{\exp\{ \ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}}
-\pi_2(\theta_2)d\theta_2 \Big]^{-1} \int_{\Theta_2} \theta_2 \textcolor{red}{\exp \{\ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}}
+\Big[\int_{\Theta_2}\exp\{ \ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}
+\pi_2(\theta_2)d\theta_2 \Big]^{-1} \int_{\Theta_2} \theta_2 \exp \{\ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}
 \pi_2(\theta_2)d\theta_2
@@
 Consider again the model \eqref{eq:model1} with the same values for the parameters.
-In order to perform Bayesian estimation, we need to prepare the prior densities for the parameters.
+In order to perform Bayesian estimation, we prepare prior densities for the parameters.
 For simplicity we use uniform distributions in $[0,1]$
@@
 
-\subsubsection{Theoretical remarks Adaptive Bayes}
-\textcolor{red}{XXXXX}
-
-\subsubsection{The Effect of Small Sample Size on Drift Estimation}
+The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical integration, otherwise MCMC method is used.
+
+\subsubsection{Theoretical Remarks on Adaptive Bayes Estimator}
+Under the same conditions, the adaptive Bayes estimator $\tilde{\theta}_1$ and
+$\tilde{\theta}_2$ perform in the same way as $\hat{\theta}_1$ and $\hat{\theta}_2$, respectively. 
+See the remark in Section \ref{section-qmle} 
+and also \citep{Yoshida92} and \citep{UchidaYoshida12} for details.
+
+\subsubsection{The Effect of Small Sample Size on Drift Estimation}\label{small.sample}
 It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
-In our example above, we took $T=n^\frac13$, with $n = \Sexpr{n}$, which is approximatively  \Sexpr{round(n^(1/3),2)}. Now we reduce the sample size to $n=500$ and then $T=\Sexpr{round(500^(1/3),2)}$.
+In our example above, we took $T=n^\frac13$, with $n = \Sexpr{n}$, which is approximatively   \Sexpr{round(n^(1/3),2)}.
+Now we reduce the sample size to $n=500$ and then $T=\Sexpr{round(500^(1/3),2)}$.
 We then apply both quasi-maximum likelihood and adaptive Bayes type estimators to these data
@@
 irregularly spaced, in general. 
-In particular, \textcolor{red}{\code{cce}} can apply to estimation of the quadratic variation of a single 
+In particular, \code{cce} can apply to estimation of the quadratic variation of a single
 stochastic process sampled regularly/irregularly.
@@
 \subsubsection{Example: Data Generation and Estimation by \pkg{yuima} Package}
-We will demonstrate how to apply \textcolor{red}{\code{cce}} function to 
+We will demonstrate how to apply \code{cce} function to
 nonsynchronous high-frequency data by simulation. 
@@
 \end{equation}
-
 <<>>=
@@
 $$
-where $W_t$ \textcolor{red}{is an} $r$-dimensional Wiener process and $a_t$ and $X_t$ are multidimensional processes, $\theta\in\Theta\subset \mathbb{R}^p$,  $b:\mathbb{R}^d\times\Theta \to \mathbb{R}^d\times \mathbb{R}^r$, is the diffusion coefficient (volatility) matrix.
+where $W_t$ is an $r$-dimensional Wiener process and $a_t$ and $X_t$ are multidimensional processes, $\theta\in\Theta\subset \mathbb{R}^p$,  $b:\mathbb{R}^d\times\Theta \to \mathbb{R}^d\times \mathbb{R}^r$, is the diffusion coefficient (volatility) matrix.
 
@@
 \end{equation}
-\textcolor{red}{and $S=b^{\otimes 2}$.}
+and $S=b^{\otimes 2}$.
 Suppose that there exists an estimator 
@@
 \de W_{1,t}\\ \de W_{2,t} 
-\end{array}\right),\quad \textcolor{red}{ t\in[0,T],}
+\end{array}\right),\quad  t\in[0,T],
 $$
@@
  \de W_{1,t}\\ \de W_{2,t} 
- \end{array}\right),\quad \textcolor{red}{ t\in[0,T],}
+ \end{array}\right),\quad  t\in[0,T],
 $$
@@
 $$
-with change-point instant at time $\tau=0.4$ \textcolor{red}{and $T=10$}.   First, we describe the model to be simulated
+with change-point instant at time $\tau=0.4$  and $T=10$.   First, we describe the model to be simulated
 <<cpoint1, echo=TRUE,results=hide>>=
@@
 @
-and then simulate two trajectories. One up to the change-point $\tau=4$ with parameters \textcolor{red}{$\theta_{1.0}=0.5$ and  $\theta_{2.0}=0.3$}
+and then simulate two trajectories. One up to the change-point $\tau=4$ with parameters $\theta_{1.0}=0.5$ and  $\theta_{2.0}=0.3$
 <<cpoint3,results=hide>>=
@@
  with parameters
-\textcolor{red}{$\theta_{1.1}=0.2$ and  $\theta_{2.1}=0.4$}.
+$\theta_{1.1}=0.2$ and  $\theta_{2.1}=0.4$.
 For this trajectory, the initial value is set to the last value of the first trajectory stored in \code{x1} and \code{x2} for the two component of the process.
@@
 @
-Now we proceed first with the estimation of the parameters before and after the change-point. The \pkg{yuima} package contains two  functions
+\subsubsection{An example of two stage estimation}
+In practical situations, the initial values of the parameters are not known and there is the need to provide a preliminary estimators of them.
+One possible approach is the two stage change-point estimation approach as explained in  \citet{iacyos09}.
+The idea is to take a small subsets of observations at the very beginning and the end of the time series, estimate a
+change point and and then refine the estimation.
+
+To this aim, the \pkg{yuima} package contains two  functions
 which are useful in the framework of change-point or sequential analysis.
@@
 Notice that, even if the estimated parameters are not too accurate because we use a small subsets of observations, the change-point estimate remains good.
-A two stage change-point estimation approach is also possible as explained in  \citet{iacyos09}.
-
-\subsubsection{Add example of two stage estimation}
-\textcolor{red}{XXXXX}
+We can now refine the estimate of the change point and the parameters but iterating the above procedure.
+<<>>=
+qmleL(noDriftModel, t=t.est3$tau, start=list(theta1.k=0.1, theta2.k=0.1),
+lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), 
+method="L-BFGS-B") -> estL
+qmleR(noDriftModel, t=t.est3$tau, start=list(theta1.k=0.1, theta2.k=0.1),
+lower=list(theta1.k=0, theta2.k=0), upper=list(theta1.k=1, theta2.k=1), 
+method="L-BFGS-B") -> estR
+t02s.est <- coef(estL)
+t12s.est <- coef(estR)
+t2s.est3 <- CPoint(noDriftModel,param1=t02s.est,param2=t12s.est)
+t2s.est3
+@
 
@@
 \code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper.
-
-%\section*{Acknowledgements}
-%The YUIMA Project was initially started by Nakahiro Yoshida as principal 
-%investigator and then Stefano M. Iacus joined as coordinator of the \proglang{R}
-%implementation. \textcolor{red}{PLEASE ADD SOME HISTORICAL FACTS AND ACKNOWLEDGMENTS OF DIFFERENT FUNDS}.
+We also plan to open the the contribution to the YUIMA project to external developers in the near future.
+
+\section*{Acknowledgements}
+This work was in part supported by Japan Society for the Promotion of Science Grants-in-Aid for Scientific Research No. 24340015 (Scientific Research), No. 24650148 (Challenging Exploratory Research);
+the Global COE program ``The Research and Training Center for New Development in Mathematics'' of the Graduate School of Mathematical Sciences, University of Tokyo;  by a Cooperative Research Program of the Institute of Statistical Mathematics and by the project PRIN 2009JW2STY, Ministero dell'Istruzione dell'Universit\`a e della Ricerca.
+
+Project YUIMA I was in part supported by JST Basic Research Programs PRESTO.
+Project YUIMA II is in part supported by NS Solutions Corporation. 
+NS Solutions Corporation and The Graduate School of Mathematical Sciences, University of Tokyo are conducting a joint study (Azzurro project) on ``The Application of Advanced Mathematical Statistics Theories in the Financial Industry''.
 

### 2013-03-05 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
update revision
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@233 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\subsection{User Specified State and Time Variables: NEW SECTION}
+\subsection{User Specified State and Time Variables}
 Suppose now that the user wants to specify her own model using a prescribed notation, e.g., some sdes like
@@
 
+\subsection{Asymptotic Expansion for General Stochastic Processes}
 Of course, \pkg{yuima} approach is more general in that the above Levy approximation only holds when the process $X_t$ is the geometric Brownian motion.
@@
 @
-As it can be seen, the contribution of the term corresponding to the second order of the asymptotic expansion give a real contribution to the approximation and the final approximated value \Sexpr{round(ae.value2,5)} can be compared with a Monte Carlo experiment with 1000000 replications
-<<>>=
-time3 <- proc.time()
-
-time4 <- proc.time()
-mc.time <- time4 - time3
-mc.time/ae.time*100
-@
+As it can be seen, the contribution of the term corresponding to the second order of the asymptotic expansion give a real contribution to the approximation and the final approximated value \Sexpr{round(ae.value2,5)} can be compared with a Monte Carlo estimate based on 1000000 replications
+which is equal to \code{0.561059}, but more demanding in terms of cpu time. The relative difference among the two estimates is \Sexpr{round((ae.value2-0.561059)/0.561059*100,1)}\%.
 
@@
 
-\subsection{How to input data into a \code{yuima} object}
+\subsection{How to Input Data Into a \code{yuima} Object}
 Although most of the examples in this section are given on simulated data, the main way to fill up the \code{data} slot of a \code{yuima} object is to use the function \code{setYuima}.
@@
 
-\subsection{Quasi Maximum Likelihood Estimation}
+\subsection{Quasi Maximum Likelihood Estimation}\label{section-qmle}
 Consider a multidimensional diffusion process
@@
 
-\subsubsection{QMLE in practice}
+\subsubsection{QMLE in Practice}
 
@@
 
-\subsubsection{Theoretical remarks on QMLE}%, High-freq \& ergodicity}
+\subsubsection{Theoretical Remarks on QMLE}%, High-freq \& ergodicity}
 Here are some theoretical remarks. 
@@
 @
-\subsubsection{An example of two stage estimation}
+\subsubsection{An Example of Two Stage Estimation}
 In practical situations, the initial values of the parameters are not known and there is the need to provide a preliminary estimators of them.

### 2013-03-09 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@234 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
         Stefano M. Iacus\\University of Milan \AND 	
-        Kengo Kamatani\\The University of Tokyo \And
-        Yuta Koike\\The University of Tokyo\And
+        Kengo Kamatani\\University of Tokyo \And
+        Yuta Koike\\University of Tokyo\And
         Hiroki Masuda\\Kyushu University \And
-        Ryosuke Nomura\\The University of Tokyo\AND
+        Ryosuke Nomura\\University of Tokyo\AND
         Teppei Ogihara\\Osaka University \And
@@
         Masayuki Uchida\\Osaka University \And
-        Nakahiro Yoshida\\The University of Tokyo 
+        Nakahiro Yoshida\\University of Tokyo 
         }
@@
 \code{install.packages("yuima",repos="http://R-Forge.R-project.org")}.
+If, the R-Forge system does not provide binary builds of the \pkg{yuima} package, the
+user can also try 
+\code{install.packages("yuima",repos="http://R-Forge.R-project.org", type="source")}.
 
@@
 \subsection{Multidimensional Processes}\label{sec:multi}
-Next is an example of a system of two stochastic differential equations for the couple $(X_{t,1}, X_{t,2})$ driven by three independent Brownian motions $(W_{t,1}, W_{t,2}, W_{t,3})$
+Next is an example of a system of two stochastic differential equations for the couple $(X_{1,t}, X_{2,t})$ driven by three independent Brownian motions $(W_{1,t}, W_{2,t}, W_{3,t})$
 $$
@@
 \left(\begin{array}{ccc}1 & 0 & X_{2,t} \\X_{1,t} & 3 & 0\end{array}\right)
-\left(\begin{array}{c}\de W_{t,1} \\ \de W_{t,2} \\ \de W_{t,3}\end{array}\right)
+\left(\begin{array}{c}\de W_{1,t} \\ \de W_{2,t} \\ \de W_{3,t}\end{array}\right)
 $$
@@
 \begin{aligned}
-\de X^1_t &= -\theta X^1_t \de t + \de W^1_t + X^2_t \de W^3_t\\
-\de X^2_t &= -(X^1_t + \gamma X^2_t) \de t + X^1_t \de W^1_t + \delta \de W^2_t
+\de X_{1,t} &= -\theta X_{1,t} \de t + \de W_{1,t} + X_{2,t} \de W_{3,t}\\
+\de X_{2,t} &= -(X_{1,t} + \gamma X_{2,t}) \de t + X_{1,t} \de W_{1,t} + \delta \de W_{2,t}
 \end{aligned}
@@
 The \pkg{yuima}  package provides functions to construct the functional $F^{(\ve)}$, and automatic asymptotic expansion based on the Malliavin calculus.  starting from a  \pkg{yuima} object. 
-This asymptotic expansion approach to option pricing was proposed in \citep{Yoshida92b} 
-and many related papers are available today.
+This asymptotic expansion approach to option pricing was proposed in early 1990's 
+( \citep{Yoshida92b}, \citep{Takahashi99}, \citep{Kunitomo01}), and 
+many related papers are available today.
 
@@
 
-\subsection{Asymptotic Expansion for General Stochastic Processes}
+\subsubsection{Asymptotic Expansion for General Stochastic Processes}
 Of course, \pkg{yuima} approach is more general in that the above Levy approximation only holds when the process $X_t$ is the geometric Brownian motion.

### 2013-03-09 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@235 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \section{Miscellanea and Roadmap of YUIMA Project}\label{sec6}
-Other statistical techniques are already implemented or will be shortly released in the \pkg{yuima}.
-For example, a nice utility is the function \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}.
-A simple writing like
+Other statistical techniques are already implemented in the developer version of the \pkg{yuima} package.
+Among these: the QMLE approach for stochastic differential equations with L\'evy component, the estimation for the fractional Ornstein-Uhlembeck model \citep{brouiac12},
+
+or will be shortly released in the \pkg{yuima}. For example, a nice utility is the function \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}. A simple writing like
 \code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper.

### 2013-03-10 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@236 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2013-03-09 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
@@
 \section{Miscellanea and Roadmap of YUIMA Project}\label{sec6}
-Other statistical techniques are already implemented or will be shortly released in the \pkg{yuima}.
-For example, a nice utility is the function \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}.
-A simple writing like
-\code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper.
-We also plan to open the the contribution to the YUIMA project to external developers in the near future.
+Other statistical techniques have been are already implemented in the developer version of the \pkg{yuima} package although not yet released into the current distribution.
+Among these we mention: the QMLE approach for stochastic differential equations with L\'evy noise \citep{shiyos06, ogiyos11}; the parametric estimation for the fractional Ornstein-Uhlembeck model \citep{brouiac12}; different simulation schemes as in \citep{Iacus08} for multidimensional diffusion processes; Lead-lag estimation \citep{HoffRosYo09}.
+ For example, a nice (yet incomplete) utility is the method \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}. A simple writing like
+\code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper. For example,
+<<>>=
+a <- c("-3*x1","-x1-2*x2")
+b <- matrix(c("1","x1","0","3","x2","0"),2,3)
+modtex <- setModel(drift = a, diffusion = b, solve.variable = c("x1","x2"))
+toLatex(modtex)
+@
+which can be typesetted with  \LaTeX{} to produce
+<<echo=FALSE,results=tex>>=
+toLatex(modtex)
+@
+which can be used as a quick starting point for more complex editing of a mathematical paper.
+Another major plan is to open the  contribution to the YUIMA project to external developers in the near future as well as to include other estimation procedures for low frequency and/or sparse data and parallelization of the general infrastructure.
 

### 2013-07-18 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
final version of submission
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@243 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
         Masaaki Fukasawa\\Osaka University\And
-        Hideitsu Hino\\Waseda University\And
+        Hideitsu Hino\\University of Tsukuba\And
         Stefano M. Iacus\\University of Milan \AND 	
@@
 \section{Introduction}
-The plan of the YUIMA Project is to construct the bases for a complete environment for simulation and inference for stochastic processes via an \proglang{R} \citep{ERRE} package called \pkg{yuima}.
+The plan of the YUIMA\footnote{YUIMA is both the acronym for \textit{Yoshida-Uchida-Iacus-Masuda-Andothers} but also the name of an important character in Buddhism religion (\url{http://www.kyohaku.go.jp/eng/syuzou/meihin/kaiga/chuugoku/item01.html}) whose approach to problems fits well this project.} Project is to construct the bases for a complete environment for simulation and inference for stochastic processes via an \proglang{R} \citep{ERRE} package called \pkg{yuima}.
 % The package \pkg{yuima} provides an object-oriented programming environment for simulation and statistical inference for stochastic processes by \proglang{R}.
@@
 Section \ref{sec3} describes the way in which models are specified in \pkg{yuima}.
+Section \ref{sec:simul} describes how to simulate \code{yuima} models.
 Section \ref{sec4} explains asymptotic expansion methods.
@@
 For example, in case one wants to simulate a stochastic process, only the slots \code{model} and \code{sampling} should be present, while the slot \code{data} will be filled by the simulator.
-We discuss in details the different objects separately in the next sections.
+We discuss in detail the different objects separately in the next sections.
 
@@
 If the aim is the simulation of the solution of the stochastic  differential equation specified in the \code{yuima.model} object then, using the method \code{simulate}, it is possible to obtain one trajectory of the process.
-As an output, a \code{yuima} object is created which contains the original model specified in the \code{yuima.model} object in the slot named \code{model} and two additional slots named \code{data}, for the simulated data, and \code{sampling} which contains the description of the simulation scheme used as well as other informations.
+As an output, a \code{yuima} object (by ``\code{yuima} object'' we mean a, possibly incomplete, object of class \pkg{yuima}) is created which contains the original model specified in the \code{yuima.model} object in the slot named \code{model} and two additional slots named \code{data}, for the simulated data, and \code{sampling} which contains the description of the simulation scheme used as well as other informations.
 The details of \code{simulate} will be explained in Section \ref{sec:simul} along with the use of method \code{setSampling} which allows to specify the type of sampling scheme to be used by the \code{simulate} method.
@@
 In this case the method \code{setData} is used to transform most types  of \proglang{R} time series object into a a proper \code{yuima.data} object.
-When the slots \code{data} and \code{model} are available, many other methods can be used to perform statistical analysis on these sde's models. These methods will be discussed in Section \ref{sec5}.
-Further, functionals of stochastic differential equations can be defined using the \code{setFunctional} method and evaluated using asymptotc expansion methods as explained in Section \ref{sec4}.
+When the slots \code{data} and \code{model} are available, many other methods can be used to perform statistical analysis on these SDE's (stochastic differential equation) models.
+These methods will be discussed in Section \ref{sec5}.
+Further, functionals of stochastic differential equations can be defined using the \code{setFunctional} method and evaluated using asymptotic expansion methods as explained in Section \ref{sec4}.
 The \code{setFunctional} method creates a \code{yuima.functional} object which is included along with a \code{yuima.model}  into a \code{yuima} object in order to be used for the evaluation of its asymptotic expansion.
@@
 At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified.  Here we present  a brief
-overview of these models as they will be described in details in Section \ref{sec3}, but this allow to introduce an overall view of the slots of the \code{yuima.model} class.
+overview of these models as they will be described in detail in Section \ref{sec3}, but this allows to introduce an overall view of the slots of the \code{yuima.model} class.
 In \pkg{yuima} one can describe three main families of stochastic processes at present.
@@
  where $W_t$ is a standard Brownian motion;
-\item sde's driven by fractional Gaussian noise, with $H$ the Hurst parameter, described as
+\item SDE's driven by fractional Gaussian noise, with $H$ the Hurst parameter, described as
 $$ \de X_t=a(t,X_t,\theta)dt + b(t,X_t,\theta)\de W_t^{H}, \quad X_0=x_0;$$ 
@@
 Detailed hypotheses and regularity conditions on the coefficients $a(\cdot)$, $b(\cdot)$ and $c(\cdot)$ for each class of models will be given in the next sections.
-Nevertheless, it is important to remark that these notations only matter the mathematical description of the model while each coefficient is passed to \pkg{yuima} methods as \proglang{R} mathematical expressions.
-It means that, for example, $a(t, X_t, \theta) = t\cdot \sqrt{\theta X_t}$ will be passed as \code{t*sqrt(x*theta)}, therefore, the order of the arguments is not relevant to \proglang{R} as well as its mathematical description as long as it is consistent through each specific section.
-Furhter, \pkg{yuima} is able to accept any user-specified notation for the state variable $x$ (for $X_t$) and the time variable $t$ and the remaining term of an \proglang{R} expression will be interpreted as parameter as will explained in Section \ref{sec:diff}.
-We are now able to give an overview of the mian slots of the most important class of the \pkg{yuima} package.
+Nevertheless, it is important to remark that these notations only matter to the mathematical description of the model while each coefficient is passed to \pkg{yuima} methods as \proglang{R} mathematical expressions.
+This means that, for example, $a(t, X_t, \theta) = t\cdot \sqrt{\theta X_t}$ will be passed as \code{t*sqrt(x*theta)}, therefore, the order of the arguments is not relevant to \proglang{R} as well as its mathematical description as long as it is consistent through each specific section.
+Furhter, \pkg{yuima} is able to accept any user-specified notation for the state variable $x$ (for $X_t$) and the time variable $t$ and the remaining term of an \proglang{R} expression will be interpreted as parameter as will be explained in Section \ref{sec:diff}.
+We are now able to give an overview of the main slots of the most important class of the \pkg{yuima} package.
 
@@
 \item \code{parameter},  which is a short name for ``parameters'',  
-  is a list with the following entries (explained details in Section \ref{sec:par}):
+  is a list with the following entries (more details in Section \ref{sec:par}):
 \begin{itemize}
@@
 \item \code{state.variable} and \code{time.variable}, by default,
-  are assumed to be \code{x} and \code{t} but the user can freely choose them and they matter the right hand-side of the equation of the sde.
+  are assumed to be \code{x} and \code{t} but the user can freely choose them and they matter to the right hand-side of the equation of the SDE.
   The \code{yuima.model} class assumes that the user either uses default
-  names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
-  the rest of the symbols are considered parameters and distributed accordingly
+  names for \code{state.variable} and \code{time.variable} variables or specifies his own names. All
+  other  symbols are considered parameters and distributed accordingly
   in the \code{parameter} slot. Example of use will be given in Section \ref{sec:diff};
-\item \code{jump.variable} the name of the variable used in the description of the L\'evy component (see Section \ref{sec:jump});
+\item \code{jump.variable} indicates the name of the variable used in the description of the L\'evy component (see Section \ref{sec:jump});
 \item \code{solve.variable} contains a vector of variable names, each element corresponds to the
@@
 \item \code{noise.number} indicates the number of sources of noise.
-\item \code{xinit} initial value of the stochastic differential equation;
+\item \code{xinit} is the initial value of the stochastic differential equation;
 \item \code{equation.number} represents the number of equations, i.e., the number of one dimensional
@@
   same length of \code{parameter} with the same names.
-\item \code{J.flag} for internal use only.
+\item \code{J.flag} is for internal use only.
 \end{itemize}
@@
 Usually, the rate of convergence for estimators in the diffusion coefficient are similar to the ones in the i.i.d.
-sampling while estimators of parameters in the drift coefficient are slower or, in some cases, not even consistent.
+(independent and idensitcally distributed) sampling while estimators of parameters in the drift coefficient are slower or, in some cases, not even consistent.
 The \pkg{yuima} always tries to implement the best statistical inference for the given model under the observed sampling scheme.
@@
 \section{Model Specification}\label{sec3}
-In order to show how general the approach is in the \pkg{yuima} package we present some examples.
-Throughout this section we assume that all the stochastic differential equations exists while in Section \ref{sec5} we will give precise regularity conditions needed to have a properly defined statistical model.
+In order to show how general the approach of the \pkg{yuima} package is, we present some examples.
+Throughout this section we assume that all the stochastic differential equations exist while in Section \ref{sec5} we will give  regularity conditions needed to have a properly defined statistical model.
 
@@
 By default, \pkg{yuima} assumes that the state variable is \code{x} and the time variable is \code{t} and the solve variable is again \code{x}.
-Notice that the left hand-side of the equation is implicit, this is why \code{yuima.model} has the slot \code{solve.variable.
-The user should not we worried about the warning raised by \pkg{yuima} at this stage, as this is just to inform the user about the implicit assumption on the solution variable.}
+Notice that the left hand-side of the equation is implicit, this is why \code{yuima.model} has the slot \code{solve.variable}.
+The user should not be worried about the warning raised by \pkg{yuima} at this stage, as this is just to inform her on the implicit assumption on the solution variable.
 At this point, the package fills the proper slots of the \code{yuima} object
@@
 a fractional Brownian motion if Hurst index $H=\frac12$.
-Now, with \code{mod1} in hands, it is very easy to simulate a trajectory by  Euler-Maruyama scheme of the process as follows
+Now, with \code{mod1} in hands, it is extremely easy to simulate a trajectory by  Euler-Maruyama scheme of the process as follows
 <<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
@@
 \subsection{User Specified State and Time Variables}
-Suppose now that the user wants to specify her own model using a prescribed notation, e.g., some sdes like
+Suppose now that the user wants to specify her own model using a prescribed notation, e.g., some SDE's like
 $$\de Y_s = -3 s Y_s \de s + \frac{1}{1+Y_s^2}\de W_s,$$
@@
 $$
-For this system it now necessary to instruct \pkg{yuima} about the state variable on bot the left-hand side of the equation and the right-hand side of the equation, i.e. the \code{solve.variable}.
+For this system it is now necessary to instruct \pkg{yuima} about the state variable on both the left-hand side of the equation and the right-hand side of the equation, i.e.
+there is the need to specify also the \code{solve.variable} for the left hand side of the SDE:
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 
+Notice that the role of \code{solve.variable} is essential because it allows to specify the left hand side of an SDE equation.
+For example, if one wants to specify this model instead of the previous one
+
+$$
+\begin{aligned}
+\de X_{2,t} &= -3 X_{1,t} \de t + \de W_{1,t} + X_{2,t} \de W_{3,t}\\
+\de X_{1,t} &= -(X_{1,t} + 2 X_{2,t}) \de t + X_{1,t} \de W_{1,t} + 3 \de W_{2,t}
+\end{aligned}
+$$
+the \code{solve.variable} argument should be specified as \code{solve.variable=c("x2","x1")} in place of  \code{solve.variable=c("x1","x2")}, all the rest being the same as in model \code{mod3}.
+
 \noindent
 But it is also possible to specify more complex models like the following
- \begin{equation} \label{sabr}
+$$
  \begin{cases}
@@
  \end{cases}
- \end{equation}
+$$
 $$ (X_{1,0}, X_{2,0}, X_{3,0}) = (1.0,0.1,1.0)$$
@@
  This is a stochastic integral equation defined as
- \begin{equation}
+ $$
  X_t = X_0 +  \int_0^t a(s,X_s) \mbox{d}s + \int_0^t b(s,X_s) \mbox{d}W_s
- \end{equation}
+ $$
  with
@@
 @
-The functions \code{f1}, \code{f2} and \code{f3} are defined in a way that, when the trajectory of the processes crosso zero from above, the trajectory is stopped ad zero.
+The functions \code{f1}, \code{f2} and \code{f3} are defined in a way that, when the trajectory of the processes crosses zero from above, the trajectory is stopped at zero.
 Notice that in this way the only visible parameter for \pkg{yuima} is \code{mu} as \code{rho} and \code{sig} are inside the bodies of the functions \code{f2} and \code{f3}.
-If we want to instruct \pkg{yuima} about these parameters, they should appear explicitly as arguments of the functions as explained by this \proglang{R} code
+If we want to instruct \pkg{yuima} about these parameters, they should appear explicitly as arguments of the functions as shown in the following \proglang{R} code
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 where $W^H=\left( W^H_t, t\geq 0 \right)$ is a normalized fractional Brownian motion (fBM), {\it i.e.} the zero mean Gaussian processes with covariance function
-$$\dE( W_s^H W_t^H )= \frac{1}{2} \left[ |s|^{2H}+|t|^{2H}- |t-s|^{2H}\right]$$ with Hurst exponent $H\in(0,1)$.
+$$\dE( W_s^H W_t^H )= \frac{1}{2} \left( |s|^{2H}+|t|^{2H}- |t-s|^{2H}\right)$$ with Hurst exponent $H\in(0,1)$.
 The fractional Brownian motion process is  neither Markovian nor a semimartingale for $H\neq\frac{1}{2}$ but
 remains Gaussian.
-For $H > \frac12$, the solution $X_t$ above  presents the long-range dependance property that makes it useful for different applications in biology, physics, ethernet traffic or finance.
+For $H > \frac12$, the solution $X_t$ above  presents the long-range dependence property that makes it useful for different applications in biology, physics, ethernet traffic or finance.
 In order to specify a stochastic differential equation driven by fractional Gaussian noise it is necessary to specify the value of the Hurst parameter.
@@
 Let $Z_t$ be a  Compound Poisson Process (i.e., jumps size follow some distribution, like the Gaussian law and jumps occur at Poisson times).
-Then it is possible to consider the following SDE which involves jumps
-$$\de X_t =  a(X_t)\de t + b(X_t)\de W_t + \de Z_t$$
-In the next example we consider a compound Poisson process with intensity $\lambda=10$ with Gaussian jumps.
-This model can be specified in \code{setModel} using the argument  \code{measure.type="CP"} 
-A simple Ornstein-Uhlenbeck process with Gaussian jumps 
+Then it is possible to consider the simple following SDE which involves jumps
+$$\de X_t =  a(t,X_t,\theta)\de t + b(t,X_t,\theta)\de W_t + \de Z_t,$$
+or the more general representation
+\begin{equation}\label{eq:levy}
+\begin{aligned}
+\de X_t = & \,\,\, a(t,X_t,\theta)\de t + b(t,X_t,\theta)\de W_t + \int\limits_{|z|>1}\!\!\! c(X_{t-},z)\mu(\de t,\de z) \\
+&{}+\!\! \int\limits_{0<|z|\le 1}\!\!\!  c(X_{t-},z)\{\mu(\de t,\de z) - \nu(\de z)\de t\}
+\end{aligned}, \quad X_0=x_0,
+\end{equation}
+with $\mu$ a random measure associated with the jumps of $X$, that is,
+$$
+\mu(dt,dz)=\sum_{s>0}\mathbf{1}_{\{\Delta Z_s \ne 0\}}\delta_{(s,\Delta Z_s)}(dt,dz),
+$$
+where $\delta$ denotes the Dirac measure and $Z$ the driving pure-jump L\'evy process of the form
+$$
+Z_{t}=\int_{0}^{t}\int_{|z|\le 1}z\{\mu(ds,dz) - \nu(dz)ds\}+\int_{0}^{t}\int_{|z|>1}z\mu(ds,dz).
+$$
+The L\'evy measure $\nu$ is any measure satisfying $\nu(\{0\})=0$ and $\int(1\wedge |z|^2)\nu(dz)< \infty$.
+In the next example we consider a compound Poisson process with intensity $\lambda=10$ with Gaussian jumps as driving L\'evy process.
+This model can be specified in \code{setModel} using the argument  \code{measure.type="CP"} (fro Compound Poisson).
+A simple Ornstein-Uhlenbeck process with Gaussian jumps like this
 $$\de X_t = -\theta X_t \de t + \sigma \de W_t + \de Z_t$$
-is specified as
+is then specified as follows
 <<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
@@
 \noindent
-Another possibility is to specify the jump component via the L\'evy measure.
-Without going into too much details, here is an example of specification of a simple Ornstein-Uhlenbeck  process with IG (Inverse Gaussian) L\'evy measure
+Another possibility is to specify the jump component via the L\'evy measure $\nu(\cdot)$. 
+ Without going into too much details, here is an example of specification of a simple Ornstein-Uhlenbeck  process with IG (Inverse Gaussian) L\'evy measure $\nu(\cdot)$ and no Poisson component
 $$\de X_t = -x \de t + \de Z_t$$
@@
 
-\subsection{Specification of Generic Models}
+\noindent
+The argument \code{type="code"} stands for user defined code.
+\subsection{Specification of Generic Models in \pkg{yuima}}
 In general, the \pkg{yuima} package allows to specify a large family of models solutions to
-$$\de X_t =a(X_t)\de t +  b(X_t)\de W_t + c(X_t)\de Z_t$$
+$$\de X_t =a(t,X_t,\theta)\de t +  b(t,X_t,\theta)\de W_t + c(t,X_t,\theta)\de Z_t$$
 using the following interface
@@
 @
+The coefficient $c()$ may also depend on the process $Z$, in this case the admissible form for the coefficient is $c(t,x,\theta, z) = c(t,x,\theta) \cdot z$.
+In the integral representation correspoding to formula \eqref{eq:levy} the coefficient $c(\cdot)$ is not allowed to depend on time though.
 The \pkg{yuima} package implements many multivariate Random Numbers Generators (RNG) which are needed to simulate L\'evy paths including
@@
 The \code{simulate} function simulates \code{yuima} models according to Euler-Maruyama scheme in the presence of non-fractional diffusion noise and L\'evy jumps and uses the Cholesky or the Wood and Chan methods for the fractional Gaussian noise.
+For diffusion models without jumps, \pkg{yuima} also implements the space discretized Euler-Maruyama method, which is more efficient, in the sense of mean squared error approximation, than the usual time-discretized Euler-Maruyama scheme.
+The (time discretized) Euler-Maruyama simulation scheme defines a grid of times
+$0 = \tau_0 < \tau_1 < \cdots < \tau_j < \tau_{j+1} < \cdots$,
+and constructs an approximative solution $\{\tilde{X}_t, t\geq 0\}$ of the original process $\{X_t, t\geq 0\}$ given by
+$$
+ \tilde{X}_{\tau_{j+1}} = \tilde{X}_{\tau_j} +
+ b(\tau_j,\tilde{X}_{\tau_j})(\tau_{j+1} - \tau_j) + 
+ c(\tau_j, \tilde{X}_{\tau_j} )(W_{\tau_{j+1}}- W_{\tau_j}), \ \ j \geq 0.
+$$
+ As the interval $\tau_{j+1} - \tau_j$ is independent to  $\{W_t\}_{t \geq \tau_j}$ for each $j$,
+ $W_{\tau_{j+1}} - W_{\tau_j}$ has the same distribution as $\sqrt{\tau_{j+1} - \tau_j}N_j$,
+ where $N_j$ is an i.i.d. sequence of standard normal variables.
+ The basic discretization scheme is  equidistant, i.e. 
+ $\tau_{j+1} - \tau_j = T/n = \Delta_n$, for all $j \geq 0$.
+But the \code{simulate} method provides also space-discretized Euler-Maruyama method to solve SDE.
+ This is at the moment designed for 1 factor SDE only, i.e., the case the driving
+ Brownian motion $W$ is 1-dimensional. In this case, the discretization scheme $\{\tau_j\}$ 
+ is defined as
+$$
+\tau_0 = 0, \ \ \tau_{j+1} = \inf\{ t > \tau_j; |W_t - W_{\tau_j}| = \epsilon \}
+$$ for each $j \geq 0$. Internally, \code{simulate} takes
+$\epsilon^2 = T/n = \Delta_n$
+which coincides with the mean of the interval  $\tau_{j+1} - \tau_j$.
+This space-discretizing scheme is known to be 3 times efficient than the usual time-discretized scheme
+one in the sense of the mean squared error \citep{Fukasawa11}.
+To make use of the space-discretized Euler-Maruyama scheme, one should use the argument \code{space.discretized = TRUE} which by default, is set to \code{FALSE}.
+Other simulation scheme for 1-dimensional diffusion processes as explained in \citet{Iacus08} will be implemented in the near future.
+
 The \code{simulate} function accepts several arguments including the description of the sampling structure, which is an object of type \code{yuima.sampling}.
 The \code{setSampling} allows for the specification of different sampling parameters including random sampling.
-Further, the \code{subsampling} allows us to subsample a trajectory of a simulated stochastic differential equation or a given time series in the \code{yuima.data} slot of a \code{yuima} object.
+Further, the \code{subsampling} allows  to subsample a trajectory of a simulated stochastic differential equation or a given time series in the \code{yuima.data} slot of a \code{yuima} object.
 Sampling and subsampling can be specified jointly as arguments to the \code{simulate} function.
@@
 In what follows we explain how to specify  arguments of these \pkg{yuima} functions.
-Complete details can be found in  the man pages of the \pkg{yuima} package.
-
-Assume that we want to simulate this model
+Complete details can be found in  the manual pages of the \pkg{yuima} package.
+
+Assume that we want to simulate the following model
 $$
@@
 @
-and look inside it
+and let us analyize its content
 <<>>=
@@
 As seen from the output, the sampling structure is quite rich and we will show how to specify some of the slots later on.
-We now simulate this process specifying the \code{sampling} argument to \code{simulate}
+We  simulate this process by specifying the \code{sampling} argument in the \code{simulate} method
 <<>>=
@@
 @
-now the sampling structure is recorded along with the data in the \code{yuima} object \code{X2}
+the sampling structure is recorded along with the data in the \code{yuima} object \code{X2}
 <<>>=
@@
 @
-In the above we have specified two independent exponential  distributions to represent Poisson arrival times. Notice that the slot \code{regular} is now set to \code{FALSE}.
-Now we subsample the original trajectory of \code{X2} using the \code{subsampling} function
+In the above we have specified two independent exponential  distributions to represent Poisson arrival times. Notice that the slot \code{regular} is  set to \code{FALSE}.
+We subsample the original trajectory of \code{X2} using the \code{subsampling} function
 <<sub2,fig=TRUE,width=9,height=5>>=
@@
 
-Or we can operate a deterministic sampling specifying two different regular frequencies
+We can also operate a deterministic sampling by specifying two different regular frequencies
 <<sub3,fig=TRUE,width=9,height=5>>=
@@
 What is usually done in simulation studies, is to simulate the process at very high frequency but then use data for estimation at a lower frequency.
-This can be done in a single step in the following way.
+This can be done in a single step in the following way:
 <<sub4,fig=TRUE,width=9,height=5>>=
@@
 
-In the previous code we have simulated the process twice just to show the effect of the subsampling but the reader should use only the line which outputs the simulation to \code{Y.sub} as seen in the next plot.
+In the previous code we have simulated the process twice just to show the effect of the subsampling, but the reader should use only the line which outputs the simulation to \code{Y.sub} as seen in the next plot.
 <<sub5,fig=TRUE,width=9,height=5>>=
@@
 the Monte-Carlo method gives a universal solution although it is time consuming and involves 
- stochastic errors of certain scale 
-depending on the number of repetitions. 
-An alternative method is the asymptotic expansion, that can often give a solution with accuracy comparable or superior to  
-Monte-Carlo approaches, 
-besides, it has advantage in the computation time because of the approximation by an analytic formula.
+ stochastic errors of a certain scale 
+depending on the number of replications. 
+An alternative tool is the asymptotic expansion method that can often give a solution with accuracy comparable or superior to  
+Monte-Carlo methods. The asymptotic expansion method
+has an advantage in the computational time because the approximation is given through an analytic formula.
 
@@
 the functional becomes $F^{(\ve)}=X^{(\ve)}_T$ and the price of the European call option is 
-$\dE[\max(X^{(\ve)}_T-K,0)]$. This value has a closed form in the Black-Sholes economy but 
+$\dE[\max(X^{(\ve)}_T-K,0)]$. This value has a closed form in the Black-Sholes economy, but 
 it is necessary to apply some numerical method for pricing the Asian option even in this linear case.
@@
 %
-Since $X^{(0)}_t$ is deterministic solution to the ordinary differential equation 
+Since $X^{(0)}_t$ is the deterministic solution to the ordinary differential equation
 \beas 
@@
 \eea
-as $\ep\downarrow0$ for measurable functions $g$ at most polynomial growth or 
-more generally for Schwartz distributions, 
-under the uniform nondegeneracy of the Malliavain covariance of $\tilde{F}^{(\ep)}$.
-\footnote{This condition ensures smoothness of the distribution of $\tilde{F}^{(\ep)}$. 
+as $\ep\downarrow0$ for measurable functions $g$ at most polynomial growth or, 
+more generally, for Schwartz distributions, 
+under the uniform nondegeneracy of the Malliavin covariance of $\tilde{F}^{(\ep)}$.
+\footnote{This condition ensures the smoothness of the distribution of $\tilde{F}^{(\ep)}$. 
 It should be remarked that the Watanabe's theory is more general than the present use 
@@
 %
-In the present situation, each $d_i(g)$ is expresses as 
+In the present situation, each $d_i(g)$ is expressed as 
 \beas 
@@
 and accurate approximation by analytic formulation.
-The \pkg{yuima}  package provides functions to construct the functional $F^{(\ve)}$, and automatic asymptotic expansion based on the Malliavin calculus.  starting from a  \pkg{yuima} object. 
+The \pkg{yuima}  package provides functions to construct the functional $F^{(\ve)}$ and perform automatic asymptotic expansion based on the Malliavin calculus  starting from a  \pkg{yuima} object. 
 This asymptotic expansion approach to option pricing was proposed in early 1990's 
 ( \citep{Yoshida92b}, \citep{Takahashi99}, \citep{Kunitomo01}), and 
-many related papers are available today.
+several related papers are available today.
 
 Though our method can be applied to the nonlinear system (\ref{asy01})-(\ref{yuima:func1}), 
-just for an example, we shall consider 
+just as an example, we shall consider 
 the Asian call option of the geometric Brownian motion of equation \eqref{eqyuima:gbm1} with $\mu=1$ and $x_0=1$, 
@@
 $$
-\de X_t =  0.9 X_t \de t + \ep \sqrt{X_t} \de W_t, X0 = 1
+\de X_t =  0.9 X_t \de t + \ep \sqrt{X_t} \de W_t,\quad X_0 = 1
 $$
@@
 @
-As it can be seen, the contribution of the term corresponding to the second order of the asymptotic expansion give a real contribution to the approximation and the final approximated value \Sexpr{round(ae.value2,5)} can be compared with a Monte Carlo estimate based on 1000000 replications
+As it can be seen, the contribution of the term corresponding to the second order of the asymptotic expansion gives a real contribution to the approximation and the final approximated value \Sexpr{round(ae.value2,5)} can be compared with a Monte Carlo estimate based on 1000000 replications
 which is equal to \code{0.561059}, but more demanding in terms of cpu time. The relative difference among the two estimates is \Sexpr{round((ae.value2-0.561059)/0.561059*100,1)}\%.
@@
 The function \code{setYuima} sets various slots of the \code{yuima} object.
-In particular, to estimate a \code{yuima.model} called \code{mod} on the data \code{X} one can use a code like this \code{my.yuima <- setYuima(data=setData(X), model=mod)} and then pass \code{my.yuima} to the inference functions as described in what follows.
+In particular, to estimate a \code{yuima.model} called \code{mod} on the data \code{X} one can use a code like this \code{my.yuima <- setYuima(data=setData(X), model=mod)} and then pass \code{my.yuima} to the inference functions as described in the following.
 
@@
 The naming of $\theta_2$ and $\theta_1$ is theoretically natural in view of the optimal convergence rates of the estimators for these parameters.
-Given sampled data ${\bf X}_n = (X_{t_i})_{i=0,\ldots,n}$ with $t_i = i\Delta_n$,
+Given sampled data ${\bf X}_n = (X_{t_i})_{i=0,\ldots,n}$ with $t_i = i\Delta_n$, $\Delta_n \to 0$ as $n\to \infty$,
 QMLE (Quasi Maximum Likelihood Estimator) makes use of the following approximation of the true log-likelihood for multidimensional diffusions
@@
 @
-Now the \code{yuima} object contains both the \code{model} and \code{data} slots.
+Now the \code{yuima} object contains both the \code{model} and the \code{data} slots.
 We set the initial values for the optimizer as $\theta_1=\theta_2=0.5$ and we specify them as a named list called \code{param.init}.
@@
 \subsubsection{Theoretical Remarks on QMLE}%, High-freq \& ergodicity}
-Here are some theoretical remarks. 
-Consistency must be required when an estimator is applied; otherwise 
-the estimation would lose mathematical backing because 
+Consistency of an estimator is always a  required property; otherwise 
+the estimation would loose mathematical backing because 
 the more data the observer obtains, the worse the estimator behaves. 
-For consistency of $\hat{\theta}_1$, we are assuming 
+For the consistency of $\hat{\theta}_1$, we are assuming 
 $\Delta_n\to0$ as $n\to\infty$. Indeed, under this condition, $\hat{\theta}_1$ 
@@
 \citep{GenonCatalotJacod93, UchidaYoshida12}. 
-On the other hand, one needs $T=n\Delta_n\to\infty$ for consistency of 
-$\hat{\theta}_2$ because 
+On the other hand, one needs $T=n\Delta_n\to\infty$ for the consistency of 
+$\hat{\theta}_2$ since 
 the Fisher information for $\theta_2$ is finite for a finite $T$ 
@@
 When $T\to\infty$, usually ergodicity is assumed to ensure a law of large numbers and as a result 
-we obtain 
-consistency of $\hat{\theta}_2$, moreover asymptotic normality. 
+the  
+consistency of $\hat{\theta}_2$ is obtained. Moreover, asymptotic normality is also established. 
 We assume the condition $n\Delta_n^p\to0$ for $p=2$ 
@@
 In practical applications, reduction of the parameter's dimension and relaxing 
-the above condition to $n\Delta_n^p\to0$ for $p$ bigger than $2$ are very important. 
+the above condition to $n\Delta_n^p\to0$ for $p$ larger than $2$ are extremely important.
 Adaptive estimation methods were proposed for $p=3$ and for any $p$ 
@@
 First we choose an initial arbitrary value $\theta_2^\star\in\Theta_2$ and 
-pretend $\theta_1$ is the unknown parameter to 
-make the Bayesian type estimator $\tilde{\theta}_1$ as 
+pretend $\theta_1$ is the unknown parameter for which we construct the Bayesian type estimator $\tilde{\theta}_1$ as  follows
 \begin{equation} 
@@
 =  
-\Big[\int_{\Theta_1}\exp\{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))\}
-\pi_1(\theta_1)d\theta_1 \Big]^{-1} \int_{\Theta_1} \theta_1 \exp \{\ell_n(\mathbf{ x}_n,(\theta_1,\theta_2^\star))
+\Big[\int_{\Theta_1}\exp\{\ell_n(\mathbf{ X}_n,(\theta_1,\theta_2^\star))\}
+\pi_1(\theta_1)d\theta_1 \Big]^{-1} \int_{\Theta_1} \theta_1 \exp \{\ell_n(\mathbf{ X}_n,(\theta_1,\theta_2^\star))
 \}\pi_1(\theta_1)d\theta_1
@@
 According to the asymptotic theory 
-under hight-frequent samplings, any function $\pi_1$ can be used
+under high-frequency samplings, any function $\pi_1$ can be used
 if it is positive on $\Theta_1$. 
-For estimation of $\theta_2$, we use $\tilde{\theta}_1$ 
+For the estimation of $\theta_2$, we use $\tilde{\theta}_1$ 
 to reform the quasi-likelihood function. That is,
@@
 = 
-\Big[\int_{\Theta_2}\exp\{ \ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}
-\pi_2(\theta_2)d\theta_2 \Big]^{-1} \int_{\Theta_2} \theta_2 \exp \{\ell_n(\mathbf{ x}_n,(\tilde{\theta}_1,\theta_2))\}
+\Big[\int_{\Theta_2}\exp\{ \ell_n(\mathbf{ X}_n,(\tilde{\theta}_1,\theta_2))\}
+\pi_2(\theta_2)d\theta_2 \Big]^{-1} \int_{\Theta_2} \theta_2 \exp \{\ell_n(\mathbf{ X}_n,(\tilde{\theta}_1,\theta_2))\}
 \pi_2(\theta_2)d\theta_2
@@
 
-The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical integration, otherwise MCMC method is used.
-
 \subsubsection{Theoretical Remarks on Adaptive Bayes Estimator}
@@
 See the remark in Section \ref{section-qmle} 
-and also \citep{Yoshida92} and \citep{UchidaYoshida12} for details.
+and also \citet{Yoshida92} and \citet{UchidaYoshida12} for details.
 
 \subsubsection{The Effect of Small Sample Size on Drift Estimation}\label{small.sample}
-It is known from the theory that the estimation of the drift in a diffusion process strongly depend on the length of the observation interval $[0,T]$.
+It is known from the theory that the estimation of the drift in a diffusion process strongly depends on the length of the observation interval $[0,T]$.
 In our example above, we took $T=n^\frac13$, with $n = \Sexpr{n}$, which is approximatively   \Sexpr{round(n^(1/3),2)}.
@@
 
-The process $X_l$ is supposed to be observed at 
+The process $X_l$ is supposed to be observed  
 over the increasing sequence of times 
@@
 Thus, the observables are $(T^{l,i},X_{l,i})$ with $T^{l,i}\leq T$. 
-Each $T^{l,j}$ may be a stopping time, so possibly depends on 
-the history of $(X_1,X_2)$ as well as the precedent stopping times.  
+Each $T^{l,j}$ may be a stopping time, so it possibly depends on 
+the history of $(X_1,X_2)$ as well as on the precedent stopping times.  
 Two sequences of stopping times 
@@
 \end{equation}
-The target variable $\theta$ is random in general.
-
-It can be estimated with the nonsynchronous covariance estimator 
+The target variable $\theta$ is random in general and
+it can be estimated with the nonsynchronous covariance estimator 
 (Hayashi-Yoshida estimator) 
@@
 to the sum only when 
-the respective observation intervals $(T^{1,i-1},T^{1,i}] $ and $ (T^{2,j-1},T^{2,j}] $ are overlapping 
-with each other. 
+the respective observation intervals $(T^{1,i-1},T^{1,i}] $ and $ (T^{2,j-1},T^{2,j}] $ are overlapping.
 %The estimator $U_n$ was proposed by Hayashi and Yoshida  and 
-It is known that $U_n$ is consist and has asymptotically mixed normal 
+It is known that $U_n$ is consistent and has asymptotically mixed normal 
 distribution 
@@
 "diff.coef.2(t,x1,x2) * sqrt(1-cor.rho(t,x1,x2)^2)"),2,2)
-# Model sde using yuima.model
+# Model SDE using yuima.model
 cor.mod <- setModel(drift = c("",""), diffusion = diff.coef.matrix,
@@
 and the two random measures are independent each other as well as 
-the stochastic processes. Under this scheme data become asynchronous. It is known that 
+the stochastic processes. Under this scheme the data become asynchronous. It is known that 
 \begin{equation}
@@
 \code{cce} takes the sample and returns 
-an estimate of the quadratic covariation. For example, for the complete data
+an estimate of the quadratic covariation. For example, for the complete data, we obtain the
+following estimates
 <<>>=
@@
 $$
-where $W_t$ is an $r$-dimensional Wiener process and $a_t$ and $X_t$ are multidimensional processes, $\theta\in\Theta\subset \mathbb{R}^p$,  $b:\mathbb{R}^d\times\Theta \to \mathbb{R}^d\times \mathbb{R}^r$, is the diffusion coefficient (volatility) matrix.
-
-When $Y=X$ the problem is a diffusion model.
+where $W_t$ is a $r$-dimensional Wiener process and $a_t$ and $X_t$ are multidimensional processes, $\theta\in\Theta\subset \mathbb{R}^p$,  $b:\mathbb{R}^d\times\Theta \to \mathbb{R}^d\times \mathbb{R}^r$, is the diffusion coefficient (volatility) matrix.
+
+When $Y=X$ the model above is a diffusion model.
 The process $a_t$ may have jumps but should not explode and it is treated as a nuisance in this model.
@@
 $$
-with change-point instant at time $\tau=0.4$  and $T=10$.   First, we describe the model to be simulated
+with change-point instant at time $\tau=4$  and $T=10$.   First, we describe the model to be simulated
 <<cpoint1, echo=TRUE,results=hide>>=
@@
 
-As said, the change-point analysis do not consider the information coming from the drift part of the model and \pkg{yuima} ignores this internally.
+As said, the change-point analysis does not consider the information coming from the drift part of the model and \pkg{yuima} ignores this internally.
 Just to make clear that the information on the drift term is not considered by the function \code{CPoint}, we redefine the \code{yuima} model removing the information coming from the drift and then adding back the data.
@@
 @
-First we show that there is no difference in using the complete model or the model without drift. For simplicity, we assume the true values of the parameters for $\theta_{1.k}$ and $\theta_{1.k}$
+First we show that there is no difference in using the complete model or the model without drift.
+For simplicity, we assume to know the true values of the parameters for $\theta_{1.k}$ and $\theta_{1.k}$
 <<cpoint5>>=
@@
 @
+As it can be seen, the above estimates of the change point are the same for the complete model \code{yuima} and the model without drift \code{noDriftModel}.
 \subsubsection{An Example of Two Stage Estimation}
@@
 One possible approach is the two stage change-point estimation approach as explained in  \citet{iacyos09}.
-The idea is to take a small subsets of observations at the very beginning and the end of the time series, estimate a
+The idea is to take a small subset of observations at the very beginning and the end of the time series, estimate a
 change point and and then refine the estimation.
@@
 Notice that, even if the estimated parameters are not too accurate because we use a small subsets of observations, the change-point estimate remains good.
-We can now refine the estimate of the change point and the parameters but iterating the above procedure.
+We can now refine the estimate of the change point and the parameters by iterating the above procedure.
 <<>>=
@@
 @
-
 \subsection{LASSO Model Selection}
@@
 We denote by $\theta=(\alpha,\beta)\in\Theta_p\times \Theta_q=\Theta$ the parametric vector and with $\theta_0=(\alpha_0,\beta_0)$ its unknown true value. 
-Let $\mathbb{H}_n({\bf X}_n,\theta) = \ell_n({\bf X}_n,\theta)$ from equation \eqref{qlik}.
-The quasi-MLE $\hat{\theta}$ for this model is the solution of the following problem
+Let $\mathbb{H}_n({\bf X}_n,\theta) = -\ell_n({\bf X}_n,\theta)$ from equation \eqref{qlik}.
+The QMLE $\hat{\theta}$ for this model is the solution of the following problem
 $$
@@
 $$
+and $\ddot{\mathbb{H}}_n$ is the matrix of second partial derivatives of $\mathbb{H}$ with respect to the vector $\theta$.
 For more details see \citet{DegIac10b}.
@@
 \end{equation}
-where $\hat \alpha_{j}$ and  $\hat \beta_{k}$ are the unpenalized QML estimator of $\alpha_j$ and $\beta_k$ respectively, $\delta_1, \delta_2>0$ and usually taken unitary.
+where $\hat \alpha_{j}$ and  $\hat \beta_{k}$ are the unpenalized QMLE's of $\alpha_j$ and $\beta_k$ respectively, $\delta_1, \delta_2>0$ and usually taken unitary.
 
@@
 The \code{lasso} method is implemented in the \pkg{yuima} package.
-Let us consider the full CKLS model
+Let us consider the full CKLS model \citep{CKLS92}
 $$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
@@
 @
-and now we apply the \code{lasso} function
+where \code{lambda10} stores the penaly terms $\lambda_0$ and $\gamma_0$ of \eqref{eq:penalty} of the LASSO method.
+Now we apply the \code{lasso} function
 <<echo=TRUE,results=hide>>=
@@
 Other statistical techniques have been are already implemented in the developer version of the \pkg{yuima} package although not yet released into the current distribution.
-Among these we mention: the QMLE approach for stochastic differential equations with L\'evy noise \citep{shiyos06, ogiyos11}; the parametric estimation for the fractional Ornstein-Uhlembeck model \citep{brouiac12}; different simulation schemes as in \citep{Iacus08} for multidimensional diffusion processes; Lead-lag estimation \citep{HoffRosYo09}.
+Among these we mention: the QMLE approach for stochastic differential equations with L\'evy noise \citep{shiyos06, ogiyos11}; the parametric estimation for the fractional Ornstein-Uhlembeck model \citep{brouiac12}; different simulation schemes as in \citep{Iacus08} for multidimensional diffusion processes; Lead-lag estimation \citep{HoffRosYo09}; hypotheses testing via $\phi$-divergence \citep{degiac12}.
  For example, a nice (yet incomplete) utility is the method \code{toLatex} for objects of class \code{yuima} and \code{yuima.model}. A simple writing like
-\code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copy and pasted in a mathematical paper. For example,
+\code{toLatex(my-yuima-obj)} produces the related \LaTeX{} code which can be copied and pasted in a mathematical paper. For example,
 <<>>=

### 2013-10-10 pkg/yuimadocs/inst/doc/JSS/final-edited.Rnw
print methods
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@250 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2013-07-18 pkg/yuimadocs/inst/doc/JSS/article-new.Rnw
@@
 If the aim is the simulation of the solution of the stochastic  differential equation specified in the \code{yuima.model} object then, using the method \code{simulate}, it is possible to obtain one trajectory of the process.
-As an output, a \code{yuima} object (by ``\code{yuima} object'' we mean a, possibly incomplete, object of class \pkg{yuima}) is created which contains the original model specified in the \code{yuima.model} object in the slot named \code{model} and two additional slots named \code{data}, for the simulated data, and \code{sampling} which contains the description of the simulation scheme used as well as other informations.
+As an output, a \code{yuima} object (by ``\code{yuima} object'' we mean a, possibly incomplete, object of class \code{yuima}) is created which contains the original model specified in the \code{yuima.model} object in the slot named \code{model} and two additional slots named \code{data}, for the simulated data, and \code{sampling} which contains the description of the simulation scheme used as well as other informations.
 The details of \code{simulate} will be explained in Section \ref{sec:simul} along with the use of method \code{setSampling} which allows to specify the type of sampling scheme to be used by the \code{simulate} method.
@@
 This means that, for example, $a(t, X_t, \theta) = t\cdot \sqrt{\theta X_t}$ will be passed as \code{t*sqrt(x*theta)}, therefore, the order of the arguments is not relevant to \proglang{R} as well as its mathematical description as long as it is consistent through each specific section.
-Furhter, \pkg{yuima} is able to accept any user-specified notation for the state variable $x$ (for $X_t$) and the time variable $t$ and the remaining term of an \proglang{R} expression will be interpreted as parameter as will be explained in Section \ref{sec:diff}.
+Further, \pkg{yuima} is able to accept any user-specified notation for the state variable $x$ (for $X_t$) and the time variable $t$ and the remaining term of an \proglang{R} expression will be interpreted as parameter as will be explained in Section \ref{sec:diff}.
 We are now able to give an overview of the main slots of the most important class of the \pkg{yuima} package.
@@
 An object of class \code{yuima.model} contains several slots listed below. 
-To see inside its structure, one can use the \proglang{R} command \code{str} on a \pkg{yuima} object.
+To see inside its structure, one can use the \proglang{R} command \code{str} on a \code{yuima} object.
 \begin{itemize}
@@
 Usually, the rate of convergence for estimators in the diffusion coefficient are similar to the ones in the i.i.d.
-(independent and idensitcally distributed) sampling while estimators of parameters in the drift coefficient are slower or, in some cases, not even consistent.
-The \pkg{yuima} always tries to implement the best statistical inference for the given model under the observed sampling scheme.
+(independent and identically distributed) sampling while estimators of parameters in the drift coefficient are slower or, in some cases, not even consistent.
+The \pkg{yuima} package always tries to implement the best statistical inference for the given model under the observed sampling scheme.
 
@@
 $$\de X_t = -3 X_t \de t + \frac{1}{1+X_t^2}\de W_t.$$
-In the above $a(x) = -3 x$ and $b(x) = \frac{1}{1+x^2}$ according to the notaion of previous section and $W_t$ is a standard Wiener process.
+In the above $a(x) = -3 x$ and $b(x) = \frac{1}{1+x^2}$ according to the notation of previous section and $W_t$ is a standard Wiener process.
 This can be described in \pkg{yuima} by specifying the drift and diffusion coefficients as plain

### 2013-10-11 pkg/yuimadocs/inst/doc/JSS/final-edited.Rnw
edited for JSS final
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@251 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\usepackage{float}
 \def\be{\begin{equation}}
@@
 
-\section{The \pkg{yuima} Package}\label{sec2}
+\section{The \pkg{yuima} package}\label{sec2}
 The package \pkg{yuima} depends on some other packages, like  \pkg{zoo} \citep{zoo}, which can be installed separately.
@@
 
-\subsection{How to Obtain the Package}
+\subsection{How to obtain the package}
 The \pkg{yuima} package is hosted on R-Forge and the web page
@@
 
-\subsection{The Main Object and Classes}
+\subsection{The main object and classes}
 Before discussing the methods for simulation and inference for stochastic processes solutions to stochastic differential equations, here we discuss the main classes in the package.
@@
 the main class is called the \code{yuima-class}. This class  is composed of several slots.
-Figure \ref{fig:classes} represents the different classes and their slots.
+Figure~\ref{fig:classes} represents the different classes and their slots.
 \begin{figure}[th]
@@
 
-\subsection{The \code{yuima.model} Class}\label{sec:model}
+\subsection{The \code{yuima.model} class}\label{sec:model}
 At present, in \pkg{yuima}  three main classes of stochastic differential equations  can be easily specified.  Here we present  a brief
@@
 
-\section{Model Specification}\label{sec3}
+\section{Model specification}\label{sec3}
 In order to show how general the approach of the \pkg{yuima} package is, we present some examples.
@@
 
-\subsection{One Dimensional Diffusion Processes}\label{sec:diff}
+\subsection{One dimensional diffusion processes}\label{sec:diff}
 Assume that we want to describe the following stochastic differential equation
@@
 <<print=FALSE,echo=FALSE,results=hide>>=
-library(yuima)
+library("yuima")
 @ 
@@
 
-From the above, it is possible to see that the jump coefficient is void and the Hurst parameter is set to 0.5, because this is a model where the driving process is the standard Brownian motion, i.e.
-a fractional Brownian motion if Hurst index $H=\frac12$.
+From the above, it is possible to see that the jump coefficient is void and the Hurst parameter is set to 0.5, because this is a model where the driving process is the standard Brownian motion, i.e., a fractional Brownian motion if Hurst index $H=\frac12$.
 Now, with \code{mod1} in hands, it is extremely easy to simulate a trajectory by  Euler-Maruyama scheme of the process as follows
-<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+<<sim-mod1,echo=TRUE, print=FALSE,results=hide>>=
 set.seed(123)
 X <- simulate(mod1)
+@
+which can be plotted using the command \code{plot}
+<<plot-mod1,echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 plot(X)
 @
+and the result is shown in Figure~\ref{fig:plot-mod1}.
+\begin{figure}[H]
+<<echo=FALSE, print=FALSE,results=hide>>=
+pdf("plot-mod1.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(X)
+dev.off()
+@
+\includegraphics[width=\textwidth]{plot-mod1}
+\caption{The \code{plot} function is used to draw a trajectory of a simulated \code{yuima} object.}
+\label{fig:plot-mod1}
+\end{figure}
 
@@
 
-\subsection{User Specified State and Time Variables}
+\subsection{User specified state and time variables}
 Suppose now that the user wants to specify her own model using a prescribed notation, e.g., some SDE's like
@@
 
-\subsection{Specification of Parametric Models}\label{sec:par}
+\subsection{Specification of parametric models}\label{sec:par}
 Assume now that we want to specify a parametric model like this
@@
 In order to simulate the parametric model it is necessary to specify the values of the parameters $\theta$ and $\gamma$ as shown in the next code chunk
-<<echo=TRUE, print=FALSE,fig=TRUE,height=4,results=hide>>=
+<<sim-mod2,echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 set.seed(123)
@@
 @
-
-\subsection{Multidimensional Processes}\label{sec:multi}
+and the trajectory can be seen in Figure~\ref{fig:plot-mod2}.
+\begin{figure}[H]
+<<plot-mod2,echo=FALSE, print=FALSE,results=hide>>=
+pdf("plot-mod2.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(X)
+dev.off()
+@
+\includegraphics[width=\textwidth]{plot-mod2}
+\caption{A trajectory simulated from the parametric model  \code{mod2}.}
+\label{fig:plot-mod2}
+\end{figure}
+
+\subsection{Multidimensional processes}\label{sec:multi}
 Next is an example of a system of two stochastic differential equations for the couple $(X_{1,t}, X_{2,t})$ driven by three independent Brownian motions $(W_{1,t}, W_{2,t}, W_{3,t})$
@@
 $$
-For this system it is now necessary to instruct \pkg{yuima} about the state variable on both the left-hand side of the equation and the right-hand side of the equation, i.e.
-there is the need to specify also the \code{solve.variable} for the left hand side of the SDE:
+For this system it is now necessary to instruct \pkg{yuima} about the state variable on both the left-hand side of the equation and the right-hand side of the equation, i.e., there is the need to specify also the \code{solve.variable} for the left hand side of the SDE:
 <<echo=TRUE, print=FALSE,results=hide>>=
@@
 Looking at the structure of the \code{noise.number} slot in \code{mod3}, one can see that this is now set to 3 which is taken as the number of columns of the diffusion matrix.
-Again, this model can be easily simulated
-<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+Again, this model can be easily simulated and the trajectory can be seen in Figure~\ref{fig:plot-mod3}.
+<<sim-mod3,echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 set.seed(123)
@@
 @
-
+<<plot-mod3,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-mod3.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(X, plot.type="single",lty=1:2)
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-mod3}
+\caption{A trajectory of the multidimensional SDE described in \code{mod3}.}
+\label{fig:plot-mod3}
+\end{figure} 
 Notice that the role of \code{solve.variable} is essential because it allows to specify the left hand side of an SDE equation.
 For example, if one wants to specify this model instead of the previous one
-
 $$
@@
 
-\subsection{Fractional Gaussian Noise}\label{sec:fgn}
+\subsection{Fractional gaussian noise}\label{sec:fgn}
 The \pkg{yuima} allows for the description of stochastic differential equations driven by fractional Brownian motion of the following type
@@
 $$
-where $W^H=\left( W^H_t, t\geq 0 \right)$ is a normalized fractional Brownian motion (fBM), {\it i.e.} the zero mean Gaussian processes with covariance function
+where $W^H=\left( W^H_t, t\geq 0 \right)$ is a normalized fractional Brownian motion (fBM), {\it i.e.}, the zero mean Gaussian processes with covariance function
 $$\dE( W_s^H W_t^H )= \frac{1}{2} \left( |s|^{2H}+|t|^{2H}- |t-s|^{2H}\right)$$ with Hurst exponent $H\in(0,1)$.
@@
 In order to specify a stochastic differential equation driven by fractional Gaussian noise it is necessary to specify the value of the Hurst parameter.
-For example, if we want to specify the following model
+For example, if we want to specify the following  fractional Ornstein-Uhlenbeck model
 $$\de Y_t =  3 Y_t \de t + \de W_t^H$$
-we proceed as follows
-<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=6,results=hide>>=
+we can proceed as follows
+<<sim-mod4AB, echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 mod4A <- setModel(drift="3*y", diffusion=1, hurst=0.3, solve.var="y")
@@
 @
+and the two trajectories can be seen in Figure~\ref{fig:plot-mod4AB}.
+\begin{figure}[H]
+<<plot-mod4AB,echo=FALSE, print=FALSE,results=hide>>=
+pdf("plot-mod4AB.pdf",width=9,height=6)
+par(mfrow=c(2,1))
+par(mar=c(2,3,1,1))
+plot(X1,main="H=0.3")
+plot(X2,main="H=0.7")
+dev.off()
+@
+\includegraphics[width=\textwidth]{plot-mod4AB}
+\caption{Trajectories of the fractional Ornstein-Uhlenbeck process for different values of the Hurst parameter.}
+\label{fig:plot-mod4AB}
+\end{figure}
 
@@
 
-\subsection{L\'evy Processes}\label{sec:jump}
+\subsection{L\'evy processes}\label{sec:jump}
 Jump processes can be specified in different ways in mathematics and hence in \pkg{yuima} package. 
@@
 $$\de X_t = -\theta X_t \de t + \sigma \de W_t + \de Z_t$$
-is then specified as follows
-<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+is then specified as follows (the trajectory is shown in Figure~\ref{fig:plot-mod5})
+<<sim-mod5,echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 mod5 <- setModel(drift=c("-theta*x"), diffusion="sigma",
@@
 @
-
+\begin{figure}[H]
+<<plot-mod5,echo=FALSE, print=FALSE,results=hide>>=
+pdf("plot-mod5.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(X)
+dev.off()
+@
+\includegraphics[width=\textwidth]{plot-mod5}
+\caption{A trajectory of the Ornstein-Uhlenbeck process with jumps following a compound Poisson process with Gaussian jumps as defined in \code{mod5}.}
+\label{fig:plot-mod5}
+\end{figure} 
+%%
 \noindent
 Another possibility is to specify the jump component via the L\'evy measure $\nu(\cdot)$. 
- Without going into too much details, here is an example of specification of a simple Ornstein-Uhlenbeck  process with IG (Inverse Gaussian) L\'evy measure $\nu(\cdot)$ and no Poisson component
+ Without going into too much details, here is an example of specification of a simple Ornstein-Uhlenbeck  process with IG (Inverse Gaussian) L\'evy measure $\nu(\cdot)$ and no Poisson component (see Figure~\ref{fig:plot-mod6})
 $$\de X_t = -x \de t + \de Z_t$$
-<<echo=TRUE, print=FALSE,fig=TRUE,width=9,height=4,results=hide>>=
+<<sim-mod6,echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 mod6 <- setModel(drift="-x", xinit=1, jump.coeff="1", 
@@
 @
-
 \noindent
 The argument \code{type="code"} stands for user defined code.
-\subsection{Specification of Generic Models in \pkg{yuima}}
+\begin{figure}[H]
+<<plot-mod6,echo=FALSE, print=FALSE,results=hide>>=
+pdf("plot-mod6.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(X)
+dev.off()
+@
+\includegraphics[width=\textwidth]{plot-mod6}
+\caption{A trajectory of the pure jump process with inverse Gaussian L\'evy measure for jumps as defined in \code{mod6}.}
+\label{fig:plot-mod6}
+\end{figure} 
+\subsection{Specification of generic models in \pkg{yuima}}
 In general, the \pkg{yuima} package allows to specify a large family of models solutions to
@@
 
-\section{Simulation, Sampling and Subsampling}\label{sec:simul}
+\section{Simulation, sampling and subsampling}\label{sec:simul}
 The \code{simulate} function simulates \code{yuima} models according to Euler-Maruyama scheme in the presence of non-fractional diffusion noise and L\'evy jumps and uses the Cholesky or the Wood and Chan methods for the fractional Gaussian noise.
@@
  where $N_j$ is an i.i.d. sequence of standard normal variables.
- The basic discretization scheme is  equidistant, i.e. 
+ The basic discretization scheme is  equidistant, i.e., 
  $\tau_{j+1} - \tau_j = T/n = \Delta_n$, for all $j \geq 0$.
@@
 In the above we have specified two independent exponential  distributions to represent Poisson arrival times. Notice that the slot \code{regular} is  set to \code{FALSE}.
-We subsample the original trajectory of \code{X2} using the \code{subsampling} function
-<<sub2,fig=TRUE,width=9,height=5>>=
+We subsample the original trajectory of \code{X2} using the \code{subsampling} function (see Figure~\ref{fig:plot-sub2})
+<<sub2,echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 newdata <- subsampling(X2, sampling=newsamp)
@@
 @
-
-We can also operate a deterministic sampling by specifying two different regular frequencies
-<<sub3,fig=TRUE,width=9,height=5>>=
+<<plot-sub2,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-sub2.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(X2,plot.type="single", lty=c(1,3),ylab="X2")
+points(get.zoo.data(newdata)[[1]],col="red")
+points(get.zoo.data(newdata)[[2]],col="green",pch=18)
+dev.off()
+@
+We can also operate a deterministic sampling by specifying two different regular frequencies (see Figure~\ref{fig:plot-sub3})
+<<sub3,echo=TRUE, print=FALSE,fig=FALSE,results=hide>>=
 newsamp <- setSampling(delta=c(0.1,0.2))
@@
 @
-
+<<plot-sub3,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-sub3.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(X2,plot.type="single", lty=c(1,3),ylab="X2")
+points(get.zoo.data(newdata)[[1]],col="red")
+points(get.zoo.data(newdata)[[2]],col="green", pch=18)
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-sub2}
+\caption{An example of Poisson random sub-sampling: green and red dots are sampled according to two different and independent Poisson processes.}
+\label{fig:plot-sub2}
+\end{figure} 
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-sub3}
+\caption{An example of deterministic sub-sampling: the frequency of red dots is two times the one of the green dots.}
+\label{fig:plot-sub3}
+\end{figure} 
 Again one can look at the structure of the sampling structure.
@@
 Subsampling can be used within the \code{simulate} function.
-What is usually done in simulation studies, is to simulate the process at very high frequency but then use data for estimation at a lower frequency.
+What is usually done in simulation studies, is to simulate the process at very high frequency but then use data for estimation at a lower frequency (see Figure~\ref{fig:plot-sub4}).
 This can be done in a single step in the following way:
-<<sub4,fig=TRUE,width=9,height=5>>=
+<<sub4,fig=FALSE>>=
 set.seed(123)
@@
 @
-
-In the previous code we have simulated the process twice just to show the effect of the subsampling, but the reader should use only the line which outputs the simulation to \code{Y.sub} as seen in the next plot.
-<<sub5,fig=TRUE,width=9,height=5>>=
+<<plot-sub4,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-sub4.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(Y, plot.type="single")
+points(get.zoo.data(Y.sub)[[1]],col="red")
+points(get.zoo.data(Y.sub)[[2]],col="green",pch=18)
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-sub4}
+\caption{An example of sub-sampling used within the \code{simulate} command.}
+\label{fig:plot-sub4}
+\end{figure} 
+In the previous code we have simulated the process twice just to show the effect of the subsampling, but the reader should use only the line which outputs the simulation to \code{Y.sub} as seen in Figure~\ref{fig:plot-sub5}.
+<<sub5,fig=FALSE>>=
 plot(Y.sub, plot.type="single")
 @
-
-\section{Asymptotic Expansion}\label{sec4}
+<<plot-sub5,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-sub5.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(Y.sub, plot.type="single")
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-sub5}
+\caption{Plotting directly the sub-sampled trajectory  \code{Y.sub}.}
+\label{fig:plot-sub5}
+\end{figure}
+
+\section{Asymptotic expansion}\label{sec4}
 For numerical computation of the expectation of a random variable,
@@
 and accurate approximation by analytic formulation.
-The \pkg{yuima}  package provides functions to construct the functional $F^{(\ve)}$ and perform automatic asymptotic expansion based on the Malliavin calculus  starting from a  \pkg{yuima} object. 
+The \pkg{yuima}  package provides functions to construct the functional $F^{(\ve)}$ and perform automatic asymptotic expansion based on the Malliavin calculus  starting from a  \code{yuima} object. 
 This asymptotic expansion approach to option pricing was proposed in early 1990's 
-( \citep{Yoshida92b}, \citep{Takahashi99}, \citep{Kunitomo01}), and 
+ \citep{Yoshida92b,Takahashi99,Kunitomo01}, and 
 several related papers are available today.
@@
 One can compare the result of the asymptotic expansion with other well known techniques like Edgeworth series expansion for the log-normal distribution as proposed, e.g.,  in \citet{Levy92}.
-This approximation is available through the package \pkg{fExoticOptions}.
+This approximation is available through the package \pkg{fExoticOptions} \citep{fExoticOptions}.
 %%% https://www.rocq.inria.fr/mathfi/Premia/free-version/doc/premia-doc/pdf_html/asian_doc/index.html#x1-80005.1.1
 <<>>=
-library(fExoticOptions)
+library("fExoticOptions")
 levy <- LevyAsianApproxOption(TypeFlag = "c", S = xinit, SA = xinit, X = K, 
@@
 
-\subsubsection{Asymptotic Expansion for General Stochastic Processes}
+\subsubsection{Asymptotic expansion for general stochastic processes}
 Of course, \pkg{yuima} approach is more general in that the above Levy approximation only holds when the process $X_t$ is the geometric Brownian motion.
@@
 
-\section{Inference for Stochastic Processes}\label{sec5}
+\section{Inference for stochastic processes}\label{sec5}
 The \pkg{yuima} implements several optimal techniques for parametric, semi- and non-parametric estimation of (multidimensional) stochastic differential equations.
-Most of the techniques presented below apply to high frequency data, i.e.
-when $\Delta_n$, the time lag between to consecutive observations of the process, converges to zero as  the number  $n$ of observations       increases.
-
-\subsection{How to Input Data Into a \code{yuima} Object}
+Most of the techniques presented below apply to high frequency data, i.e., when $\Delta_n$, the time lag between to consecutive observations of the process, converges to zero as  the number  $n$ of observations       increases.
+
+\subsection{How to input data into a \code{yuima} object}
 Although most of the examples in this section are given on simulated data, the main way to fill up the \code{data} slot of a \code{yuima} object is to use the function \code{setYuima}.
@@
 
-\subsection{Quasi Maximum Likelihood Estimation}\label{section-qmle}
+\subsection{Quasi maximum likelihood estimation}\label{section-qmle}
 Consider a multidimensional diffusion process
@@
 
-\subsubsection{QMLE in Practice}
+\subsubsection{QMLE in practice}
 
@@
 
-\subsubsection{Theoretical Remarks on QMLE}%, High-freq \& ergodicity}
+\subsubsection{Theoretical remarks on QMLE}%, High-freq \& ergodicity}
 Consistency of an estimator is always a  required property; otherwise 
@@
 
-\subsection{Adaptive Bayes Estimation}
+\subsection{Adaptive Bayes estimation}
 Consider again the  diffusion process in \eqref{eq:sdemle}
@@
 
-\subsubsection{Theoretical Remarks on Adaptive Bayes Estimator}
+\subsubsection{Theoretical remarks on adaptive Bayes estimator}
 Under the same conditions, the adaptive Bayes estimator $\tilde{\theta}_1$ and
@@
 
-\subsubsection{The Effect of Small Sample Size on Drift Estimation}\label{small.sample}
+\subsubsection{The effect of small sample size on drift estimation}\label{small.sample}
 It is known from the theory that the estimation of the drift in a diffusion process strongly depends on the length of the observation interval $[0,T]$.
@@
 
-\subsection{Asynchronous Covariance Estimation}
+\subsection{Asynchronous covariance estimation}
 Suppose that two It\^o processes are observed
@@
 
-\subsubsection{Example: Data Generation and Estimation by \pkg{yuima} Package}
+\subsubsection{Example: data generation and estimation by \pkg{yuima} package}
 We will demonstrate how to apply \code{cce} function to
@@
 \code{cce} takes the sample and returns 
-an estimate of the quadratic covariation. For example, for the complete data, we obtain the
+an estimate of the quadratic covariation. For example, for the complete data in Figure~\ref{fig:plot-cceplot1}, we obtain the
 following estimates
@@
 @
-<<cceplot1,fig=TRUE,width=9,height=5>>=
+<<cceplot1,fig=FALSE>>=
 plot(X,main="complete data")
 @
-
+<<plot-cceplot1,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-cceplot1.pdf",width=9,height=5)
+par(mar=c(4,4,0,0))
+plot(X,main="complete data")
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-cceplot1}
+\caption{The complete simulated data.}
+\label{fig:plot-cceplot1}
+\end{figure} 
 We now apply random sampling in the following way: we define a new sampling structure via \code{setSampling} specifying in the argument \code{random} 
@@
 @
-Now we use the \code{subsampling} function to subsample the original data \code{X} into new asynchronous data \code{Y}
+Now we use the \code{subsampling} function to subsample the original data \code{X} into new asynchronous data \code{Y} (see Figure~\ref{fig:plot-cceplot2})
 <<>>=
@@
 @ 
-<<cceplot2,fig=TRUE,width=9,height=5>>=
+<<cceplot2,fig=FALSE>>=
 plot(Y,main="asynchronous data")
 @
-
-\subsubsection{Asynchronous Estimation for Nonlinear Systems}
+<<plot-cceplot2,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-cceplot2.pdf",width=9,height=5)
+par(mar=c(4,4,0,0))
+plot(Y,main="asynchronous data")
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-cceplot2}
+\caption{The asynchronous data generated from the simulated ones using Poisson random sub-sampling.}
+\label{fig:plot-cceplot2}
+\end{figure}
+
+\subsubsection{Asynchronous estimation for nonlinear systems}
 Consider now the two-dimensional system with nonlinear feedback
@@
 @
-We apply the same Poisson random sampling so that the object \code{Y} will contain  asynchronous data
+We apply the same Poisson random sampling so that the object \code{Y} will contain  asynchronous data (see Figure~\ref{fig:plot-cceplot3})
 <<>>=
@@
 @
-<<cceplot3,fig=TRUE,width=9,height=5>>=
+<<cceplot3,fig=FALSE>>=
 plot(Y,main="asynchronous data (non linear case)")
 @
-
+<<plot-cceplot3,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-cceplot3.pdf",width=9,height=5)
+par(mar=c(4,4,0,0))
+plot(Y,main="asynchronous data (non linear case)")
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-cceplot3}
+\caption{The asynchronous data for the nonlinear system.}
+\label{fig:plot-cceplot3}
+\end{figure} 
 The estimated covariance for the complete trajectory \code{yuima}
@@
 
-\subsection{Change-Point Analysis}
+\subsection{Change-point analysis}
 Consider a multidimensional stochastic differential equation of the form
@@
 
-\subsubsection{Example  of Volatility Change-Point Estimation for 2-Dimensional SDE's}
+\subsubsection{Example  of volatility change-point estimation for 2-dimensional SDE's}
 One example of model that can be analyzed by this technique is, for example, the 2-dimensional stochastic differential equation
@@
 $\theta_{1.1}=0.2$ and  $\theta_{2.1}=0.4$.
-For this trajectory, the initial value is set to the last value of the first trajectory stored in \code{x1} and \code{x2} for the two component of the process.
+For this trajectory, the initial value is set to the last value of the first trajectory stored in \code{x1} and \code{x2} for the two component of the process (see Figure~\ref{fig:plot-cpoint4})
 <<cpoint3b,results=hide>>=
@@
 @ 
-
 The composed trajectory appears as follows
-<<cpoint4,fig=TRUE,width=9,height=5>>=
+<<cpoint4,fig=FALSE>>=
 plot(yuima)
 @
-
+<<plot-cpoint4,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-cpoint4.pdf",width=9,height=5)
+par(mar=c(4,4,0,0))
+plot(yuima)
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-cpoint4}
+\caption{The 2-dimensional trajectory with change point around $\tau=4$.}
+\label{fig:plot-cpoint4}
+\end{figure} 
 As said, the change-point analysis does not consider the information coming from the drift part of the model and \pkg{yuima} ignores this internally.
@@
 As it can be seen, the above estimates of the change point are the same for the complete model \code{yuima} and the model without drift \code{noDriftModel}.
-\subsubsection{An Example of Two Stage Estimation}
+\subsubsection{An example of two stage estimation}
 In practical situations, the initial values of the parameters are not known and there is the need to provide a preliminary estimators of them.
@@
 @
-\subsection{LASSO Model Selection}
-The Least Absolute Shrinkage and Selection Operator (LASSO) is a useful and well studied approach to the problem of model selection and its major advantage is the simultaneous execution of both parameter estimation and variable selection (see \citet{Tib96}; \citet{Knight00}; \citet{Efron}).
+\subsection{LASSO model selection}
+The Least Absolute Shrinkage and Selection Operator (LASSO) is a useful and well studied approach to the problem of model selection and its major advantage is the simultaneous execution of both parameter estimation and variable selection \citep{Tib96,Knight00,Efron}.
 
@@
 $$ Y = \theta_0 +  \theta_1 X_1+  \theta_2 X_2 + \cdots +  \theta_k X_k$$
-perform least squares estimation under $L_1$ constraints, i.e.
+perform least squares estimation under $L_1$ constraints, i.e.,
 
@@
 
-\subsubsection{An Example of Model Selection for Interest Rates Data}
+\subsubsection{An example of model selection for interest rates data}
 The \code{lasso} method is implemented in the \pkg{yuima} package.
@@
 $$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
-and let us try to estimate the parameter on  the U.S. Interest Rates monthly data from 06/1964 to 12/1989. We prepare the data, the model and the constraints for optimization
-<<results=hide>>=
-library(Ecdat)
-data(Irates)
+and let us try to estimate the parameter on  the U.S.
+Interest Rates monthly data from 06/1964 to 12/1989 (see Figure~\ref{fig:plot-irates}).
+We prepare the data, the model and the constraints for optimization
+<<results=hide,fig=FALSE>>=
+library("Ecdat")
+data("Irates")
 rates <- Irates[,"r1"]
@@
 where \code{lambda10} stores the penaly terms $\lambda_0$ and $\gamma_0$ of \eqref{eq:penalty} of the LASSO method.
+<<plot-irates,echo=FALSE, fig=FALSE, print=FALSE,results=hide>>=
+pdf("plot-irates.pdf",width=9,height=4)
+par(mar=c(4,4,0,0))
+plot(rates)
+dev.off()
+@
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{plot-irates}
+\caption{The U.S. Interest Rates monthly data from 06/1964 to 12/1989.}
+\label{fig:plot-irates}
+\end{figure} 
 Now we apply the \code{lasso} function
@@
 
-\section{Miscellanea and Roadmap of YUIMA Project}\label{sec6}
+\section{Miscellanea and roadmap of YUIMA project}\label{sec6}
 Other statistical techniques have been are already implemented in the developer version of the \pkg{yuima} package although not yet released into the current distribution.

### 2013-10-11 pkg/yuimadocs/inst/doc/JSS/final-edited.R
edit
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@252 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@ -1,0 +1,523 @@
+### R code from vignette source 'final-edited.Rnw'
+### Encoding: UTF-8
+
+
+##################################
+##################################
+##################################
+### Sec 3. Model specification ###
+##################################
+##################################
+##################################
+
+
+
+library("yuima")
+mod1 <- setModel(drift = "-3*x", diffusion = "1/(1+x^2)")
+str(mod1)
+set.seed(123)
+X <- simulate(mod1)
+plot(X)
+str(X,vec.len=2)
+
+mod1b <- setModel(drift = "-3*s*y", diffusion = "1/(1+y^2)",
+state.var="y", time.var="s")
+str(mod1b)
+
+mod2 <- setModel(drift = "-theta*x", diffusion = "1/(1+x^gamma)")
+str(mod2)
+set.seed(123)
+X <- simulate(mod2,true.param=list(theta=1,gamma=3))
+plot(X)
+
+# 2-dim Models
+
+sol <- c("x1","x2") # variable for numerical solution
+a <- c("-3*x1","-x1-2*x2")   # drift vector 
+b <- matrix(c("1","x1","0","3","x2","0"),2,3)  #  diffusion matrix
+mod3 <- setModel(drift = a, diffusion = b, solve.variable = sol)
+set.seed(123)
+X <- simulate(mod3)
+plot(X, plot.type="single",lty=1:2)
+
+# Non liner models
+mu <- 0.1
+sig <- 0.2
+rho <- -0.7
+g <- function(t) {0.4 + (0.1 + 0.2*t)* exp(-2*t)}
+
+f1 <- function(t, x1, x2, x3) {
+    ret <- 0
+    if(x1 > 0 && x2 > 0) ret <- x2*exp(log(x1)*2/3)
+    return(ret)
+}
+
+f2 <- function(t, x1, x2, x3) {
+     ret <- 0
+     if(x3 > 0) ret <- rho*sig*x3
+     return(ret)
+}
+
+f3 <- function(t, x1, x2, x3) {
+     ret <- 0
+     if(x3 > 0) ret <- sqrt(1-rho^2)*sig*x3
+     return(ret)
+}
+
+diff.coef.matrix <- matrix(c("f1(t,x1,x2,x3)",
+ "f2(t,x1,x2,x3) * g(t)", "f2(t,x1,x2,x3)", "0", 
+     "f3(t,x1,x2,x3)*g(t)", "f3(t,x1,x2,x3)"),  3, 2)
+
[…]
